### 2007-09-26 Developer/eqPly/chapters/abstract.tex
Added LaTeX sources of semester thesis about the new eqPly in /doc/Developer/eqPly.
[ ] May break build
[ ] Breaks existing applications
[ ] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[x] Documentation

% Copyright (c) 2007, Tobias Wolf <twolf@access.unizh.ch>
% All rights reserved.

%
% Abstrakt Datei
%

\clearpage
\null
\vfil

\begin{center}
\sffamily
\textbf{Zusammenfassung}
\end{center}
\normalfont
Equalizer ist ein Open Source Framework für verteiltes, paralleles Rendering. 
Es setzt auf OpenGL auf, einer plattformunabhängigen Programmierschnittstelle 
für Computergrafik. Im Rahmen dieser Semesterarbeit wird auf der Basis von 
Equalizer und OpenGL ein Programm zur effizienten Darstellung von polygonalen
Daten entwickelt.

In einem ersten Schritt wird eine effiziente Datenstruktur für die Verwaltung
der polygonalen Daten umgesetzt. In einem zweiten Schritt wird ein optisch
ansprechendes Rendering für die Darstellung dieser Daten implementiert. Das
Ergebnis dieser Arbeit ist eine neue Beispielapplikation für Equalizer.

\vfil

\begin{center}
\sffamily
\textbf{Abstract}
\end{center}
\normalfont
Equalizer is an open source framework for distributed, parallel rendering.
It is built upon OpenGL, a platform independant application programming 
interface for computer graphics. In the course of this semester thesis an
application for the efficient presentation of polygonal data is developed, 
based upon Equalizer and OpenGL.

In a first step, an efficient data structure for the storage of the polygonal
data is developed. In a second step, a visually appealing rendering for the 
presentation of that data is implemented. The result of this thesis is a new
example application for Equalizer.

\vfil
\clearpage

%
% EOF
%