### 2010-02-22 Developer/ProgrammingGuide/paper.tex
Programming Guide: add slave object commit section
[ ] May break build
[ ] May break existing applications
[x] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[ ] Documentation

\NeedsTeXFormat{LaTeX2e}
\documentclass[10pt,a4]{scrartcl}

\usepackage{tabularx,longtable,graphicx,a4,listings,wrapfig,subfigure,textcomp,ccaption}
\usepackage[absolute]{textpos}

\ifx\pdfoutput\undefined
  % We're not running pdftex
  % european (better) fonts -- does not look good with pdflatex
  \usepackage[T1]{fontenc}
  \newcommand{\href}[2]{#2\\{\hspace*{5mm}\scriptsize <#1>}\\}
\else
  \pdfcompresslevel=9
  \def\pdfBorderAttrs{/Border [0 0 0] } % No border around Links
  \usepackage{hyperref}
\fi

\title{Equalizer\\Programming and User Guide}
\author{Eyescale Software GmbH}

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications based on
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.7.4 for Equalizer 0.9.1+svn\\[\medskipamount]
%  \today
}

\newcommand{\tm}{\texttrademark~}
\newcommand{\rc}{\raise 1ex\hbox{{\tiny\textregistered}}~}
\newcommand{\fig}[1]{Figure~\ref{#1}}
\newcommand{\sref}[1]{Section~\ref{#1}}
\newcommand{\aref}[1]{Appendix~\ref{#1}}
\newcommand{\link}[1]{\htmladdnormallink{#1}{#1}}
\indentcaption{2em}

% suppress  single floating lines on top (widow) and bottom(club)
%  10000 is infinity
%  tradeoff: maybe underfull vboxes
\clubpenalty=10000
\widowpenalty=10000

\begin{document}

\pagestyle{empty}
\maketitle
\thispagestyle{empty}

%\clearpage

\lstset{language=C++}
\section*{Equalizer 0.9.1 Programming and User Guide}
\today
\subsection*{Contributors}

Written by Stefan Eilemann.\\
Engineering contributions by Maxim Makhinya, Jonas B\"osch, Christian
Marten, Sarah Amsellem, Patrick Bouchaud, Philippe Robert and Robert Hauck.

\subsection*{Copyright}

\textcopyright 2007-2010 \htmladdnormallink{Eyescale Software GmbH}
{http://www.eyescale.ch}. All rights reserved. No permission is
granted to copy, distribute, or create derivative works from the
contents of this electronic documentation in any manner, in whole or in
part, without the prior written permission of Eyescale Software GmbH.

\subsection*{Trademarks and Attributions}

OpenGL is a registered Trademark, OpenGL Multipipe is a Trademark of
Silicon Graphics, Inc. Linux is a registered Trademark of Linus
Torvalds.  Mac OS is a Trademark of Apple Inc. CAVELib is a registered
Trademark of the University of Illinois. The CAVE is a registered
Trademark of the Board of Trustees of the University of Illinois at
Chicago. Qt is a registered Trademark of Trolltech. TripleHead2Go is a
Trademark of Matrox Graphics. PowerWall is a Trademark of Mechdyne
Corporation. CUDA is a Trademark of NVIDIA Corporation. All other
trademarks and copyrights herein are the property of their respective
owners.

\subsection*{Feedback}

If you have comments about the content, accuracy or comprehensibility of
this Programming and User Guide, please contact
\htmladdnormallink{eile@eyescale.ch}
{mailto:eile@eyescale.ch?subject=Equalizer\%20Programming\%20Guide}.

\vfill

\subsection*{Front and Back Page}

The images on the front page show the following applications build using
Equalizer: 
RTT DeltaGen\footnote{Image copyright Realtime Technology AG, 2008} [top
right], 
\textsf{eqRaster}, a scalable terrain visualization\footnote{Image
  courtesy of Visualization and Multimedia Lab, University of Zurich}
[middle center],
a flow visualization application for climate research\footnote{Image
  courtesy of Computer Graphics and Multimedia Systems, University of
  Siegen} [middle right],
the \textsf{eqPly} polygonal renderer in a three-sided CAVE [bottom
left],
the \textsf{eVolve} volume renderer\footnote{Data set courtesy of
  General Electric, USA} [bottom center],
and the \textsf{eqNBody} gravitational n-body simulation, based on the
N-Body CUDA example [bottom right].

The images on the back page show the following scalable rendering modes:
Subpixel (FSAA) [top left], DPlex [middle left], Pixel [middle center],
2D [bottom left], DB [bottom center] and stereo [bottom right].

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.7.0   & Sep 09, 2009 & Updates based on hardcopy print\\
    1.7.1   & Oct 20, 2009 & Updates for new eq::GLWindow\\
    1.7.2   & Nov 18, 2009 & Add \sref{sSubpixel} and \sref{sFSAA} for Subpixel compounds\\
    1.7.3   & Feb 02, 2010 & Revision for Equalizer 0.9.1\\
    1.7.4   & Feb 22, 2010 & Add \sref{sSlaveCommit}\\
    \hline
  \end{tabularx}}
\clearpage

\pagenumbering{arabic}
\pagestyle{headings}

\part{User Guide}
\section{Introduction}

Equalizer is the standard middleware for the development and deployment
of parallel OpenGL applications. It enables applications to benefit from
multiple graphics cards, processors and computers to scale rendering
performance, visual quality and display size. An Equalizer-based
application runs unmodified on any visualization system, from a simple
workstation to large scale graphics clusters, multi-GPU workstations and
Virtual Reality installations.

This User and Programming Guide introduces parallel rendering concepts,
the configuration of Equalizer-based applications and programming using
the Equalizer parallel rendering framework.

Equalizer is the most advanced middleware for scalable 3D visualization,
providing the broadest set of parallel rendering features available in
an open source library to any OpenGL application. Many commercial and
open source applications in a variety of different markets rely on
Equalizer for flexibility and scalability.

Equalizer provides the domain-specific parallel rendering know-how and
abstracts configuration, threading, synchronization, windowing and event
handling. It is a `GLUT on steroids', providing parallel and distributed
execution, scalable rendering features, network data distribution and
fully customizable event handling.

If you have any question regarding Equalizer programming, this guide, or
other specific problems you encountered, please direct them to the
\textsf{eq-dev} mailing
list\footnote{\link{http://www.equalizergraphics.com/lists.html}}.

\subsection{Parallel Rendering}

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-1ex}\includegraphics[width=.618\textwidth]{images/executionFlow.pdf}
  {\caption{\label{fExecutionFlow}Parallel Rendering}}
  \vspace{-4ex}
\end{wrapfigure}

\fig{fExecutionFlow} illustrates the basic principle of any parallel
rendering application. The typical OpenGL application, for example GLUT,
has an event loop which redraws the scene, updates data based on
received events, and eventually redraws a new frame.

A parallel rendering application uses the same basic execution model and
extends it by separating the rendering code from the main event
loop. The rendering code is then executed in parallel on different
resources, depending on the configuration chosen at runtime.

This model is naturally followed by Equalizer, thus making application
development as easy as possible.

\subsection{Installing Equalizer and Running \textsf{eqPly}}

Equalizer can be installed by downloading the
distribution\footnote{\link{http://www.equalizergraphics.com/downloads.html}}
and compiling the source code. After installing Equalizer, please take a
look at the Quickstart
Guide\footnote{\link{http://www.equalizergraphics.com/documents/EqualizerGuide.html}}
to get familiar with the capabilities of Equalizer and the
\textsf{eqPly} example.

Compiling Equalizer is as simple as running \textsf{make} on Linux or
Mac OS X, building the provided XCode project on Mac OS X 10.5, or
building the Equalizer Visual Studio 2005 solution on Windows. Note that
on Mac OS X 10.4 (Tiger), some prerequisites have to be installed before
running \textsf{make}, as explained in \textsf{README.Darwin}.

\subsection{Equalizer Processes}

The Equalizer architecture is based on a client-server model. The client
library exposes all functionality discussed in this document to the
programmer, and provides communication between the different Equalizer
processes. The underlying network layer uses a peer-to-peer structure,
allowing render clients to communicate directly with each other, e.g.,
during parallel compositing.

\subsubsection{Server}
Each Equalizer server is responsible for managing one visualization
system, i.e., a shared memory system or graphics cluster.  It controls
and launches the application's rendering clients.  Currently, Equalizer
only supports one application per server, but it will provide
concurrent and efficient multi-application support in the future.

\subsubsection{Application}

The application connects to an Equalizer server and receives a
configuration.  Furthermore, the application also provides its render
client, which will be controlled by the server. The application reacts
on events, updates its data and controls the rendering.

\subsubsection{Render Clients}

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/processes.pdf}
  {\caption{\label{fProcesses}Equalizer Processes}}
\end{wrapfigure}
The render client implements the rendering part of an application. Its
execution is passive, it has no main loop and is completely driven by
Equalizer. It executes the rendering tasks received from the server by
calling the appropriate task methods (see \sref{sTaskMethods}) in the
correct thread and context. The application either implements the task
methods with application-specific code or uses the default methods
provided by Equalizer.

The application can also be a rendering client, in which case it can
also contribute to the rendering. If it does not implement any render
client code, it is reduced to be the application's `master' process
without any OpenGL windows and 3D rendering.

The rendering client can be the same executable as the application, as
it is the case with all provided examples. When it is started as a
render client, the Equalizer initialization routine does not return and
takes over the control by calling the render client task
methods. Complex applications usually implement a separate, lightweight
rendering client.

\section{Scalable Rendering}

Real-time visualization is an inherently parallel
problem. Unfortunately, different applications have different rendering
algorithms, which require different scalable rendering modes to address
the bottlenecks correctly. Equalizer supports all important algorithms as
listed below, and will continue to add new ones over time to meet
application requirements.

This section gives an introduction to scalable rendering, providing some
background for end users and application developers. The scalability
modes offered by Equalizer are discussed, along with their advantages
and disadvantages.

Choosing the right mode for the application profile is critical for
performance. Equalizer uses the concept of compounds to describe the
task decomposition and result recomposition. It allows the combination
of the different compound modes in any possible way, which allows to
address different bottlenecks in a flexible way.

\subsection{2D or Sort-First Compounds}

2D decomposes the rendering in screen-space, that is, each contributing
rendering unit processes a tile of the final view. The recomposition
simply assembles the tiles side-by-side on the destination view. This
mode is also known as sort-first or SFR.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/2D.pdf}
  {\caption{ A 2D Compound}}
\end{wrapfigure}
The advantage of this mode is a low, constant IO overhead for the pixel
transfers, since only color information has to be transmitted. The upper
limit is the amount of pixel data for the destination view.

Its disadvantage is that it relies on view frustum culling to reduce the
amount of data submitted for rendering. Depending on the application
data structure, the overlap of some primitives between individual tiles
limits the scalability of this mode, typically to around eight graphics
cards. Each node has to potentially hold the full database for
rendering.

2D decompositions can be used by all types of applications, but should
be combined with DB compounds to reduce the data per node, if
possible. In most cases, a \textsf{loadBalancer} should be used to
automatically adjust the tiling each frame, based on the current
rendering load.

2D compounds in Equalizer are configured using the \textsf{viewport}
parameter, using the values \textsf{[ x y width height]} in normalized
coordinates. The viewport defines the area of the parent (destination)
channel to be rendered for each child. Each child compound uses an
output frame, which is connected to an input frame on the destination
channel. The destination channel can also be used as a source channel,
in which case it renders in place and no output frame is needed.

\subsection{DB or Sort-Last Compounds}

DB, as shown in \fig{fDB}\footnote{3D model courtesy of AVS, USA.},
decomposes the rendered database so that all rendering units process a
part of the scene in parallel. This mode is also known as sort-last, and
is very similar to the data decomposition approach used by HPC
applications.

Volume rendering applications use an ordered alpha-based blending to
composite the result image. The depth buffer information is used to
composite the individual images correctly for polygonal data.

This mode provides very good scalability, since each rendering unit
processes only a part of the database. This allows to lower the
requirements on all parts of the rendering pipeline: main memory usage,
IO bandwidth, GPU memory usage, vertex processing and fill rate.

Unfortunately, the database recomposition has linear increasing IO
requirements for the pixel transfer. Parallel recomposition algorithms,
such as direct-send address this problem by keeping the per-node IO
constant (see \fig{fDirectSend}).

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/DB.pdf}
  {\caption{\label{fDB}\small A Database Compound}}
\end{wrapfigure}
The application has to partition the database so that the rendering units
render only part of the database. Some OpenGL features do not work
correctly (anti-aliasing) or need special attention (transparency).

The best use of database compounds is to divide the data to a manageable
size, and then to use other decomposition modes to achieve further
scalability. Volume rendering is one of the applications which can
profit from database compounds.

DB compounds in Equalizer are configured using the \textsf{range}
parameter, using the values \textsf{[ begin end ]} in normalized
coordinates. The range defines the start and end point of the
application's database to be rendered. The value has to be interpreted
by the application's rendering code accordingly.  Each child compound
uses an output frame, which is connected to an input frame on the
destination channel. For more than two contributing channels, it is
recommended to configure streaming or parallel direct send compositing,
as described in \sref{sDirectSend}.

\subsection{Stereo Compounds}

Stereo compounds, as shown in \fig{fStereoCmp}\footnote{3D model
  courtesy of Stereolithography Archive at Clemson University.}, assign
each eye pass to individual rendering units. The resulting images are
copied to the appropriate stereo buffer. This mode supports a variety of
stereo modes, including active (quad-buffered) stereo, anaglyphic stereo
and auto-stereo displays with multiple eye passes.

Due to the frame consistency between the eye views, this modes scales
very well. The IO requirements for pixel transfer are small and
constant.

The number of rendering resources used by stereo compounds is limited by
the number of eye passes, typically two.

Stereo compounds are used by all applications when rendering in stereo,
and is often combined with other modes.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/EYE.pdf}
  {\caption{\label{fStereoCmp}\small A Stereo Compound}}
\end{wrapfigure}
Eye compounds in Equalizer are configured using the \textsf{eye}
parameter, limiting the child to render the \textsf{[ LEFT ]} or
\textsf{[ RIGHT ]} eye. Each child compound uses an output frame, which
is connected to an input frame on the destination channel. The
destination channel can also be used to render an eye pass, in which case
it renders in the correct stereo buffer and no output frame is needed.

\subsection{\label{sDPlex}DPlex Compounds}

DPlex compounds assign full, alternating frames to individual rendering
units. The resulting images are copied to the destination channel, and
Equalizer load-balancing is used to ensure a steady framerate on the
destination window. This mode is also known as time-multiplex or AFR.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/DPlex.pdf}
  {\caption{ A DPlex Compound}}
\end{wrapfigure}

Due to the frame consistency between consecutive frames, this mode scales
very well. The IO requirements for pixel transfer are small and
constant.

DPlex requires a latency of at least n frames. This increased latency
might be disturbing to the user, but it is often compensated by the
higher frame rate. The frame rate typically increases linearly with the
number of source channels, and therefore linearly with the latency.

DPlex compounds in Equalizer are configured using the \textsf{period}
and \textsf{phase} parameter, limiting each child to render a subset of
the frames. Each child compound uses an output frame, which is connected
to an input frame on the destination channel. The destination channel
uses a DPlex load-balancer to smoothen the framerate.

\subsection{Pixel Compounds}

Pixel compounds are similar to 2D compounds. The frusta of the
source rendering units are modified so that each unit renders an evenly
distributed subset of pixels, as shown in \fig{fPixel}\footnote{3D model courtesy of AVS, USA.}

As 2D compounds, pixels compounds have low, constant IO requirements for
the pixel transfers during recomposition.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/Pixel.pdf}
  {\caption{\label{fPixel}\small A Pixel Compound}}
\end{wrapfigure}
OpenGL functionality influenced by the raster position will not work
correctly with pixel compounds, or needs at least special
attention. Among them are: lines, points, sprites, glDrawPixels,
glBitmap, glPolygonStipple. The application can query the current pixel
parameters at runtime to adjust the rendering accordingly.

Pixel compounds work well for purely fill-limited applications.
Techniques like frustum culling do not reduce the rendered data for the
source rendering resources. Pixel compounds are ideal for ray-tracing,
which is highly fill-limited and needs the full database for rendering
anyway. Volume rendering applications are also well suited for this
mode, and should choose it over 2D compounds.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/pixelKernel.pdf}
  {\caption{\label{fPixelKernel}\small Pixel Compound Kernel}}
\end{wrapfigure}
Pixel compounds in Equalizer are configured using the \textsf{pixel}
parameter, using the values \textsf{[ x y width height]}  to configure
the size and offset of the sampling kernel. The width and height of the
sampling kernel define how many pixels are skipped in the x and y
direction, respectively. The x and y offset define the index of the
source channel within the kernel. They have to be smaller than the size
of the kernel. \fig{fPixelKernel} illustrates these parameters, and
\fig{fPixelKernels} shows some example kernels for a four-to-one pixel decomposition.

The destination channel can also be used as a source channel. Contrary
to the other compound modes, it also has to use an output and
corresponding input frame. During rendering, the frustum is 'squeezed'
to configure the pixel decomposition. The destination channel can
therefore not be rendered in place, like with the other compound modes.
\begin{figure}[ht!]\center
  \includegraphics[width=.9\textwidth]{images/pixelKernels.pdf}
  {\caption{\label{fPixelKernels}Example Pixel Kernels for a
      four-to-one Pixel Compound}}
\end{figure}

\subsection{\label{sSubpixel}Subpixel Compounds}

Subpixel decomposes the rendering of multiple samples for one pixel, e.g., for
anti-aliasing or depth-of-field rendering. The default implementation provides
transparent, scalable software idle-antialiasing when configuring a subpixel
compound.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/Subpixel.pdf}
  {\caption{\label{fSubpixel}\small A Subpixel Compound}}
\end{wrapfigure}
Applications can use subpixel compounds to accelerate depth-of-field effects and
software anti-aliasing, with potentially a different number of idle or non-idle
samples per pixel.

As for the DB compound, the subpixel recomposition has linear increasing 
IO requirements for the pixel transfer with the difference that only 
color information has to be transmitted.

Subpixel compounds are configured using the \textsf{subpixel} parameter, using
the values \textsf{[ index size ]}. The \textsf{index} parameter corresponds to
the current resource index and the \textsf{size} is the total number of
resources used.

The \textsf{index} has to be smaller than the \textsf{size}. 
\fig{fSubpixel} illustrates a three-way subpixel decomposition.

The destination channel can also be used as a source channel. 
As for the pixel compound, it has to use an output and corresponding 
input frame.

\subsection{\label{sLoadBalancing}Automatic Runtime Adjustments}

Some scalable rendering parameters are updated at runtime to provide
optimal performance or influence other scalable rendering
features. These adjustments are often referred to as load-balancing, but
are called \textsf{equalizers}, since their functionality is not limited
to load-balancing in Equalizer.

The server provides a number of equalizers, which automatically update
certain parameters of compounds based on runtime information. They
balance the load of 2D and DB compounds, optimally distribute render
resources for segment projections systems, adjust the resolution
to provide a constant framerate or zoom images to allow monitoring of
another view.

Equalizers are described in more detail in \sref{sEqualizers}.

\section{\label{sConfig}Writing Configurations}

\subsection{Preparation}

Before writing a configuration, it is useful to assemble the following
information:

\begin{itemize}
\item A list of all \textbf{computers} in your rendering cluster,
  including the \textbf{IP addresses} of all network interfaces to be
  used.
\item The number of \textbf{graphics cards} in each computer.
\item The \textbf{physical dimensions} of the display system, if
  applicable. These are typically the bottom-left, bottom-right and
  top-left corner points of each display surface in meters.
\item The relative coordinates of all the \textbf{segments} belonging to
  each display surface, and the graphics card \textbf{output} used for
  each segment. For homogenous setups, it is often enough to know the
  number of rows and columns on each surface, as well as the overlap or
  underlap percentage, if applicable.
\item The number of desired \textbf{application windows}. Application
  windows are typically destination windows for scalable rendering or
  'control' windows paired with a view on a display system.
\item \textbf{Characteristics} of the application, e.g., supported
  scalability modes and features.
\end{itemize}

\subsection{Overview}

Equalizer applications are configured at runtime by the Equalizer
server. The server loads its configuration from a text file, which is a
one-to-one representation of the configuration data structures at
runtime.

For an extensive documentation of the file format please refer to
\aref{aFileFormat}. This section gives an introduction on how to write
configuration files.

A configuration consists of the declaration of the rendering resources
and the usage description for these resources. It can optionally contain
the description of the physical layout of the projection system and
logical layouts on the projection canvases.

The rendering resources are represented in a hierarchical tree structure
which corresponds to the physical and logical resources found in a 3D
rendering environment: nodes (computers), pipes (graphics cards),
windows, channels.

The resource usage is configured using a compound tree, which is a
hierarchical representation of the rendering decomposition and
recomposition across the resources. It is explained in \sref{sCompounds}.

Physical layouts of display systems are configured using canvases with
segments, which represent 2D rendering areas composed of multiplex
displays or projectors. Logical layouts are applied to canvases and
define views on a canvas.

\begin{figure}[ht!]\center
  \includegraphics[width=\textwidth]{images/cave.pdf}
  {\caption{\label{fConfig}An Example Configuration}}
\end{figure}

\fig{fConfig} shows an example configuration for a four-side
CAVE, running on two machines (nodes) using three graphics
cards (pipes) with one window each to render to the four output channels
connected to the projectors for each of the walls. The compound
description is only used by the server to compute the rendering
tasks. The application is not aware of compounds, and does not need to
concern itself with the parallel rendering logics of a configuration.

For testing and development purposes it is possible to use multiple
instances for one resource, e.g. to run multiple render client nodes on
one computer. For optimal performance during deployment, one node and
pipe should be used for each computer and graphics card, respectively.

\subsection{Node}

For each machine in your cluster, create one \textsf{node}. Create one
\textsf{appNode} if your application process should also render, which
is often the case. List all nodes, even if you are not planning to use
them at first. Equalizer will only instantiate and access used nodes,
that is, nodes which are referenced by an active compound.

In each node, list all \textsf{connection}s through which this node is
reachable. Typically a node uses only one connection, but it is possible
to configure multiple connections if the machine and cluster is set up
to use multiple, independent network interfaces. Make sure the
configured \textsf{hostname} is reachable from all nodes. An IP address
may be used as the hostname.

The \textsf{eq::Node} class is the representation of a single computer
in a cluster. One operating system process of the render client will be
used for each node. Each configuration might also use an application
node, in which case the application process is also used for
rendering. All node-specific task methods are executed from the main
application thread.

\subsection{Pipe}

For each node, create one \textsf{pipe} for each graphics card in the
machine. Set the \textsf{device} number to the correct index. On
operating systems using X11, e.g., Linux, also set the port number if
your X-Server is running on a nonstandard port.

The \textsf{eq::Pipe} class is the abstraction of a graphics card (GPU).
In the current implementation it is also one operating system
thread. Non-threaded pipes are supported for integrating with
thread-unsafe libraries, but have various performance caveats. They
should only be used if using a different, synchronized rendering thread
is not an option.

All pipe, window and channel task methods are executed from the pipe
thread, or in the case of non-threaded pipes from the main application
thread\footnote{see
  \link{http://www.equalizergraphics.com/documents/design/nonthreaded.html}}.

Further versions of Equalizer might introduce threaded windows, where
all window-related task methods are executed in a separate operating
system thread.

\subsection{Window}

Configure one \textsf{window} for each desired application window on the
appNode. Configure one full-screen window for each display
segment. Configure one off-screen window, typically a pbuffer, for each
graphics card used as a source for scalable rendering. Provide a useful
name to each on-screen window if you want to easily identify it at runtime.

Sometimes display segments cover only a part of the graphics card
output. In this case it is advised to configure a non-fullscreen window
without window decorations, using the correct window viewport.

The \textsf{eq::Window} class encapsulates a drawable and an OpenGL
context. The drawable can be an on-screen window or an off-screen
PBuffer or FrameBuffer Object (FBO). The window uses an
\textsf{eq::OSWindow}, which abstracts and manages
window-system-specific handles to the drawable and context, e.g., an X11
window \textsf{XID} and \textsf{GLXContext} for the glX window system.

\subsection{Channel}

Configure one \textsf{channel} for each desired rendering area in each
window. Typically one full-screen channel per window is used. Name the
channel using a unique, easily identifiable name, e.g., 'source-1',
'control-2' or 'segment-2\_3'.

Multiple channels in application windows may be used to view the model
from different viewports.

Sometimes, a single window is split across multiple projectors, e.g., by
using an external splitter such as the Matrox TripleHead2Go. In this
case configure one channel for each segment, using the channel's
\textsf{viewport} to configure its position relative to the window.

The \textsf{eq::Channel} class is the abstraction of an OpenGL viewport
within its parent window. It is the entity executing the actual
rendering. The channel's viewport is overwritten when it is rendering
for another channel during scalable rendering.

\subsection{\label{sCanvas}Canvases}

If you are writing a configuration for workstation usage you can skip
the following sections and restart with \sref{sCompounds}.

Configure one \textsf{canvas} for each display surface. For planar
surfaces, e.g., a display wall, configure a frustum. For non-planar
surfaces, the frustum will be configured on each display segment.

The frustum can be specified as a wall or projection description. Take
care to choose your reference system for describing the frustum to be
the same system as used by the head-tracking matrix calculated by the
application.

A wall is completely defined by the bottom-left, bottom-right and
top-left coordinates relative to the origin.

A projection is defined by the position and head-pitch-roll orientation
of the projector, as well as the horizontal and vertical field-of-view
and distance of the projection wall.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/frusta.pdf}
  {\caption{\label{fFrusta}Wall and Projection Parameters}}
\end{wrapfigure}

\fig{fFrusta} illustrates the wall and projection frustum parameters.

Canvases provide a convenient way to configure projection surfaces. A
canvas uses layouts, which describe logical views. The usage of canvases
and layouts is optional. Typically, each desktop window uses one canvas.

A canvas represents one logical projection surface, e.g., a PowerWall, a
curved screen or an immersive installation. One configuration might
drive multiple canvases, for example an immersive installation and an
operator station.

A canvas consists of one or more segments. A planar canvas typically has
a frustum description (see \sref{sFrustum}), which is inherited by the
segments. Non-planar frusta are configured using the segment
frusta. These frusta typically describe a physically correct display
setup for Virtual Reality installations.

A canvas has one or more layouts. One of the layouts is the active
layout, that is, this set of views is currently used for rendering. It
is possible to specify \textsf{OFF} as a layout, which deactivates the
canvas. It is possible to use the same layout on different canvases.

\subsubsection{Segments}

Configure one \textsf{segment} for each display or projector of each
canvas. Configure the \textsf{viewport} of the segment to match the area
covered by the segment on the physical canvas. Set the output
\textsf{channel} to the resource driving the described projector.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/canvas.pdf}
  {\caption{\label{fCanvas}A Canvas using four Segments}}
\end{wrapfigure}

For non-planar displays, configure the frustum as described in
\sref{sCanvas}. For passive stereo installations, configure one segment
per eye pass, where the segment for the left and right eye have the same
viewport.

A segment represents one output channel of the canvas, e.g., a projector
or an LCD. A segment has an output channel, which references the channel to
which the display device is connected.

A segment covers a part of its parent canvas, which is configured using
the segment viewport. The viewport is in normalized coordinates with
respect to the canvas. Segments might overlap (edge-blended projectors)
or have gaps between each other (display walls,
\fig{fCanvas}\footnote{Dataset courtesy of VolVis distribution of SUNY
  Stony Brook, NY, USA.}). The viewport is used to configure the
segment's default frustum from the canvas frustum description, and to
place layout views correctly.

\subsection{\label{sLayout}Layouts}

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/layout.png}
  {\caption{\label{fLayout}A Layout with four Views}}
\end{wrapfigure}

Configure one \textsf{layout} for each configuration of logical
views. Name the layout using a unique name. Often only one layout with a
one view is used for all canvases.

Enable the layout on each desired canvas by adding it to the
canvas. Since canvases reference layouts by name or index, layouts have
to be configured before their respective canvases in the configuration
file.

A layout is the grouping of logical views. It is used by one or more
canvases. For all given layout/canvas combinations, Equalizer creates
destination channels when the configuration file is loaded. These
destination channels can be referenced by compounds to configure
scalable rendering.

\begin{figure}[ht!]\center
  \includegraphics[width=.9\textwidth]{images/wallLayout.jpg}
  {\caption{A Display Wall using a six-Segment Canvas with a two-View Layout}}
\end{figure}

Layouts can be switched at runtime by the application. Switching a
layout will activate different destination channels for rendering.

\subsubsection{Views}

Configure one \textsf{view} for each logical view in each layout. Set
the \textsf{viewport} to position the view.

A view is a logical view of the application data, in the sense used by
the Model-View-Controller pattern. It can be a scene, viewing mode,
viewing position, or any other representation of the application's data.

A view has a fractional viewport relative to its layout.  A layout
is often fully covered by its views, but this is not a requirement.

Each view can have a frustum description. The view's frustum overrides
frusta specified at the canvas or segment level. This is typically used
for non-physically correct rendering, e.g., to compare two models
side-by-side on a canvas. If the view does not specify a frustum, it
will use the sub-frustum resulting from the covered area on the canvas.

A view might have an observer, in which case its frustum is tracked by
this observer.

\subsection{Observers}

Unless you have multiple tracked persons, or want to disable tracking on
certain views, you can skip this section.

Configure one \textsf{observer} for each tracked person in the
configuration. Most configurations have at most one observer. Assign the
observer to all views which belong to this observer. Since the observer
is referenced by its name or index, it has to be specified before the
layout in the configuration file.

Views with no observer are not tracked. The config file loader will
create one default observer and assign it to all views if the
configuration has no observer.

An observer represents an actor looking at multiple views. It has a head
matrix, defining its position and orientation within the world, and an
eye separation. Typically, a configuration has one
observer. Configurations with multiple observers are used if multiple,
head-tracked users are in the same configuration session, e.g., a
non-tracked control host with two tracked head-mounted displays.

\subsection{\label{sCompounds}Compounds}

Compound trees are used to describe how the rendering resources are
combined to produce the desired output, and how they are aggregated to
increase the performance.

It is advised to study and understand the basic configuration files
shipped with the Equalizer configuration, before attempting to write
compound configurations. The command line program \textsf{configTool},
shipped with the Equalizer distribution, creates some standard
configurations automatically.

For configurations using canvases and layouts without scalability, the
configuration file loader will create the appropriate compounds. It is
typically not necessary to write compounds for this use case.

The following subsection outlines the basic approach to writing
compounds. The remaining subsections provide an in-depth explanation of
the compound structure to give the necessary background for compound
configuration.

\subsubsection{Writing Compounds}
The following steps are typically taken when writing compound
configurations:
\begin{description}
\item [Root compound] Define an empty top-level compound when
  synchronizing multiple destination views. Multiple destination views
  are used for multi-display systems, e.g., a PowerWall or CAVE. All
  windows used for one display surface should be swap-locked (see below)
  to provide a seamless image. A single destination view is typically
  used for providing scalability to a single workstation window.
\item [Destination compound(s)] Define one compound for each destination
  channel, either as a child of the empty group, or as a top-level
  compound.
  \begin{description}
  \item[Standalone channels] Set the channel by using the channel name
    from the resource section. Define a frustum as explained in
    \sref{sCanvas}.
  \item[View/Segment channels] Set the channel by using the canvas,
    segment, layout and view name. The compound frustum will be
    calculated automatically based on the segment or view frustum. Note
    that one segment may created multiple view/segment channels, one for
    each view intersection of each layout used on the canvas. Only the
    compounds belonging to the active layout of a canvas are activate at
    runtime.
    \begin{description}
    \item [Software swap lock] When using software swap synchronization,
      swap-lock the destination compounds using a swapbarrier. All
      windows with a swapbarrier of the same name synchronize their
      swapbuffers.
    \item [Hardware swap lock] When using hardware swap synchronization,
      use swapbarriers for the destination compounds, setting
      \textsf{NV\_group} and \textsf{NV\_barrier} appropriately. The
      swap barrier name is ignored in this case. All windows of the same
      group on a single node synchronize their swap buffer. All groups
      of the same barrier synchronize their swap buffer across
      nodes. Please note that the driver typically limits the number of
      groups and barriers to one, and that multiple windows per swap
      group are not supported by all drivers. Please refer to your
      OpenGL driver documentation for details.
    \end{description}
  \end{description}
\item[Scalability] If desired, define scalability for each of your
  destination compounds. Add one compound using a source channel for
  each contributor to the rendering. The destination channel may also be
  used as a source. 
  \begin{description}
  \item[Decomposition] On each child compound, limit the rendering task of that
    child by setting the \textsf{viewport}, \textsf{range}, \textsf{period} and
    \textsf{phase}, \textsf{pixel}, \textsf{subpixel}, \textsf{eye} or
    \textsf{zoom} as desired.
  \item[Runtime Adjustments] A \textsf{load\_equalizer} may be used on
    the destination compounds to set the \textsf{viewport} or
    \textsf{range} of all children each frame, based on the current
    load. A \textsf{view\_equalizer} may be used on the root compound to
    assign resources to all destination compounds, which have to use
    \textsf{load\_equalizer}s. A \textsf{framerate\_equalizer} should be
    used to smoothen the framerate of DPlex compounds. A
    \textsf{DFR\_equalizer} may be used to set the zoom of a compound to
    achieve a constant framerate. One compound may have multiple
    equalizers, e.g., a \textsf{load\_equalizer} and a
    \textsf{DFR\_equalizer} for a 2D compound with a constant framerate.
  \item[Recomposition] For each source compound, define an
    \textsf{output\_frame} to read back the result. Use this output frame
    as an \textsf{input\_frame} on the destination compound. The frames
    are connected with each other by their name, which has to be unique
    within the root compound tree. For parallel compositing, describe
    your algorithm by defining multiple input and output frames across
    all source compounds.
  \end{description}
\end{description}

\subsubsection{Compound Channels}
Each compound has a channel, which is used by the compound to execute
the rendering tasks. One channel might be used by multiple
compounds. Unused channels, windows, pipes and nodes are not
instantiated during config initialization. The rendering tasks for the
channels are computed by the server and send to the appropriate render
client nodes at the beginning of each frame.

\subsubsection{\label{sFrustum}Frustum}

Compounds have a frustum description to define the physical layout of
the display environment. The frustum specification is described in
\sref{sCanvas}.

The frustum description is inherited by the children, therefore the
frustum is typically defined on the topmost compound.

\subsubsection{Compound Classification}
The channels of the leaf compounds in the compound tree are designated
as source channels. The topmost channel in the tree is the destination
channel. One compound tree might have multiple destination channels,
e.g., for a swap-synchronized immersive installation.

All channels in a compound tree work for the destination channel. The
destination channel defines the 2D pixel viewport rendered by all leaf
compounds. The destination channel and pixel viewport cannot be
overridden by child compounds.

\subsubsection{Tasks}
Compounds execute a number of tasks: clear, draw, assemble and
readback. By default, a leaf compound executes all tasks and a non-leaf
compound assemble and readback. A non-leaf compound never executes the
draw task.

A compound can be configured to execute a specific set of tasks, for
example to configure the multiple steps used by binary-swap compositing.

\subsubsection{Decomposition - Attributes}
Compounds have attributes which configure the decomposition of the destination
channel's rendering, which is defined by the viewport, frustum and database. A
\textsf{viewport} decomposes the destination channel and frustum in screen
space. A \textsf{range} tells the application to render a part of its database,
and an \textsf{eye} rendering pass can selectively render different stereo
passes. A \textsf{pixel} parameter adjusts the frustum so that the source
channel renders an even subset of the parent's pixels. A \textsf{subpixel}
parameter tells the source channels to render different samples for one pixel to
perform anti-aliasing or depth-of-field rendering. Setting one or multiple
attributes causes the parent's view to be decomposed accordingly. Attributes are
cumulative, that is, intermediate compound attributes affect and therefore
decompose the rendering of all their children.

\subsubsection{Recomposition - Frames}
Compounds use output and input frames to configure the recomposition of
the resulting pixel data from the source channels. An output frame
connects to an input frame of the same name. The selected frame buffer
data is transported from the output channel to the input channel. The
assembly routine of the input channel will block on the availability of
the output frame. This composition process is extensively described in
\sref{sCompositing}. Frame names are only valid within the compound
tree, that is, an output frame from one compound tree cannot be used as
an input frame of another compound tree.

\subsubsection{\label{sSwapBarrier}Synchronization - Swapbarriers}
Compounds may have a software or hardware swapbarrier to synchronize the
buffer swap of multiple channels. A software swapbarrier is configured
by giving it a name. Windows with a swapbarrier of the same name
synchronize with each other before executing the swap buffer
task. Before entering the barrier, \textsf{Window::finish} is called to
ensure that all OpenGL commands have been executed.

Note that a software swap barrier is not accurate enough for
edge-blended projection system or immersive installations. Such
installations typically use a hardware synchronization, e.g., nVidia
G-Sync cards. It is however sufficient for display walls made out of
LCD's and for some non-immersive projection systems.

Swapbarrier names are only valid within the compound tree, that is, a
compound from one compound tree cannot be synchronized with a compound
from another compound tree.

A hardware swapbarrier is configured by setting the \textsf{NV\_group}
and \textsf{NV\_barrier} parameters. These parameters follow the
\textsf{NV\_swap\_group} extension, which synchronizes all windows bound
to the same group on a single machine, and all groups bound to the same
barrier across systems.

\subsubsection{\label{sEqualizers}Adjustments - Equalizers}

Equalizers are used to update compound parameters based on runtime
data. They are attached to a compound (sub-)tree, on which they
operate. The Equalizer distribution contains numerous example
configuration files using equalizers.

\paragraph{Load Equalizer}
\vspace{-2ex}\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/lb.pdf}
  {\caption{\label{fLoadBalancing}\small 2D Load-Balancing}}
\end{wrapfigure}
While pixel, subpixel and stereo compounds are naturally load-balanced, 2D and
DB compounds often need load-balancing for optimal rendering performance. Using
a load equalizer is transparent to the application, and can be used on any
application for 2D, and on most applications for DB load-balancing. Some
applications do not support dynamic updates of the database range, and therefore
cannot be used with DB load-balancing.

Using a 2D or DB load-balancer will adjust the 2D split or database range
automatically each frame. The 2D load-balancer exists in three flavors:
\textsf{2D} using tiles, \textsf{VERTICAL} using columns and
\textsf{HORIZONTAL} using rows.

2D load-balancing increases the framerate over a static decomposition in
virtually all cases. It works best if the application data is relatively
uniformly distributed in screen-space. A damping parameter can be used
to fine-tune the algorithm.

DB load-balancing is beneficial for applications which cannot precisely
predict the load for their scene data, e.g., when the data is
nonuniform. Volume rendering is a counterexample, where the data is
uniform and a static DB decomposition typically results in a better
performance.

\paragraph{View Equalizer}

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/viewLB.pdf}
  {\caption{\label{fViewLoadBalancing}\small Cross-Segment
      Load-Balancing for two Segments using eight GPU's}}
\end{wrapfigure}
Depending on the model position and data structure, each segment of a
multi-display system has a different rendering load. The segment with
the biggest load determines the overall performance when using a static
assignment of resources to segments. The view equalizer analyzes the
load of all segments, and adjusts the resource usage each frame. It
equalizes the load on all segments of a view.

\fig{fViewLoadBalancing}\footnote{Image Copyright Realtime Technology
  AG, 2008} illustrates this process. On the left side, a static
assignment of resources to display segments is used. The right-hand
segment has a higher load than the left-hand segment, causing
sub-optimal performance. The configuration on the left uses a view
equalizer, which assigns two GPU's to the left segment and four GPU's to
the right segment, which leads to optimal performance for this model and
camera position.

\begin{wrapfigure}{r}{.382\textwidth} 
    \includegraphics[width=.382\textwidth]{images/viewLB.png}
    {\caption{\label{fViewLoadBalancing2}\small Cross-Segment
        Load-Balancing for a CAVE}}
\end{wrapfigure}
The view equalizer can also use resources from another display resource,
if this resource has little rendering load by itself. It is therefore
possible to improve the rendering performance of a multi-display system
without any additional resources. This is particularly useful for
installations with a higher number of displays where the rendering load
is typically in a few segments only, e.g., for a CAVE.

\fig{fViewLoadBalancing2} shows cross-usage for a five-sided CAVE driven
by five GPU's. The front and left segments show the model and have a
significant rendering load. The view equalizer assigns the GPU's from
the top, bottom and right wall for rendering the left and front wall in
this configuration.

Cross-segment load-balan\-cing is configured hierarchically. On the top
compound level, a view equalizer assigns resources to each of its
children, so that the optimal number of resources is used for each
segment. On the next level, a load equalizer on each child computes the
resource distribution within the segment, taking the resource usage
given by the view equalizer into account.

\paragraph{Framerate Equalizer}
Certain configurations, in particular DPlex compounds, require a
smoothing of the framerate at the destination channel, otherwise the
framerate will become periodically faster and slower. Using a framerate
equalizer will smoothen the swap buffer rate on the destination window
for optimal user experience.

\paragraph{DFR Equalizer}

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/dfr.pdf}
  {\caption{\label{fDFR}Dynamic Frame Resolution}}
\end{wrapfigure}
Dynamic Frame Resolution (DFR) trades rendering performance for visual
quality. The rendering for a channel is done at a different resolution
than the native channel resolution to keep the framerate constant. The
DFR equalizer adjusts the zoom of a channel, based on the target and
current framerate. It is typically used for fill-rate bound
applications, such as volume rendering and ray-tracing.

\fig{fDFR}\footnote{Data set courtesy of Olaf Ronneberger, Computer
  Science Institute, University of Freiburg, Germany} shows DFR for
volume rendering. To achieve 10 frames per second, the model is rendered
at a lower resolution, and upscaled to the native resolution for
display. The rendering quality is slightly degraded, while the rendering
performance remains interactive. When the application is idle, it
renders a full-resolution view.

The dynamic frame resolution is not limited to downscaling the
rendering resolution, it will also upscale the resolution if the
parameters allow for it. Upscaled rendering, which will down-sample the
result for display, provides dynamic anti-aliasing at a constant
framerate.

\paragraph{Monitor Equalizer}
The monitor equalizer allows the observation of another view,
potentially made of multiple segments, in a different channel at a
different resolution. This is typically used to reuse the rendering of
a large-scale display on an operator station.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/monitorEq.pdf}
  {\caption{\label{fMonitorEq}Monitoring a Projection Wall}}
\end{wrapfigure}
A monitor equalizer adjusts the frame zoom of the output frames used to
observe the rendering, depending on the destination channel size. The
output frames are downscaled on the GPU before readback, which results
in optimal performance.

\fig{fMonitorEq} shows a usage of the monitor equalizer. A two-segment
display wall is driven by a separate control station. The rendering
happens only on the display wall, and the control window receives the
correctly downscaled version of the rendering.

\clearpage
\part{Programming Guide}

To modify an application for Equalizer, the programmer structures the
source code so that the OpenGL rendering can be executed in parallel,
potentially using multiple processes for cluster-based execution.

\section{Hello, World!}

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/eqHello.png}
  {\caption{\label{fHello}Hello, World!}}
\end{wrapfigure}
The \textsf{eqHello} example is a minimal application to illustrate the
basic principle of an Equalizer application: The application developer
has to implement the rendering method \textsf{Channel::frameDraw},
similar to the \textsf{glutDisplayFunc} in GLUT applications. It can be
run as a stand-alone application from the command line.

The \textsf{eqHello} redraw function renders six colored quads, rotating
around the origin. The \textsf{frameDraw} meth\-od provided by the
\textsf{eq::Channel} can be used as a convenience function to setup the
frustum and other OpenGL state. After setting up some lighting
parameters, \textsf{eqHello} rotates the scene and renders the quads
using immediate mode:

{\footnotesize\begin{lstlisting}
void Channel::frameDraw( const uint32_t spin )
{
    // setup OpenGL State
    eq::Channel::frameDraw( spin );

    const float lightPos[] = { 0.0f, 0.0f, 1.0f, 0.0f };
    glLightfv( GL_LIGHT0, GL_POSITION, lightPos );

    const float lightAmbient[] = { 0.2f, 0.2f, 0.2f, 1.0f };
    glLightfv( GL_LIGHT0, GL_AMBIENT, lightAmbient );

    // rotate scene around the origin
    glRotatef( static_cast< float >( spin ) * 0.5f, 1.0f, 0.5f, 0.25f );

    // render six axis-aligned colored quads around the origin
    [...]
}
\end{lstlisting}}

The \textsf{eqHello} main function sets up the communication with the
server, initializes and drives the rendering. The details of this setup
are explained in \sref{sEqPly}.

\section{The Programming Interface}

Equalizer uses a C++ programming interface. The API is minimally
invasive, so Equalizer imposes only a minimal, natural execution
framework upon the application. It does not provide a scene graph, or
interferes in any other way with the application's rendering code. The
restructuring work enforced by Equalizer is the minimal refactoring
needed to parallelize the application for rendering.

Methods called by the application have the form \textsf{verb[Noun]},
whereas methods called by Equalizer (`Task Methods') have the form
\textsf{nounVerb}. For example, the application calls
\textsf{Config::startFrame} to render a new frame, which causes, among
many other things, \textsf{Node::frameStart} to be called in all active
render clients.

\subsection{\label{sNamespaces}Namespaces}

Application developers are exposed to the following namespaces:

\begin{description}
\item[eq::base] The eq::base namespace provides C++ classes to abstract
  the underlying operating system and implements common helper
  functionality. One example is the \textsf{eq::base::Clock} class
  providing a high-resolution timer. Classes in this namespace are fully
  documented in the API documentation on the Equalizer website, and are
  not subject of this Programming Guide.
\item[eq::net] The Equalizer network layer provides basic functionality
  for network communication, such as \textsf{Connection} and
  \textsf{ConnectionSet}, as well as higher-level functionality such as
  \textsf{Node}, \textsf{Session} and \textsf{Object}. Please refer to
  \sref{sNetwork} for an introduction into the network layer, and to
  \sref{sNetObject} for distributed objects.
\item[eq] The core namespace for the Equalizer client library. The
  classes and their relationship in this namespace closely model the
  configuration file format. The classes in the \textsf{eq} namespace
  are the main subject of this Programming Guide.
\end{description}

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/namespaces}
  {\caption{\label{fNamespaces}Namespace Layering for eqPly and eqServer}}
\end{wrapfigure}
Examples shipped with Equalizer are implemented in their own namespaces,
e.g., \textsf{eqPly} or \textsf{eVolve}. They rely mostly on subclassing
classes from the \textsf{eq} namespace, with the occasional usage of
functionality from the \textsf{eq::net} and \textsf{eq::base} namespace.

The functionality of the Equalizer server is implemented in the
\textsf{eq::server} namespace. This namespace does not yet expose an API
intended to be used by developers.

\fig{fNamespaces} shows the namespace layering for \textsf{eqPly} and
\textsf{eqServer}.

\subsection{\label{sTaskMethods}Task Methods}

The application inherits from Equalizer classes and overrides virtual
functions to implement certain functionality, e.g., the application's
OpenGL rendering in \textsf{eq::Chan\-nel::frameDraw}. These task
methods are similar in concept to C function callbacks. \sref{sEqPly}
will discuss the important task methods. A full list can be found on the
website\footnote{see
  \link{http://www.equalizergraphics.com/documents/design/taskMethods.html}}.

\subsection{\label{sExecution}Execution Model and Thread Safety}

Using threading correctly in OpenGL-based applications is easy with
Equalizer. Equalizer creates one rendering thread for each graphics
card. All task methods for a pipe, and therefore all OpenGL commands,
are executed from this thread. This threading model is the OpenGL
`threading model', which maintains a current context for each
thread. If structured correctly, the application rarely has to take care
of thread synchronization or protection of shared data.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/model.pdf}
  {\caption{\label{fModel}Simplified Execution Model}}
\end{wrapfigure}
The main thread is responsible for maintaining the application logic. It
reacts on user events, updates the data model and requests new frames to
be rendered. It drives the whole application, as shown in \fig{fModel}.

The rendering threads concurrently render the application's
database. The data\-base should be accessed in a read-only fashion
during rendering to avoid threading problems. This is normally the case,
for example all modern scene graphs use read-only render traversals.

All rendering threads in the configuration run asynchronously to the
application's main thread. Depending on the configuration's latency,
they can fall $n$ frames behind the last frame finished by the
application thread. A latency of one frame is usually not perceived by
the user, but can increase rendering performance substantially since
operations are pipelined better.

Rendering threads on a single node are synchronized when using the
default thread model \textsf{draw\_sync}. When a frame is finished, all
local rendering threads are done drawing. Therefore the application can
safely modify the data between the end of a frame and the beginning of a
new frame. Furthermore, only one instance of the scene data has to be
maintained within a process, since all rendering threads are guaranteed
to draw the same frame.

This per-node frame synchronization does not inhibit latency across
rendering nodes. Furthermore, advanced rendering software which
multi-buffers the dynamic parts of the database can disable the per-node
frame synchronization, as explained in \sref{sThreadModel}. Some scene
graphs implement multi-buffered data, and can profit from relaxing the
frame synchronization.

\section{\label{sEqPly}The eqPly Polygonal Renderer}

\begin{figure}[ht!]\center
  \includegraphics[width=.9\textwidth]{images/uml}
  {\caption{\label{fUml}UML Diagram eqPly and relevant Equalizer Classes}}
\end{figure}

In this section the source code of \textsf{eqPly} is explained in
detail, and relevant design decisions, caveats and other remarks are
discussed.

\textsf{eqPly} is a parallel renderer for polygonal data in the
\textsf{ply} file format. It supports all relevant Equalizer features,
and can be used to render on large-scale displays, immersive
environments with head tracking and to render massive data sets using
all scalable rendering features of Equalizer.

The \textsf{eqPly} example is shipped with the Equalizer distribution
and serves as a reference implementation of an Equalizer-based
application of medium complexity. It focuses on the example usage of
core Equalizer features, not on advanced rendering features or visual
quality.

All classes in the example are in the \textsf{eqPly} namespace to avoid
type name ambiguities, in particular for the \textsf{Window} class which
is frequently used as a type in the global namespace by windowing
systems. \fig{fUml} shows how the most important Equalizer classes are
used through inheritance by the \textsf{eqPly} example.

The \textsf{eqPly} classes fall into two categories: Subclasses of the
rendering entities introduced in \sref{sConfig}, and classes for
distributing data. The function and typical usage for each of the
rendering entities is discussed in this section.

The distributed data classes are helper classes based on
\textsf{eq::Object} or its parent class \textsf{eq::net::Object}. They
illustrate the typical usage of distributed objects for static as well
as dynamic, frame-specific data. Furthermore they are used for a basic
scene graph distribution of the model data.

\subsection{The main Function}

The main function starts off with parsing the command line into the
\textsf{LocalInitData} data structure. A part of it, the base class
\textsf{InitData}, will be distributed to all render client nodes. The
command line parsing is done by the \textsf{LocalInitData} class, which
is discussed in \sref{sInitData}:

{\footnotesize\begin{lstlisting}
int main( int argc, char** argv )
{
    // 1. parse arguments
    eqPly::LocalInitData initData;
    initData.parseArguments( argc, argv );
\end{lstlisting}}

The second step is to redirect all Equalizer output to a file if this
was requested through the command line. This should happen before
initializing Equalizer, so that output during initialization is
redirected and threads launched by Equalizer will use the given log
stream. Any \textsf{std::ostream} can be used as log output:

{\footnotesize\begin{lstlisting}
    // 2. Redirect log output to file, if requested
    std::ofstream* logFile = 0;
    const std::string& logFilename = initData.getLogFilename();
    if( !logFilename.empty( ))
    {
        logFile = new ofstream( logFilename.c_str( ));
        eq::base::Log::setOutput( *logFile );
    }
\end{lstlisting}}

The third step is to initialize the Equalizer library. The
initialization function of Equalizer also parses the command line, which
is used to set certain default values based on Equalizer-specific
options\footnote{Equalizer-specific options always start with -\,-eq-},
e.g., the default server address. Furthermore, a \textsf{NodeFactory} is
provided. The \textsf{EQERROR} macro, and its counterparts
\textsf{EQWARN}, \textsf{EQINFO} and \textsf{EQVERB} allow selective
debugging outputs with various logging levels:

{\footnotesize\begin{lstlisting}
    // 2. Equalizer initialization
    NodeFactory nodeFactory;
    if( !eq::init( argc, argv, &nodeFactory ))
    {
        EQERROR << "Equalizer init failed" << endl;
        return EXIT_FAILURE;
    }
\end{lstlisting}}%>>

The node factory is used by Equalizer to create the object instances of
the configured rendering entities. Each of the classes inherits from the
same type provided by Equalizer in the \textsf{eq} namespace. The
provided \textsf{eq::NodeFactory} base class instantiates 'plain'
Equalizer objects, thus making it possible to selectively subclass
individual entity types, as it is done by \textsf{eqHello}. For each
rendering resources used in the configuration, one C++ object will be
created during initialization. Config, node and pipe objects are created and
destroyed in the node thread, whereas window and channel objects are
created and destroyed in the pipe thread:

{\footnotesize\begin{lstlisting}
class NodeFactory : public eq::NodeFactory
{
public:
    virtual eq::Config*  createConfig( eq::ServerPtr parent )
        { return new eqPly::Config( parent ); }
    virtual eq::Node*    createNode( eq::Config* parent )  
        { return new eqPly::Node( parent ); }
    virtual eq::Pipe*    createPipe( eq::Node* parent )
        { return new eqPly::Pipe( parent ); }
    virtual eq::Window*  createWindow( eq::Pipe* parent )
        { return new eqPly::Window( parent ); }
    virtual eq::Channel* createChannel( eq::Window* parent )
        { return new eqPly::Channel( parent ); }
    virtual eq::View* createView() { return new eqPly::View(); }
};
\end{lstlisting}}

The fourth step is to create an instance of the application and to
initialize it locally. The application is a subclass of
\textsf{eq::Client}, which in turn is an \textsf{eq::net::Node}. The
underlying Equalizer network layer, discussed in \sref{sNetwork}, is a
peer-to-peer network of \textsf{eq::net::Node}s. The client/server
concept is implement the higher-level \textsf{eq} client namespace.

The local initialization of a node creates at least one local listening
socket, which allows the \textsf{eq::Client} to communicate over the
network with other nodes, such as the server and the rendering
clients. The listening socket(s) can be configured using the
\textsf{--eq-listen} command line parameter, or by programmatically
adding connection descriptions to the client before the local
initialization.

{\footnotesize\begin{lstlisting}
    // 4. initialization of local client node
    RefPtr< eqPly::Application > client = new eqPly::Application( initData );
    if( !client->initLocal( argc, argv ))
    {
        EQERROR << "Can't init client" << endl;
        eq::exit();
        return EXIT_FAILURE;
    }
\end{lstlisting}}%>>

Finally everything is set up, and the \textsf{eqPly} application is executed:

{\footnotesize\begin{lstlisting}
    // 5. run client
    const int ret = client->run();
\end{lstlisting}}

After the application has finished, it is de-initialized and the
\textsf{main} function returns:

{\footnotesize\begin{lstlisting}
    // 6. cleanup and exit
    client->exitLocal();
    client = 0;

    eq::exit();

    if( logFile )
        logFile->close();
    delete logFile;

    return ret;
}
\end{lstlisting}}

\subsection{Application}

In the case of \textsf{eqPly}, the application is also the render
client. The \textsf{eqPly} executable has three runtime behaviors:

\begin{enumerate}
\item \textbf{Application}: The executable started by the user, the
  controlling entity of the rendering session.
\item \textbf{Auto-launched render client}: The typical render client,
  started by the server. The server starts the executable with special
  parameters, which cause \textsf{Client::initLocal} to never
  return. During exit, the server terminates the process. By default,
  the server starts the render client using \textsf{ssh}. The launch
  command can be used to configure another program to auto-launch render
  clients.
\item \textbf{Resident render client}: Manually pre-started render
  client, listening on a specified port for server commands. This mode
  is selected using the command-line option \textsf{--eq-client} and
  \textsf{--eq-listen \textless address\textgreater} to specify a
  well-defined listening address, and potentially \textsf{-r} to keep
  the client running across multiple runs\footnote{see
    \link{http://www.equalizergraphics.com/documents/design/residentNodes.html}}.
\end{enumerate}

\subsubsection{Main Loop}

The application's main loop starts by connecting the application to an
Equalizer server. The command line parameter \textsf{--eq-server}
explicitly specifies a server address. If no server was specified,
\textsf{Client::connectServer} tries first to connect to a server on the
local machine using the default port. If that fails, it will create a
server running within the application process with a default one-channel
configuration\footnote{see
  \link{http://www.equalizergraphics.com/documents/design/standalone.html}}.

Another configuration can be selected using the command-line parameter
\textsf{--eq-config}. This configuration may only contain an
\textsf{appNode}, i.e., no distributed configuration is possibly since
the process-local server does not create a listening socket.

{\footnotesize\begin{lstlisting}
int Application::run()
{
    // 1. connect to server
    eq::ServerPtr server = new eq::Server;
    if( !connectServer( server ))
    {
        EQERROR << "Can't open server" << endl;
        return EXIT_FAILURE;
    }
\end{lstlisting}}%>>

The second step is to ask the server for a configuration. The
\textsf{ConfigParams} are a placeholder for later Equalizer
implementations to provide additional hints and information to the
server for choosing the configuration. The configuration chosen by the
server is created locally using
\textsf{NodeFactory::createConfig}. Therefore it is of type
\textsf{eqPly::Config}, but the return value is \textsf{eq::Config},
making the cast necessary:

{\footnotesize\begin{lstlisting}
    // 2. choose config
    eq::ConfigParams configParams;
    Config* config = static_cast<Config*>(server->chooseConfig( configParams ));

    if( !config )
    {
        EQERROR << "No matching config on server" << endl;
        disconnectServer( server );
        return EXIT_FAILURE;
    }
\end{lstlisting}}%>>

Finally it is time to initialize the configuration. For statistics, the
time for this operation is measured and printed. During initialization
the server launches and connects all render client nodes, and calls the
appropriate initialization task methods, as explained in later
sections. \textsf{Config::init} returns after all nodes, pipes,
windows and channels are initialized. It returns \textsf{true} only if
all initialization task methods were successful.

The \textsf{EQLOG} macro allows topic-specific logging. The numeric
topic values are specified in the respective \textsf{log.h} header
files, and logging for various topics is enabled using the environment
variable \textsf{EQ\_LOG\_TOPICS}:

{\footnotesize\begin{lstlisting}
    // 3. init config
    eq::base::Clock clock;

    config->setInitData( _initData );
    if( !config->init( ))
    {
        EQERROR << "Config initialization failed: " 
                << config->getErrorMessage() << endl;
        server->releaseConfig( config );
        disconnectServer( server );
        return EXIT_FAILURE;
    }

    EQLOG( eq::LOG_STATS ) << "Config init took " << clock.getTimef() << " ms"
                           << endl;
\end{lstlisting}}%>>

When the configuration was successfully initialized, the main rendering
loop is executed. It runs until the user exits the
configuration, or when a maximum number of frames has been rendered,
specified by a command-line argument. The latter is useful for
benchmarks. The \textsf{Clock} is reused for measuring the overall
performance. A new frame is started using \textsf{Config::startFrame}
and a frame is finished using \textsf{Config::finishFrame}.

When a new frame is started, the server computes all rendering tasks and
sends them to the appropriate render client nodes. The render client
nodes dispatch the tasks to the correct node or pipe thread, where they
are executed in order of arrival.

\textsf{Config::finishFrame} blocks on the completion of the frame
\textsf{current - latency}. The latency is specified in the
configuration file, and allows several outstanding frames. This allows
overlapping execution in the node processes and pipe threads and
minimizes idle times.

By default, \textsf{Config::finish\-Fra\-me} also synchronizes the
completion of all local rendering tasks for the current frame. This
facilitates porting of existing rendering codes, since the database does
not have to be multi-buffered. Applications such as \textsf{eqPly}, which
do not need this per-node frame synchronization, can disable it as
explained in \sref{sThreadModel}.

\begin{figure}[ht!]\center
  \includegraphics[width=\textwidth]{images/syncAsync}
  {\caption{\label{fSyncAsync}Synchronous and Asynchronous Execution}}
\end{figure}
\fig{fSyncAsync} shows the execution of the rendering tasks of a 2-node
2D compound without latency and with a latency of one frame. The
asynchronous execution pipelines certain rendering operations and hides
imbalances in the load distribution, resulting in an improved
framerate. For example, we have observed a speedup of 15\% on a
five-node rendering cluster when using a latency of one frame instead of
no
latency\footnote{\link{http://www.equalizergraphics.com/scalability.html}}. A
latency of one or two frames is normally not perceived by the user.

\textsf{eqPly} uses event-driven execution, that is, it only request new
rendering frames if an event or animation requires an update. The
\textsf{eqPly::Config} maintains a dirty state, which is cleared after a
frame has been started, and set when an event causes a
redraw. Furthermore, when an animation is running or head tracking is
active, the config always signals the need for a new frame.

If the application detects that it is currently idle, all pending
commands are gradually flushed, while still looking for a redraw
event. Then it waits and handles one event at a time, until a redraw is
needed:

{\footnotesize\begin{lstlisting}
    // 4. run main loop
    uint32_t maxFrames = _initData.getMaxFrames();

    clock.reset();
    while( config->isRunning( ) && maxFrames-- )
    {
        config->startFrame();
        // config->renderData(...);
        config->finishFrame();

        while( !config->needsRedraw( )) // wait for an event requiring redraw
        {
            if( hasCommands( )) // execute non-critical pending commands
            {
                processCommand();
                config->handleEvents(); // non-blocking
            }
            else  // no pending commands, block on user event
            {
                const eq::ConfigEvent* event = config->nextEvent();
                if( !config->handleEvent( event ))
                    EQVERB << "Unhandled " << event << endl;
            }
        }
        config->handleEvents(); // process all pending events
    }
\end{lstlisting}}%>>

When the main rendering loop has finished,
\textsf{Config::finishAllFrames} is called to catch up with the
latency. It returns after all outstanding frames have been rendered, and
is needed to provide an accurate measurement of the framerate:

{\footnotesize\begin{lstlisting}
    const uint32_t frame = config->finishAllFrames();
    const float    time  = clock.getTimef();
    EQLOG( eq::LOG_STATS ) << "Rendering took " << time << " ms (" << frame
                           << " frames @ " << ( frame / time * 1000.f)
                           << " FPS)" << endl;
\end{lstlisting}}%>>

The remainder of the application code cleans up in the reverse order of
initialization. The config is exited, released and the connection to
the server is closed:

{\footnotesize\begin{lstlisting}
    // 5. exit config
    clock.reset();
    config->exit();
    EQLOG( eq::LOG_STATS ) << "Exit took " << clock.getTimef() << " ms" <<endl;

    // 6. cleanup and exit
    server->releaseConfig( config );
    if( !disconnectServer( server ))
        EQERROR << "Client::disconnectServer failed" << endl;
    server = 0;
    return EXIT_SUCCESS;
}
\end{lstlisting}}%>>

\subsubsection{Render Clients}

In the second and third use case of the \textsf{eqPly}, when the
executable is used as a render client, \textsf{Client::initLocal} never
returns. Therefore the application's main loop is never executed. To
keep the client resident, the \textsf{eqPly} example overrides the
client loop to keep it running beyond one configuration run:

{\footnotesize\begin{lstlisting}
bool Application::clientLoop()
{
    if( !_initData.isResident( )) // execute only one config run
        return eq::Client::clientLoop();

    // else execute client loops 'forever'
    while( true ) // TODO: implement SIGHUP handler to exit?
    {
        if( !eq::Client::clientLoop( ))
            return false;
        EQINFO << "One configuration run successfully executed" << endl;
    }
    return true;
}
\end{lstlisting}}%>>

\subsection{\label{sNetObject}Distributed Objects}

Equalizer provides distributed objects which facilitate the
implementation of data distribution in a cluster
environment. Distributed objects are created by subclassing from
\textsf{eq::Object} or \textsf{eq::net::Object}. The application
programmer implements serialization and deserialization of the
distributed data. \sref{sNetObject2} covers distributed objects in detail.

Distributed objects can be static (immutable) or dynamic. Dynamic objects are
versioned. The \textsf{eqPly} example uses static distributed objects to provide
initial data and the model to all rendering nodes, as well as a versioned object
to provide frame-specific data such as the camera position to the rendering
methods.

\subsubsection{\label{sInitData}InitData - a Static Distributed Object}

The \textsf{InitData} class holds a couple of parameters needed during
initialization. These parameters never change during one configuration
run, and are therefore static.

On the application side, the class \textsf{LocalInitData} subclasses
\textsf{InitData} to provide the command line parsing and to set the
default values. The render nodes only instantiate the distributed part
in \textsf{InitData}.

A static distributed object has to implement \textsf{getInstanceData}
and \textsf{applyInstanceData} to serialize and deserialize the object's
distributed data. These method provide an output or input stream as a
parameter, which abstracts the data transmission and can be used like a
\textsf{std::stream}.

The data streams implement efficient buffering and data transport
between nodes. They perform no type checking or transformation on the
data. It is the application's responsibility to exactly match the order
and types of variables during serialization and de-serialization.

Custom data type serializers can be implemented by providing the
appropriate serialization functions. No pointers should be directly
transmitted through the data streams. For pointers, the corresponding
object is typically a distributed object as well, and it's identifier
and potentially version is transmitted in place of its pointer.

For \textsf{InitData}, serialization in \textsf{getInstanceData}
and de-serialization in \textsf{applyInstanceData} is performed by
streaming all member variables to or from the provided data
streams:

{\footnotesize\begin{lstlisting}
void InitData::getInstanceData( eq::net::DataOStream& os )
{
    os << _frameDataID << _windowSystem << _renderMode << _useGLSL << _invFaces
       << _logo;
}

void InitData::applyInstanceData( eq::net::DataIStream& is )
{
    is >> _frameDataID >> _windowSystem >> _renderMode >> _useGLSL >> _invFaces
       >> _logo;

    EQASSERT( _frameDataID != EQ_ID_INVALID );
    EQINFO << "New InitData instance" << endl;
}
\end{lstlisting}}%>>

\subsubsection{FrameData - a Versioned Distributed Object}

Versioned objects have to override \textsf{getChangeType} to indicate
how they want to have changes to be handled. All types of versioned
objects currently implemented have the following characteristics:

\begin{itemize}
\item Only the master instance of the object generates new versions, that is,
  \textsf{eq::net::Object::com\-mit} should only be called only on the master
  instance. It is possible to commit on slave instances, but special care has to
  be taken to handle possible conflicts. \sref{sSlaveCommit} covers slave object
  commits in detail.
\item Slave instance versions can only be advanced, that is,
  \textsf{eq::net::Object::sync( version )} with a version smaller than
  the current version will fail.
\item Newly mapped slave instances are mapped to the oldest available
  version by default, or to the version specified when calling
  \textsf{Session::mapObject}.
\end{itemize}

Upon \textsf{commit} the delta data from the previous version is sent to
all mapped slave instances. The data is queued on the remote node, and
is applied when the application calls \textsf{sync} to synchronize the
object to a new version. The \textsf{sync} method might block if a
version has not yet been committed or is still in transmission.

Not syncing a mapped, versioned object creates a memory leak. The method
\textsf{Object::notifyNewHeadVersion} is called whenever a new version
is received by the node. The notification is send from the command
thread, which is different from the node main thread. The object should
not be synced from this method, but instead a message may be send to the
application, which then takes the appropriate action. The default
implementation asserts when too many versions have been queued
to detect memory leaks during development.

Besides the instance data (de-)serialization methods used to map an
object, versioned objects may implement \textsf{pack} and
\textsf{unpack} to serialize or de-serialize the changes since the last
version. If these methods are not implemented, their default
implementation forwards the (de-)serialization request to
\textsf{getInstanceData} and \textsf{applyInstanceData}, respectively.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/umlObject.pdf}
  {\caption{\label{fUMLObject}eq::Object and eq::net::Object}}
\end{wrapfigure}
The creation of distributed, versioned objects is further facilitated by
\textsf{eq::Object}, which implements the most common way of tracking
data changes in versioned objects. The concept of a dirty bit mask is
used to mark parts of the object for serialization, while preserving the
capability to inherit objects. Other ways of implementing change
tracking, e.g., using incarnation counters, can still be implemented by
using \textsf{eq::net::Object} which leaves this flexibility to the
developer. \fig{fUMLObject} shows the relationship between
\textsf{eq::Object} and \textsf{eq::net::Object}.

The \textsf{FrameData} is sub-classed from \textsf{eq::Object}, and
consequently tracks its changes by setting the appropriate dirty bit
whenever it is changed. The serialization methods are called by the
\textsf{eq::Object} with the dirty bit mask needed to serialize all
data, or with the dirty bit mask of the changes since the last
\textsf{commit}. The \textsf{FrameData} only defines its own dirty bits
and serialization code:

{\footnotesize\begin{lstlisting}
/** The changed parts of the data since the last pack(). */
enum DirtyBits
{
    DIRTY_CAMERA  = eq::Object::DIRTY_CUSTOM << 0,
    DIRTY_FLAGS   = eq::Object::DIRTY_CUSTOM << 1,
    DIRTY_VIEW    = eq::Object::DIRTY_CUSTOM << 2,
    DIRTY_MESSAGE = eq::Object::DIRTY_CUSTOM << 3,
};

void FrameData::serialize( eq::net::DataOStream& os, const uint64_t dirtyBits )
{
    eq::Object::serialize( os, dirtyBits );
    if( dirtyBits & DIRTY_CAMERA )
        os << _translation << _rotation << _modelRotation;
    if( dirtyBits & DIRTY_FLAGS )
        os << _modelID << _renderMode << _colorMode << _ortho << _statistics
           << _help << _wireframe << _pilotMode << _idleMode;
    if( dirtyBits & DIRTY_VIEW )
        os << _currentViewID;
    if( dirtyBits & DIRTY_MESSAGE )
        os << _message;
}

void FrameData::deserialize( eq::net::DataIStream& is,
                             const uint64_t dirtyBits )
{
    eq::Object::deserialize( is, dirtyBits );
    if( dirtyBits & DIRTY_CAMERA )
        is >> _translation >> _rotation >> _modelRotation;
    if( dirtyBits & DIRTY_FLAGS )
        is >> _modelID >> _renderMode >> _colorMode >> _ortho >> _statistics
           >> _help >> _wireframe >> _pilotMode >> _idleMode;
    if( dirtyBits & DIRTY_VIEW )
        is >> _currentViewID;
    if( dirtyBits & DIRTY_MESSAGE )
        is >> _message;
}
\end{lstlisting}}%>>

\subsubsection{\label{sSceneData}Scene Data}

Some applications might rely on a shared filesystem to access the data,
for example when out-of-core algorithms are used. Other applications
prefer to load the data only on the application process, and use
distributed objects to synchronize the scene data with the render
clients.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/modelDist.pdf}
  {\caption{\label{fModelDist}Scene Data in eqPly}}
\end{wrapfigure}
\textsf{eqPly} uses static distributed objects to distribute the model
loaded by the application. This approach can be easily extended to
versioned objects to support dynamic data modifications.

The kD-tree data structure and rendering code for the model is strongly
separated from Equalizer, and kept in the separate namespace
\textsf{mesh}. It can also be used in other rendering software, for
example in a GLUT application. To keep this separation, an external
'mirror' hierarchy is constructed around the tree. This hierarchy of
\textsf{VertexBufferDist} nodes is responsible for cloning the model
data on the remote render clients.

The identifier of the model's root object of this distributed hierarchy
is passed as part of the \textsf{InitData} for the default model, or as
part of the \textsf{View} for each logical view. It is used on the
render clients to map the model when it is needed for
rendering. \fig{fModelDist} shows the UML hierarchy of the model and
distribution classes.

Each \textsf{VertexBufferDist} object corresponds to one node of the
model's data tree. It is serializing the data for this
node. Furthermore, it mirrors the kD-tree by having a
\textsf{VertexBufferDist} child for each child of its corresponding tree
node. During serialization, the identifier of these children is sent to
the remote nodes, which reconstruct the mirror distribution hierarchy
and model data tree based on this data.

The serialization function \textsf{getInstanceData} sends all the data
needed to reconstruct the model tree: the object identifiers of its
children, vertex data for the tree root and vertex indices for the leaf
nodes, as well as the bounding sphere and database range of each
node. The deserialization function \textsf{applyInstanceData} retrieves
the data in multiple steps, and constructs the model tree on the fly
based on this information. It is omitted here for brevity:

{\footnotesize\begin{lstlisting}
void VertexBufferDist::getInstanceData( eq::net::DataOStream& os )
{
    EQASSERT( _node );
    os << _isRoot;

    if( _left && _right )
    {
        os << _left->getID() << _right->getID();

        if( _isRoot )
        {
            EQASSERT( _root );
            const mesh::VertexBufferData& data = _root->_data;

            os << data.vertices << data.colors << data.normals << data.indices;
        }
    }
    else
    {
        os << EQ_ID_INVALID << EQ_ID_INVALID;

        EQASSERT( dynamic_cast< const mesh::VertexBufferLeaf* >( _node ));
        const mesh::VertexBufferLeaf* leaf = 
            static_cast< const mesh::VertexBufferLeaf* >( _node );

        os << leaf->_vertexStart << leaf->_vertexLength << leaf->_indexStart
           << leaf->_indexLength;
    }

    os << _node->_boundingSphere << _node->_range;
}
\end{lstlisting}}%>>

%\begin{wrapfigure}{r}{.618\textwidth}
\begin{figure}[ht!]\center
  \includegraphics[width=.618\textwidth]{images/objects.pdf}
  {\caption{\label{fObjects}Scene Graph Distribution}}
\end{figure}
Applications distributing a dynamic scene graph use the frame data
instead of the init data as the entry point to their scene graph data
structure. \fig{fObjects} shows one possible implementation, where the
identifier and version of the scene graph root are transported using the
frame data. The scene graph root then serializes and de-serializes his
immediate children by transferring their identifier and current version,
similar to the static distribution done by \textsf{eqPly}.

The objects are still created by the application, and then registered or
mapped with the session to distribute them. When mapping
objects in a hierarchical data structure, their type often has to be
known to create them. Equalizer does not currently provide
object typing, this has to be done by the application, either implicitly
in the current implementation context, or by transferring a type
identifier. In \textsf{eqPly}, object typing is implicit since it is
well-defined which object is mapped in which context.

\subsection{Config}

The configuration is driving the application's rendering, that is, it is
responsible for updating the data based on received events, requesting
new frames to be rendered and to provide the render clients with the
necessary data.

\subsubsection{Initialization and Exit}

The config initialization happens in parallel, that is, all config
initialization tasks are transmitted by the server at once and their
completion is synchronized afterwards.

\begin{figure}[ht!]\center
%\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-2ex}
  \includegraphics[width=.9\textwidth]{images/configInit.pdf}
  {\caption{\label{fConfigInit}Config Initialization Sequence}}
  \vspace{-4ex}
\end{figure}
The tasks are executed by the node and pipe threads in parallel. The
parent's initialization methods are always executed before any child
initialization method. This parallelization is necessary to allow a
speedy startup of the configuration on large-scale graphics clusters. On
the other hand, it means that initialization functions are called even
if the parent's initialization has failed. \fig{fConfigInit} shows a
sequence diagram of the config initialization.

The \textsf{eqPly::Config} class holds the master versions of the
initialization and frame data. Both objects are registered with the
\textsf{eq::Config}, which is the \textsf{eq::net::Session} used for
rendering. Equalizer takes care of the session setup and exit in
\textsf{Client::choose\-Config} and \textsf{Client::releaseConfig},
respectively.

The frame data is registered before the initialization data, since its
identifier is transmitted using the \textsf{InitData}. The identifier of
the initialization data is transmitted to the render client nodes using
the \textsf{initID} parameter of \textsf{eq::Config::init}.

Equalizer will pass this identifier to all \textsf{configInit} calls of
the respective objects:

{\footnotesize\begin{lstlisting}
bool Config::init()
{
    // init distributed objects
    _frameData.data.color      = _initData.useColor();
    _frameData.data.renderMode = _initData.getRenderMode();
    registerObject( &_frameData );

    _initData.setFrameDataID( _frameData.getID( ));
    registerObject( &_initData );

    // init config
    _running = eq::Config::init( _initData.getID( ));
    if( !_running )
        return false;
\end{lstlisting}}

If the configuration was initialized correctly, the configuration tries
to set up a tracking device for head tracking. Equalizer does not
provide extensive support for tracking devices, as this is an orthogonal
problem to parallel rendering. Tracking device support has already been
solved by a number of implementations\footnote{VRCO Trackd, VRPN, etc.},
which can and have been easily integrated with Equalizer. The example
code in \textsf{eqPly} provides a reference implementation for the
integration of such a tracking library. \sref{sTracking} provides more
background on head tracking.

{\footnotesize\begin{lstlisting}
    // init tracker
    if( !_initData.getTrackerPort().empty( ))
    {
        if( !_tracker.init( _initData.getTrackerPort() ))
            EQWARN << "Failed to initialise tracker" << endl;
        else
        {
            // Set up position of tracking system wrt world space
            // Note: this depends on the physical installation.
            vmml::Matrix4f m( vmml::Matrix4f::IDENTITY );
            m.scale( 1.f, 1.f, -1.f );
            _tracker.setWorldToEmitter( m );

            m = vmml::Matrix4f::IDENTITY;
            m.rotateZ( -M_PI_2 );
            _tracker.setSensorToObject( m );
            EQINFO << "Tracker initialized" << endl;
        }
    }

    return true;
}
\end{lstlisting}}%>>

The exit function of the configuration stops the render clients by calling
\textsf{eq::Con\-fig::exit}, and then de-registers the initialization and
frame data objects with the session:

{\footnotesize\begin{lstlisting}
bool Config::exit()
{
    _running = false;
    const bool ret = eq::Config::exit();

    _initData.setFrameDataID( EQ_ID_INVALID );
    deregisterObject( &_initData );
    deregisterObject( &_frameData );

    return ret;
}
\end{lstlisting}}

\subsubsection{Frame Control}

The rendering frames are issued by the application. The
\textsf{eqPly::Config} only overrides \textsf{startFrame} to
update its data before forwarding the start frame request to the
\textsf{eq::Config}.

If a tracker is used, the current head position and orientation is
retrieved and passed to Equalizer, which uses the head matrix together
with the wall or projection description to compute the view
frusta, as explained in \sref{sTracking}.

The camera position is updated and the frame data is committed, which
generates a new version of this object. This version is passed to the
rendering callbacks and will be used by the rendering threads to
synchronize the frame data to the state belonging to the current frame:

{\footnotesize\begin{lstlisting}
uint32_t Config::startFrame()
{
    // update head position
    if( _tracker.isRunning() )
    {
        _tracker.update();
        const eq::Matrix4f& headMatrix = _tracker.getMatrix();
        _setHeadMatrix( headMatrix );
    }

    // update database
    if( _animation.isValid( ))
    {
        const eq::Vector3f&  modelRotation = _animation.getModelRotation();
        const CameraAnimation::Step& curStep = _animation.getNextStep();

        _frameData.setModelRotation( modelRotation        );
        _frameData.setRotation(     curStep.rotation      );
        _frameData.setTranslation(  curStep.translation   );
    }
    else
    {
        if( _frameData.usePilotMode())
            _frameData.spinCamera( -0.001f * _spinX, -0.001f * _spinY );
        else
            _frameData.spinModel( -0.001f * _spinX, -0.001f * _spinY );

        _frameData.moveCamera( 0.0f, 0.0f, 0.001f*_advance );
    }

    [...]

    const uint32_t version = _frameData.commit();

    _redraw = false;
    return eq::Config::startFrame( version );
}
\end{lstlisting}}

\subsubsection{Event Handling}

Events are sent by the render clients to the application using
\textsf{eq::Config::sendEvent}. At the end of the frame,
\textsf{Config::finishFrame} calls \textsf{Config::handleEvents} to do
the event handling. The default implementation processes all pending
events by calling \textsf{Config::handleEvent} for each of them.

Since \textsf{eqPly} uses event-driven execution, the config maintains a
dirty state to allow the application to know when a redraw is needed.

The \textsf{eqPly} example implements \textsf{Config::hand\-le\-Event}
to provide the various reactions to user input, most importantly camera
updates based on mouse events. The camera position has to be handled
correctly regarding latency, and is therefore saved in the frame data:

{\footnotesize\begin{lstlisting}
bool Config::handleEvent( const eq::ConfigEvent* event )
{
    switch( event->type )
    {
        [...]
        case eq::ConfigEvent::POINTER_MOTION:
            if( event->pointerMotion.buttons == eq::PTR_BUTTON_NONE )
                return true;

            if( event->pointerMotion.buttons == eq::PTR_BUTTON1 )
            {
                _spinX = 0;
                _spinY = 0;

                _frameData.data.rotation.preRotateX( 
                    -0.005f * event->pointerMotion.dx );
                _frameData.data.rotation.preRotateY(
                    -0.005f * event->pointerMotion.dy );
                _redraw = true;
            }
            else if( event->pointerMotion.buttons == eq::PTR_BUTTON2 ||
                     event->pointerMotion.buttons == ( eq::PTR_BUTTON1 |
                                                       eq::PTR_BUTTON3 ))
            {
                _frameData.data.translation.z +=
                    .005f * event->pointerMotion.dy;
                _redraw = true;
            }
            else if( event->pointerMotion.buttons == eq::PTR_BUTTON3 )
            {
                _frameData.data.translation.x += 
                    .0005f * event->pointerMotion.dx;
                _frameData.data.translation.y -= 
                    .0005f * event->pointerMotion.dy;
                _redraw = true;
            }
            return true;

        default:
            break;
    }
    return eq::Config::handleEvent( event );
}
\end{lstlisting}}

\subsubsection{Model Handling}

Model data is handled by the config. Models in eqPly are static,
and therefore the render clients only need to map one instance of the
model per node.

Multiple models can be loaded in \textsf{eqPly}. A configuration has a
default model, stored in \textsf{InitData}, and one model per view,
stored and distributed using \textsf{View}. The
loaded models are evenly distributed over the available views of the
configuration, as shown in \fig{fLayout}.

The channel acquires the model during rendering from the config, using
the model identifier from its current view, or from the frame data if
no view is configured.

The config maintains the mapped models, and lazily maps new models,
which are registered by the application process:

{\footnotesize\begin{lstlisting}
const Model* Config::getModel( const uint32_t modelID )
{
    if( modelID == EQ_ID_INVALID ) // no model loaded by application
        return 0;

    // Accessed concurrently from pipe threads
    eq::base::ScopedMutex< eq::base::SpinLock > _mutex( _modelLock );

    const size_t nModels = _models.size();
    EQASSERT( _modelDist.size() == nModels );

    for( size_t i = 0; i < nModels; ++i )
    {
        const ModelDist* dist = _modelDist[ i ];
        if( dist->getID() == modelID )
            return _models[ i ];
    }

    _modelDist.push_back( new ModelDist );
    Model* model = _modelDist.back()->mapModel( this, modelID );
    EQASSERT( model );
    _models.push_back( model );

    return model;
}
\end{lstlisting}}

\subsubsection{Layout and View Handling}

For layout and model selection, \textsf{eqPly} maintains an active view
and canvas. The identifier of the active view is stored in the frame
data, which is used by the render client to highlight it using a
different background color. The active view can be selected by clicking
into a view, or by cycling through all the views using a keyboard
shortcut.

The model of the active view can be changed using a keyboard
shortcut. The model is view-specific, and therefore the model identifier
for each view is stored on the view, which is used to retrieve the model
on the render clients.

View-specific data is not limited to a model. Applications can choose to
make any application-specific data view-specific, e.g., cameras,
rendering modes or annotations. A view is a generic concept for an
application-specific view on data, \textsf{eqPly} is using different
models to illustrate the concept:

{\footnotesize\begin{lstlisting}
bool Config::_handleKeyEvent( const eq::KeyEvent& event )
{
    switch( event.key )
    {
        ...
        case 'm':
        case 'M':
        {
            if( _modelDist.empty( )) // no models
                return true;

            // current model
            const uint32_t viewID = _frameData.getCurrentViewID();
            View* view = static_cast< View* >( findView( viewID ));
            const uint32_t currentID = view ? 
                view->getModelID() : _frameData.getModelID();

            // next model
            ModelDistVector::const_iterator i;
            for( i = _modelDist.begin(); i != _modelDist.end(); ++i )
            {
                if( (*i)->getID() != currentID )
                    continue;

                ++i;
                break;
            }
            if( i == _modelDist.end( ))
                i = _modelDist.begin(); // wrap around

            // set identifier on view or frame data (default model)
            const uint32_t modelID = (*i)->getID();
            if( view )
                view->setModelID( modelID );
            else
                _frameData.setModelID( modelID );
            return true;
        }
        ...
    }
}
\end{lstlisting}}

The layout of the canvas with the active view can also be dynamically
switched using a keyboard shortcut. The first canvas using the layout is
found, and then the next layout of the configuration is set on this
canvas.

Switching a layout causes the initialization and de-initialization task
methods to be called on the involved channels, and potentially windows,
pipes and nodes. This operation might fail, which will cause the config
to stop running.

Layout switching is typically used to change the presentation of views
at runtime. The source code omitted for brevity.

\subsection{Node}

For each active render client, one \textsf{eq::Node} instance is
created on the appropriate machine. Nodes are only instantiated on their
render client processes, i.e., each process will only have one
instance of the \textsf{eq::Node} class. The application process might
also have a node class, which is handled in exactly the same way as the
render client nodes. The application and render clients might use a
different node factory, instantiating a different type of \textsf{eq::Node}.

All dynamic data is multi-buffered in \textsf{eqPly}. During
initialization, the node relaxes the thread synchronization between the
node and pipe threads, unless the configuration file overrides
this. \sref{sThreads} provides a detailed explanation of thread
synchronization in Equalizer.

During node initialization the static, per-config data is mapped to a
local instance using the identifier passed from
\textsf{Config::init}. No pipe, window or channel tasks methods are
executed before \textsf{Node::configInit} has returned:

{\footnotesize\begin{lstlisting}
bool Node::configInit( const uint32_t initID )
{
    if( !eq::Node::configInit( initID ))
        return false;

    // All render data is static or multi-buffered, we can run asynchronously
    if( getIAttribute( IATTR_THREAD_MODEL ) == eq::UNDEFINED )
        setIAttribute( IATTR_THREAD_MODEL, eq::ASYNC );

    Config*    config = static_cast< Config* >( getConfig( ));
    return config->mapData( initID );
}
\end{lstlisting}}

The actual mapping of the static data is done by the config. The config
retrieves the distributed \textsf{InitData}. The object is directly
unmapped since it is static, and the data has been retrieved:

{\footnotesize\begin{lstlisting}
bool Config::mapData( const uint32_t initDataID )
{
    if( _initData.getID() == EQ_ID_INVALID )
    {
        EQCHECK( mapObject( &_initData, initDataID ));
        unmapObject( &_initData ); // data was retrieved, unmap immediately
    }
    else  // appNode, _initData is registered already
        EQASSERT( _initData.getID() == initDataID );
}
\end{lstlisting}}

\subsection{Pipe}

All task methods for a pipe and its children are executed in a separate
thread. This approach optimizes usage of the GPU, since all
tasks are executed serially and therefore do not compete for resources
or cause OpenGL context switches. Later versions of Equalizer might
introduce threaded windows to allow the parallel and independent
execution of rendering tasks on a single pipe.

\subsubsection{Initialization and Exit}

Pipe threads are not explicitly synchronized with each other, that is,
pipes might be rendering different frames at any given time. Therefore
frame-specific data has to be allocated for each pipe thread, which in
the case of \textsf{eqPly} is the frame data. The frame data is a member
variable of the \textsf{eqPly::Pipe}, and is mapped to the identifier
provided by the initialization data. The initialization in
\textsf{eq::Pipe} does the GPU-specific initialization, which is
window-system-dependent. On AGL the display ID is determined, and on glX
the display connection is opened.

{\footnotesize\begin{lstlisting}
bool Pipe::configInit( const uint32_t initID )
{
    const Config*     config    = static_cast<Config*>( getConfig( ));
    const InitData& initData    = config->getInitData();
    const uint32_t  frameDataID = initData.getFrameDataID();
    eq::Config*     config      = getConfig();

    const bool mapped = config->mapObject( &_frameData, frameDataID );
    EQASSERT( mapped );

    return eq::Pipe::configInit( initID );
}
\end{lstlisting}}

The config exit function is similar to the config initialization. The
frame data is unmapped and GPU-specific data is de-initialized by
\textsf{eq::Config::exit}:

{\footnotesize\begin{lstlisting}
bool Pipe::configExit()
{
    eq::Config* config = getConfig();
    config->unmapObject( &_frameData );

    return eq::Pipe::configExit();
}
\end{lstlisting}}

\subsubsection{Window System}

Equalizer supports multiple window system interfaces, at the moment
glX/X11, WGL and AGL/Carbon. Some operating systems, and therefore some
Equalizer versions, support multiple window systems
concurrently\footnote{see
  \link{http://www.equalizergraphics.com/compatibility.html}}.

Each pipe might use a different window system for rendering, which is
determined before \textsf{Pipe::configInit} by
\textsf{Pipe::selectWindowSystem}. The default implementation of
\textsf{selectWindowSystem} loops over all window systems and returns
the first supported window system, determined by using
\textsf{supportsWindowSystem}.

The \textsf{eqPly} examples allows selecting the window system using a
command line option. Therefore the implementation of
\textsf{selectWindowSystem} is overwritten and returns the specified
window system, if supported:

{\footnotesize\begin{lstlisting}
eq::WindowSystem Pipe::selectWindowSystem() const
{
    const Config*            config = static_cast<Config*>( getConfig( ));
    const InitData&        initData = config->getInitData();
    const eq::WindowSystem ws       = initData.getWindowSystem();

    if( ws == eq::WINDOW_SYSTEM_NONE )
        return eq::Pipe::selectWindowSystem();
    if( !supportsWindowSystem( ws ))
    {
        EQWARN << "Window system " << ws 
               << " not supported, using default window system" << endl;
        return eq::Pipe::selectWindowSystem();
    }

    return ws;
}
\end{lstlisting}}%>>

\subsubsection{\label{sCarbonThread}Carbon/AGL Thread Safety}

Parts of the Carbon API used for window and event handling in the AGL
window system are not thread safe. The application has to call
\textsf{eq::Global::enterCarbon} before any thread-unsafe Carbon call,
and \textsf{eq::Global::leaveCarbon} afterwards. These functions should
be used only during window initialization and exit, not during
rendering. For various reasons \textsf{enterCarbon} might block up to 50
milliseconds. Carbon calls in the window event handling routine
\textsf{Window::processEvent} are thread-safe, since the global carbon
lock is set in this method. Please contact the Equalizer developer
mailing list if you need to use Carbon calls on a per-frame basis.

\subsubsection{Frame Control}

All task methods for a given frame of the pipe, window and channel
entities belonging to the thread are executed in one block, starting
with \textsf{Pipe::frameStart} and finished by
\textsf{Pipe::finishFrame}. The frame start callback is therefore the
natural place to update all frame-specific data to the version belonging
to the frame.

In \textsf{eqPly}, the version of the only frame-specific object
\textsf{FrameData} is passed as the per-frame id from
\textsf{Config::startFrame} to the frame task methods. The pipe uses
this version to update its instance of the frame data to the current
version, and then unlocks its child entities by calling
\textsf{startFrame}:

{\footnotesize\begin{lstlisting}
void Pipe::frameStart( const uint32_t frameID, const uint32_t frameNumber )
{
    node->waitFrameStarted( frameNumber );

    _frameData.sync( frameID );
    startFrame( frameNumber );
}
\end{lstlisting}}

\subsection{\label{sEqplyWIndow}Window}

The Equalizer window abstracts an OpenGL drawable and a rendering
context. When using the default window initialization functions, all
windows of a pipe share the OpenGL context. This allows reuse of OpenGL
objects such as display lists and textures between all windows of one
pipe.

The window class is the natural place for the application to maintain
all data specific to the OpenGL context.

\subsubsection{Window System Interface}

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-2ex}\includegraphics[width=.618\textwidth]{images/osWindow.pdf}
  {\caption{\label{fOSWindow}eq::OSWindow UML Class Hierarchy}}
\end{wrapfigure}
The particulars of creating a window and OpenGL context depend on the
window system used. One can either use the implementation provided by
the operating system, e.g., AGL, WGL or glX, or some higher-level
toolkit, e.g., Qt.

All window-system specific functionality is implemented by a
specialization of \textsf{eq::OS\-Window}. The \textsf{OSWindow} class
defines the minimal interface to be implemented for a new window
system. Each \textsf{Window} uses one \textsf{OSWindow} during
execution. This separation allows an easy implementation and adaption to
another window system or application.

Equalizer provides a generic interface and implementation for the three
most common window systems through the OpenGL specific 
\textsf{eq::GL\-Window} class: AGL, WGL and glX. The interfaces define the
minimal functionality needed to reuse other window system specific
classes, for example the AGL, WGL and glX event handlers. The separation 
of the OpenGL calls from the \textsf{OSWindow} permits to implement 
window systems which don't use any OpenGL context and therefore to use a 
different renderer. The implementation derived from these interfaces 
provides a sample implementation which honors all configurable window 
attributes.

\subsubsection{Initialization and Exit}

The initialization sequence uses multiple, override-able task
methods. The main task method \textsf{configInit} calls first
\textsf{configInitOSWindow}, which creates and initializes the
\textsf{OSWindow} for this window. The \textsf{OSWindow} initialization
code is implementation specific. If the \textsf{OSWindow} was
initialized successfully, \textsf{configInit} calls
\textsf{configInitGL}, which performs the generic OpenGL state
initialization. The default implementation sets up some typical OpenGL
state, e.g., it enables the depth test. Most nontrivial applications
do override this task method.

The \textsf{OSWindow} initialization takes into account various
attributes set in the configuration file. Attributes include the size of
the various frame buffer planes (color, alpha, depth, stencil) as well
as other framebuffer attributes, such as quad-buffered stereo,
doublebuffering, fullscreen mode and window decorations. Some of the
attributes, such as stereo, doublebuffer and stencil can be set to
\textsf{eq::AUTO}, in which case the Equalizer default implementation
will test for their availability and enable them if possible.

For the window-system specific initialization, \textsf{eqPly} uses the
default Equalizer implementation. The \textsf{eqPly} window
initialization only overrides the OpenGL-specific initialization
function \textsf{configInitGL} to initialize a state object and
an overlay logo. This function is only called if an OpenGL context was
created and made current:

{\footnotesize\begin{lstlisting}
bool Window::configInitGL( const uint32_t initID )
{
    if( !eq::Window::configInitGL( initID ))
        return false;

    glLightModeli( GL_LIGHT_MODEL_LOCAL_VIEWER, 1 );
    glEnable( GL_CULL_FACE ); // OPT - produces sparser images in DB mode
    glCullFace( GL_BACK );

    EQASSERT( !_state );
    _state = new VertexBufferState( getObjectManager( ));

    const Config*   config   = static_cast< const Config* >( getConfig( ));
    const InitData& initData = config->getInitData();

    if( initData.showLogo( ))
        _loadLogo();

    if( initData.useGLSL() )
        _loadShaders();

    return true;
}
\end{lstlisting}}%>>

The state object is used to handle the creation of OpenGL objects in a
multipipe, multithreaded execution environment. It uses the object
manager of the \textsf{eq::Window}, which is described in detail in
\sref{sObjectManager}.

The logo texture is loaded from the file system and bound to a texture
ID used later by the channel for rendering. A code listing is omitted,
since the code consists of standard OpenGL calls and is not
Equalizer-specific.

The window exit happens in the reverse order of the
initialization. First, \textsf{configExitGL} is called to de-initialize
OpenGL, followed by \textsf{configExitOSWindow} which de-initializes the
drawable and context and deletes the \textsf{OSWindow} allocated in
\textsf{configInitOSWindow}.

The window OpenGL exit function of \textsf{eqPly} de-allocates all
OpenGL objects. The object manager does not delete the object in its
destructor, since it does not know if an OpenGL context is still
current.

{\footnotesize\begin{lstlisting}
bool Window::configExitGL()
{
    if( _state )
        _state->deleteAll();

    delete _state;
    _state = 0;

    return eq::Window::configExitGL();
}
\end{lstlisting}}

\subsubsection{\label{sObjectManager}Object Manager}

The object manager is, strictly speaking, not a part of the window. It
is mentioned here since the \textsf{eqPly} window uses an object manager.

The state object in \textsf{eqPly} gathers all rendering state, which
includes an object manager for OpenGL object allocation.

The object manager (OM) is a utility class and can be used to manage
OpenGL objects across shared contexts. Typically one OM is used for each
set of shared contexts and spawns all contexts of a single
GPU\footnote{\link{http://www.equalizergraphics.com/documents/design/objectManager.html}}.

The OM is a template class. The template type is the key used to
identify objects. The same key is used by all contexts to get the
OpenGL name of an object.

Each \textsf{eq::Window} has an object manager with the key type
\textsf{const void*}, for as long as it is initialized. Each window can
have a shared context window. The OM is shared with this shared context
window. The shared context window is set by default to the first window
of each pipe, and therefore the OM will be shared between all windows of
a pipe. The method \textsf{eq::Window::setSharedContextWindow} can be
used to set up a different context sharing.

\textsf{eqPly} uses the window's object manager in the rendering code to
obtain the OpenGL objects for a given data item. The address of the data
item to be rendered is used as the key.

For the currently supported types of OpenGL objects please refer to the
API documentation on the Equalizer website. For each object, the
following functions are available:

\begin{description}
\item[supportsObjects()] returns true if the usage for this particular
  type of objects is supported. For objects available in OpenGL 1.1 or
  earlier, this function is not implemented.
\item[getObject( key )] returns the object associated with the given
  key, or FAILED.
\item[newObject( key )] allocates a new object for the given
  key. Returns FAILED if the object already exists or if the allocation
  failed.
\item[obtainObject( key )] convenience function which gets or obtains
  the object associated with the given key. Returns FAILED only if the
  object allocation failed.
\item[deleteObject( key )] deletes the object.
\end{description}

\subsection{Channel}

The channel is the heart of the application, it contains the actual
rendering code. The channel is used to perform the various rendering
operations for the compounds. Each channel has a set of task methods to
execute the clear, draw, readback and assemble stages needed to render a
frame.

\subsubsection{Initialization and Exit}

During channel initialization, the near and far planes are set to
reasonable values to contain the whole model. During rendering, the near
and far planes are adjusted dynamically to the current model position:

{\footnotesize\begin{lstlisting}
bool Channel::configInit( const uint32_t initID )
{
    setNearFar( 0.1f, 10.0f );
    return true;
}
\end{lstlisting}}

\subsubsection{Rendering}

The central rendering routine is \textsf{Channel::frameDraw}. This
routine contains the application's OpenGL rendering code, which is being
rendered using the contextual information provided by Equalizer. As most
of the other task methods, \textsf{frameDraw} is called in parallel by
Equalizer on all pipe threads in the configuration. Therefore the
rendering must not write to shared data, which is the case for all major
scene graph implementations.

In \textsf{eqPly}, the OpenGL context is first set up using various
\textsf{apply} convenience methods from the base Equalizer channel
class. Each of the \textsf{apply} methods uses the corresponding
\textsf{get} method(s) and then calls the appropriate OpenGL
function(s). It is also possible to just query the values from Equalizer
using the \textsf{get} methods, and use them to set up the OpenGL state
appropriately, for example by passing the parameters to the renderer used
by the application.

For example, the implementation for \textsf{eq::Channel::applyBuffer}
does set up the correct rendering buffer and color mask, which depends
on the current eye pass and possible anaglyphic stereo parameters:

{\footnotesize\begin{lstlisting}
void eq::Channel::applyBuffer()
{
    glReadBuffer( getReadBuffer( ));
    glDrawBuffer( getDrawBuffer( ));

    const ColorMask& colorMask = getDrawBufferMask();
    glColorMask( colorMask.red, colorMask.green, colorMask.blue, true );
}
\end{lstlisting}}

The contextual information has to be used to render the view as
expected by Equalizer. Failure to use certain information will result in
incorrect rendering for some or all configurations. The channel render
context consist of:

\begin{description}
\item[Buffer] The OpenGL read and draw buffer as well as color mask.
  These parameters are influenced by the current eye pass, eye
  separation and anaglyphic stereo settings.
\item[Viewport] The two-dimensional pixel viewport restricting the
  rendering area within the channel. For correct operations, both
  \textsf{glViewport} and \textsf{glScissor} have to be used. The pixel
  viewport is influenced by the destination channel's viewport
  definition and compound viewports set for sort-first/2D decompositions.
\item[Frustum] The same frustum parameters as defined by
  \textsf{glFrustum}. Typically the frustum used to set up the OpenGL projection
  matrix. The frustum is influenced by the destination channel's view
  definition, compound viewports, head matrix and the current eye pass. If the
  channel has a subpixel parameter, the frustum will be jittered before it is
  applied. Please refer to \ref{sFSAA} for more information.
\item[Head Transformation] A transformation matrix positioning the
  frustum. This is typically an identity matrix and is used for off-axis
  frusta in immersive rendering. It is normally used to set up the
  `view' part of the modelview matrix, before static light sources are
  defined.
\item[Range] A one-dimensional range with the interval [0..1]. This
  parameter is optional and should be used by the application to render
  only the appropriate subset of its data. It is influenced by the
  compound range attribute.
\end{description}

The \textsf{frameDraw} method in \textsf{eqPly} calls the
\textsf{frameDraw} method from the parent class, the Equalizer
channel. The default \textsf{frameDraw} method uses the apply
convenience functions to setup the OpenGL state for all render context
information, except for the range which will be used later during
rendering:

{\footnotesize\begin{lstlisting}
void eq::Channel::frameDraw( const uint32_t frameID )
{
    applyBuffer();
    applyViewport();

    glMatrixMode( GL_PROJECTION );
    glLoadIdentity();
    applyFrustum();

    glMatrixMode( GL_MODELVIEW );
    glLoadIdentity();
    applyHeadTransform();
}

void Channel::frameDraw( const uint32_t frameID )
{
    // Setup OpenGL state
    eq::Channel::frameDraw( frameID );
\end{lstlisting}}

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/DB.png}
  {\caption{\label{fDBDest}Destination View of a DB Compound using
      Demonstrative Coloring}}
\end{wrapfigure}
After the basic view setup, a directional light is configured, and the
model is positioned using the camera parameters from the frame data. The
camera parameters are transported using the frame data to ensure
that all channels render a given frame using the same position.

Three different ways of coloring the object are possible: Using the
colors of the mode, using a unique per-channel color to demonstrate the
decomposition as shown in \fig{fDBDest}, or using solid white for
anaglyphic stereo. The model colors are per-vertex and are set during
rendering, whereas the unique per-channel color is set in
\textsf{frameDraw} for the whole model:

{\footnotesize\begin{lstlisting}
    glLightfv( GL_LIGHT0, GL_POSITION, lightPosition );
    glLightfv( GL_LIGHT0, GL_AMBIENT,  lightAmbient  );
    glLightfv( GL_LIGHT0, GL_DIFFUSE,  lightDiffuse  );
    glLightfv( GL_LIGHT0, GL_SPECULAR, lightSpecular );

    glMaterialfv( GL_FRONT, GL_AMBIENT,   materialAmbient );
    glMaterialfv( GL_FRONT, GL_DIFFUSE,   materialDiffuse );
    glMaterialfv( GL_FRONT, GL_SPECULAR,  materialSpecular );
    glMateriali(  GL_FRONT, GL_SHININESS, materialShininess );

    const FrameData& frameData = getFrameData();

    glPolygonMode( GL_FRONT_AND_BACK, 
                   frameData.useWireframe() ? GL_LINE : GL_FILL );

    const eq::Vector3f& translation = frameData.getCameraTranslation();

    glMultMatrixf( frameData.getCameraRotation().array );
    glTranslatef( translation.x(), translation.y(), translation.z() );
    glMultMatrixf( frameData.getModelRotation().array );

    if( frameData.getColorMode() == COLOR_DEMO )
    {
        const eq::Vector3ub color = getUniqueColor();
        glColor3ub( color.r(), color.g(), color.b() );
    }
    else
        glColor3f( .75f, .75f, .75f );
\end{lstlisting}}

Finally the model is rendered. If the model was not loaded during node
initialization, a quad is drawn in its place:

{\footnotesize\begin{lstlisting}
    const Model* model = _getModel();
    if( model )
    {
        _drawModel( model );
    }
    else
    {
        glNormal3f( 0.f, -1.f, 0.f );
        glBegin( GL_TRIANGLE_STRIP );
        glVertex3f(  .25f, 0.f,  .25f );
        glVertex3f(  .25f, 0.f, -.25f );
        glVertex3f( -.25f, 0.f,  .25f );
        glVertex3f( -.25f, 0.f, -.25f );
        glEnd();
    }

    [...]
}
\end{lstlisting}}

To draw the model, a helper class for view frustum culling is set up
using the view frustum from Equalizer and the camera position from the
frame data. The frustum helper computes the six frustum planes from the
projection and modelView matrices. During rendering, the bounding
spheres of the model are tested against these planes to determine the
visibility with the frustum.

Furthermore, the render state from the window and the database range
from the channel is obtained. The render state manages display list or
VBO allocation:

{\footnotesize\begin{lstlisting}
void Channel::_drawModel( const Model* model )
{
    Window*                  window    = static_cast<Window*>( getWindow() );
    mesh::VertexBufferState& state     = window->getState();

    const Pipe*              pipe      = static_cast<Pipe*>( getPipe( ));
    const FrameData&         frameData = pipe->getFrameData();

    const eq::Range&         range     = getRange();
    vmml::FrustumCullerf     culler;

    if( frameData.getColorMode() == COLOR_MODEL && model->hasColors( ))
        state.setColors( true );
    else
        state.setColors( false );

    _initFrustum( culler, model->getBoundingSphere( ));
    model->beginRendering( state );
\end{lstlisting}}

The model data is spatially organized in a 3-dimensional
kD-tree\footnote{See also \link{http://en.wikipedia.org/wiki/Kd-tree}}
for efficient view frustum culling. When the model is loaded by
\textsf{Node::config\-Init}, it is preprocessed into the kD-tree. During
this preprocessing step, each node of the tree gets a database range
assigned. The root node has the range [0, 1], its left child [0, 0.5]
and its right child [0.5, 1], and so on for all nodes in the tree. The
preprocessed model is saved in a binary format for accelerating
subsequent loading.

\begin{wrapfigure}{r}{.618\textwidth}
%\begin{figure}[h!t]\center
  \includegraphics[width=.618\textwidth]{images/render.pdf}
  {\caption{\label{fRender}Main Render Loop}}
\end{wrapfigure}
The rendering loop maintains a list of candidates to render, which
initially contains the root node. Each candidate of this list is tested
for full visibility against the frustum and range, and rendered if
visible. It is dropped if it is fully invisible or fully out of
range. If it is partially visible or partially in range, the children of
the node are added to the candidate list. \fig{fRender} shows a flow
chart of the rendering algorithm, which performs efficient view frustum
and range culling.

The actual rendering uses display lists or vertex buffer objects. These
OpenGL objects are allocated using the object manager. The rendering is
done by the leaf nodes, which are small enough to store the vertex
indices in a \textsf{short} value for optimal performance with VBO's.
The leaf nodes reuse the objects stored in the object manager, or create
and set up new objects if it was not yet set up. Since one object
manager is used per thread (pipe), this allows a thread-safe sharing of
the compiled display lists or VBO's across all windows of a pipe.

The main rendering loop in \textsf{eqPly} looks like this:

{\footnotesize\begin{lstlisting}
    model->beginRendering( state );

    // start with root node
    vector< const VertexBufferBase* > candidates;
    candidates.push_back( model );

    while( !candidates.empty() )
    {
        const VertexBufferBase* treeNode = candidates.back();
        candidates.pop_back();

        // completely out of range check
        if( treeNode->getRange()[0] >= range.end || 
            treeNode->getRange()[1] < range.start )
            continue;

        // bounding sphere view frustum culling
        switch( culler.testSphere( treeNode->getBoundingSphere() ) )
        {
            case vmml::VISIBILITY_FULL:
                // if fully visible and fully in range, render it
                if( treeNode->getRange()[0] >= range.start && 
                    treeNode->getRange()[1] < range.end )
                {
                    treeNode->render( state );
                    break;
                }
                // partial range, fall through to partial visibility
            case vmml::VISIBILITY_PARTIAL:
            {
                const VertexBufferBase* left = treeNode->getLeft();
                const VertexBufferBase* right = treeNode->getRight();

                if( !left && !right )
                {
                    if( treeNode->getRange()[0] >= range.start )
                        treeNode->render( state );
                    // else drop, to be drawn by 'previous' channel
                }
                else
                {
                    if( left )
                        candidates.push_back( left );
                    if( right )
                        candidates.push_back( right );
                }
                break;
            }
            case vmml::VISIBILITY_NONE:
                // do nothing
                break;
        }
    }

    model->endRendering( state );
}
\end{lstlisting}}

\subsubsection{Assembly}

Like most applications, eqPly can use the default implementation of the
\textsf{frameReadback} and \textsf{frameAssemble} task methods.
\textsf{frameReadback} is overwritten for an optimization since eqPly
does not need the alpha channel on the destination view.
The output frames are flagged to ignore it, which allows the compressor to
drop 25\% of the data during image transfer:

{\footnotesize\begin{lstlisting}
void Channel::frameReadback( const uint32_t frameID )
{
    // OPT: Drop alpha channel from all frames during network transport
    const eq::FrameVector& frames = getOutputFrames();
    for( eq::FrameVector::const_iterator i = frames.begin(); 
         i != frames.end(); ++i )
    {
        (*i)->setAlphaUsage( false );
    }

    eq::Channel::frameReadback( frameID );
}
\end{lstlisting}}

The \textsf{frameAssemble} method is overwritten for the use of Subpixel
compound with idle software anti-aliasing, as described in \ref{sFSAA}.

\section{Advanced Features}

This section discusses important features not covered by the previous
\textsf{eqPly} section. Where possible, code examples from the Equalizer
distribution are used to illustrate usage of the feature. Its
purpose is to provide information on how to address a typical problem or
use case when developing an Equalizer-based application.

\subsection{\label{sEventHandling}Event Handling}

Event handling requires flexibility. On one hand, the implementation
differs slightly for each operating and window system due to conceptual
differences in the specific implementation. On the other hand, each
application and widget set has its own model on how events are to be
handled. Therefore, event handling in Equalizer is customizable at any
stage of the processing, to the extreme of making it possible to disable
all event-related code in Equalizer. In this aspect, Equalizer
substantially differs from GLUT, which imposes an event model and hides
most of the event handling in \textsf{glutMainLoop}. More information on
event handling can be found on the Equalizer website\footnote{see
  \link{http://www.equalizergraphics.com/documents/design/eventHandling.html}}.

The default implementation provides a convenient, easily accessible
event framework, while allowing all necessary customizations. It gathers
all events from all node processes in the main thread of the
application, so that the developer only has to implement
\textsf{Config::processEvent} to update its data based on the
preprocessed, generic keyboard and mouse events. It is very easy to use
and similar to a GLUT-based implementation.

\subsubsection{Threading}

Events are received and processed by the pipe thread a window belongs
to. For AGL, Equalizer internally forwards the event from the main
thread, where it is received, to the appropriate pipe thread.

\subsubsection{\label{sMessagePump}Message Pump}

To dispatch events, Equalizer 'pumps' the native events. On WGL and GLX,
this happens on each thread with windows, whereas on AGL it happens on
the main thread and each pipe thread. By default, Equalizer pumps these
events automatically for the application in-between executing task
methods.

The method \textsf{Pipe::useMessagePump} can be overridden to return
\textsf{false} to not enable message dispatch for the respective thread.

The method \textsf{useMessagePump} is called by Equalizer during
application and pipe thread initialization (before configInit!) to
determine if the pipe's thread shall automatically dispatch OS
events. For AGL, this affects the node threads and pipe thread, since
the node thread message pump needs to dispatch the events to the pipe
thread event queue.

If the application disables message pumping in Equalizer, it has to make
sure the events are pumped externally, as it often done within external
widget sets such as Qt.

\subsubsection{Event Data Flow}

Events are received by an event handler. The event handler finds the
\textsf{eq::OSWindow} associated to the event. It then creates a generic
\textsf{Event}, which holds the event data in an independent
format. The original native event and this generic \textsf{Event}
form the \textsf{OSWindowEvent}, which is passed to the concrete
\textsf{OSWindow} for processing.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/eventFilter.pdf}
  {\caption{\label{fEventProcessing}Event Processing}}
\end{wrapfigure}
The purpose of the \textsf{OSWindow} processing method,
\textsf{processEvent}, is to perform window-system specific event
processing. For example, \textsf{AGLWindow::processEvent} calls
\textsf{aglUpdateContext} whenever a resize event is received. For
further, generic processing, the \textsf{Event} is passed on to
\textsf{Window::processEvent}. This \textsf{Event} no longer contains
the native event.

\textsf{Window::processEvent} is responsible for handling
the event locally and for translating it into a generic
\textsf{ConfigEvent}. The config events are send to the application
thread using \textsf{Config::sendEvent}.

Events send using \textsf{Config::sendEvent} are queued in the
application thread. After a frame has been finished,
\textsf{Config::finishFrame} calls \textsf{Config::handleEvents}. The
default implementation of this method provides non-blocking event
processing, that is, it calls \textsf{Config::handleEvent} for each
queued event. By overriding \textsf{handleEvents}, event-driven
execution can easily be implemented.

If the event was processed, \textsf{processEvent} has to return
\textsf{true}. If \textsf{false} is returned to the event handler, the
event will be passed to the previously installed, window-system-specific
event handling function.

\fig{fEventProcessing} shows the overall data flow of an event.

\subsubsection{Default Implementation}

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/eventUML.pdf}
  {\caption{\label{fEventUML}UML Diagram for Event Handling Classes}}
\end{wrapfigure}
Equalizer provides an \textsf{AGLEvent\-Handler},
\textsf{GLXEventHandler} and \textsf{WGL\-Event\-Handler}, which handle
events for an \textsf{AGLWindowIF}, \textsf{GLXWindowIF} and
\textsf{WGL\-Win\-dowIF}, respectively. \fig{fEventUML} illustrates the
class hierarchy for event processing.

The concrete implementation of these window interfaces is responsible
for setting up and de-initializing event handling.

Carbon events issued for AGL windows are initially receive by the node
main thread, and automatically dispatched to the pipe thread. The pipe
thread will dispatch the event to the appropriate event handler. One
\textsf{AGLEventHandler} per window is used, which is created during
\textsf{AGLWindow::configInit}. The event handler installs a Carbon
event handler for all important events. The event handler uses an
\textsf{AGLWindowEvent} to pass the Carbon \textsf{EventRef} to
\textsf{AGLWindowIF::processEvent}. During window exit, the installed
Carbon handler is removed when the window's event handler is deleted. No
event handling is set up for PBuffers.

GLX uses one X11 Display connection for all windows on a GPU, which is
managed by the \textsf{eq::Pipe}. For each thread, one
\textsf{GLXEventHandler} is allocated. The pipe registers the display
connection with this per-thread event handler. Events for windows
created using this display connection are automatically received by the
event handler. No per-window event handling setup needs to be
done. PBuffers are handled in the same way as window drawables.

The glX event handler iterates over all windows of the registered pipe
to find the originating \textsf{GLXWindowIF}, which is then used to
process the event in a fashion similar to AGL. The
\textsf{GLXWindowEvent} passes the \textsf{XEvent} to
\textsf{GLXWindowIF::\-process\-Event}.

Each \textsf{WGLWindow} allocates one \textsf{WGLEventHandler} when the
window handle is set. The WGLEventHandler passes the native event
parameters \textsf{uMsg}, \textsf{wParam} and \textsf{lParam} to
\textsf{WGL\-Win\-dowIF::processEvent} as part of the
\textsf{WGLWindowEvent}. No event handling is set up for PBuffers.

Later Equalizer versions will introduce \textsf{Pipe::pro\-cess\-Event} and
\textsf{PipeEvent} to communicate pipe-specific events, e.g., monitor
resolution changes. It is also likely that an \textsf{OSPipe}
abstraction, similar to the \textsf{OSWindow}, will be created to
separate the window system specific code from the \textsf{eq::Pipe}.

\subsubsection{Custom Events in eqPixelBench}

The \textsf{eqPixelBench} example is a benchmark program to measure the
pixel transfer rates from and to the framebuffer of all channels within
a configuration. It uses custom config events to send the gathered data
to the application. It is much simpler than the \textsf{eqPly} example
since it does not provide any useful rendering or user interaction.

The rendering routine of \textsf{eqPixelBench} in
\textsf{Channel::frameDraw} loops through a number of pixel formats and
types. For each of them, it measures the time to readback and assemble a
full-channel image. The format, type, size and time is recorded in a
config event, which is sent to the application.

The \textsf{ConfigEvent} derives from the \textsf{eq::ConfigEvent}
structure and has the following definition:

{\footnotesize\begin{lstlisting}
struct ConfigEvent : public eq::ConfigEvent
{
public:
    enum Type
    {
        READBACK = eq::Event::USER,
        ASSEMBLE
    };

    ConfigEvent()
        {
            size = sizeof( ConfigEvent );
        }

    // channel name is in user event data
    char           formatType[64];
    vmml::Vector2i area;
    float          msec;
};
\end{lstlisting}}

The \textsf{Config::sendEvent} method transmits an
\textsf{eq::ConfigEvent} or derived class to the application. The
ConfigEvent has to be a C-type structure, and its \textsf{size}
member has to be set to the full size of the event to be transmitted.
Each event has a type which is used to identify it by the config 
processing function.

User-defined types start at \textsf{eq::ConfigEvent::USER}, and the
member variable \textsf{ConfigEvent::user} can be used to store up to
\textsf{EQ\_USER\_EVENT\_SIZE}\footnote{currently 32 bytes} bytes. In
this space, the channel's name is stored. Additional variables are used
to transport the pixel format and type, the size and the time it took
for rendering.

On the application end, \textsf{Config::handleEvent} uses the
\textsf{ostream} operator for the derived config event to output these
events in a nicely formatted way:

{\footnotesize\begin{lstlisting}
std::ostream& operator << ( std::ostream& os, const ConfigEvent* event );
...
bool Config::handleEvent( const eq::ConfigEvent* event )
{
    switch( event->type )
    {
        case ConfigEvent::READBACK:
        case ConfigEvent::ASSEMBLE:
            cout << static_cast< const ConfigEvent* >( event ) << endl;
            return true;

        default:
            return eq::Config::handleEvent( event );
    }
}
\end{lstlisting}}%>>

\subsection{\label{sThreads}Thread Synchronization}

Equalizer applications use multiple, potentially asynchronous execution
threads. The default execution model is focused on making the porting of
existing applications as easy as possible, as described in
\sref{sExecution}. The default, per-node thread synchronization provided
by Equalizer can be relaxed by advanced applications to gain better
performance through higher asynchronicity.

\subsubsection{Threads}

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/threads.pdf}
  {\caption{\label{fThreads}Threads within one Node Process}}
\end{wrapfigure}
The application or node main thread is the primary thread of each
process and executes the \textsf{main} function. The application and
render clients initialize the local node for communications with other
nodes, including the server, using \textsf{Client::initLocal}.

During this initialization, Equalizer creates and manages two threads
for communication, the receiver thread and the command thread. Normally
no application code is executed from these two threads.

The receiver thread manages the network connections to other nodes and
receives data. It dispatches the received data either to the application
threads, or to the command thread.

The command thread processes internal Equalizer request from other
nodes, for example during \textsf{eq::net::Object} mapping. In some
special cases the command thread executes application code, for example
when a remote node maps a static or unbuffered object,
\textsf{Object::getInstanceData} is called from the command thread.

The receiver and command thread are terminated when the application
stops network communications using \textsf{Client::exitLocal}.

During config initialization, one pipe thread is created for each
pipe. The pipe threads execute all render task methods for this pipe,
and therefore executes the application's rendering code. The pipe
threads are terminated during \textsf{Config::exit}.

The rest of this section discusses the thread synchronization between
the main thread and the pipe threads.

\subsubsection{\label{sThreadModel}Thread Synchronization Models}
Equalizer supports three threading models, which can be set
programmatically by the application or through the configuration file
format. Applications typically hard-code their threading model. The file
format is commonly used to change the threading model for benchmarking
and experimentation.

The following thread synchronization models are implemented:
\begin{description}
\item[ASYNC:] No synchronization happens between render threads, except
  for synchronizing the finish of frame \textsf{current-latency}. The
  \textsf{eqPly} and \textsf{eVolve} Equalizer examples use this
  threading model, by setting it in \textsf{Node::configInit}. This
  synchronization model provides the best performance and should be used
  by all applications which multi-buffer all non-static, frame-specific
  data.
\item[DRAW\_SYNC:] Additionaly to the synchronization of the async
  thread model, all local render threads are synchronized, so that the
  draw operations happen synchronously with the node main loop. This
  model allows to use the same database for rendering, and safe
  modifications of this database are possible from the node thread,
  since the pipe threads do not execute any rendering tasks between
  frames. This is the default threading model. This threading model
  should be used by applications which keep one copy of the scene graph
  per node.
\item[LOCAL\_SYNC:] Additionaly to the synchronization of the async
  thread model, all local frame operations, including readback, assemble
  and swap buffer are synchronized with the node main loop. This
  threading model should be used by applications which need to access
  non-buffered, frame-specific data after rendering, e.g., during
  \textsf{Channel::frameAssemble}.
\end{description}

\fig{fThreadModels} illustrates the synchronization and task execution
for the thread synchronization models. Please note that the thread
synchronization synchronizes all pipe render threads on a
\textbf{single} node with the node's main thread. The per-node frame
synchronization does not break the asynchronous execution across
nodes.

\begin{figure}[h!t]\center
  \includegraphics[width=.9\textwidth]{images/threadModels.pdf}
  {\caption{\label{fThreadModels}Async, draw\_sync and local\_sync
    Thread Synchronization Models}}
\end{figure}

\subsubsection{\label{sAdvDPlex}DPlex Compounds}

DPlex decomposition requires multiple frames to be rendered
concurrently. Applications using the thread synchronization models
\textsf{draw\_sync} (the default) or \textsf{local\_sync} have to use
one render client process per GPU to benefit from DPlex task
decomposition.

Non-threaded pipes should not be used for DPlex source and destination
channels. Applications using the \textsf{local\_sync} thread model can
not benefit from DPlex if the application node uses a DPlex source or
destination channel.

Applications using the \textsf{async} thread synchronization model can
fully profit from DPlex using multiple render threads on a multi-GPU
system. For these applications, all render threads run asynchronously
and can render different frames at the same time.

Synchronizing the draw operations between multiple pipe render threads,
and potentially the application thread, breaks DPlex decomposition. At
any given time, only one frame can be rendered from the same
process. The speedup of DPlex relies however on the capability to
render different frames concurrently.

If one process per GPU is configured, draw-synchronous applications can
scale the performance using DPlex compounds. The processes are not
synchronized with each other, since each process keeps its own version
of the scene data.

\subsubsection{Thread Synchronization in Detail}

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/frameSync.pdf}
  {\caption{\label{fFrameSync}Per-Node Frame Synchronization}}
\end{wrapfigure}
The application has extended control over the task synchronization
during a frame. Upon \textsf{Config::startFrame}, Equalizer invokes the
\textsf{frameStart} task methods of the various entities. The entities
unlock all their children by calling \textsf{startFrame}, e.g.,
\textsf{Node::frameStart} has to call \textsf{Node::startFrame}
to unlock the pipe threads. Note that certain \textsf{startFrame} calls,
e.g., \textsf{Window::startFrame}, are currently empty since the
synchronization is implicit due to the sequential execution within the
thread.

Each entity uses \textsf{waitFrame\-Started} to block on the parent's
\textsf{startFrame}, e.g., \textsf{Pipe::\-frame\-Start} calls
\textsf{Node::wait\-Frame\-Started} to wait for the corresponding
\textsf{Node::start\-Fra\-me}. This explicit synchronization allows to
update non-critical data before synchronizing with
\textsf{waitFrameStarted}, or after unlocking using
\textsf{start\-Fra\-me}. \fig{fFrameSync} illustrates this
synchronization model.

At the end of the frame, two similar sets of synchronization methods are
used. The first set synchronizes the local execution, while the second
set synchronizes the global execution.

The local synchronization consists of \textsf{releaseFrameLocal} to
unlock the local frame, and of \textsf{waitFrameLocal} to wait for the
unlock. For the default synchronization model \textsf{sync\_draw},
Equalizer uses the task method \textsf{frameDrawFinish} which is called
on each resource after the last \textsf{Channel::frameDraw} invocation
for this frame. Consequently, \textsf{Pipe::frameDraw\-Fi\-nish} calls
\textsf{Pipe::\-release\-Frame\-Lo\-cal} to signal that it is done
drawing the current frame, and \textsf{Node::frameDrawFinish} calls
\textsf{Pipe::waitFrameLocal} for each of its pipes to block the node
thread until the current frame has been drawn.

\fig{fFrameSync} illustrates the local frame synchronization.

The second, global synchronization is used for the frame completion
during \textsf{Config::finishFrame}, which causes \textsf{frameFinish}
to be called on all entities, passing the oldest frame number, i.e., frame
\textsf{current-latency}. The \textsf{frameFinish} task methods have to
call \textsf{releaseFrame} to signal that the entity is done with the
frame. The release causes the parent's \textsf{frameFinish} to be
invoked, which is synchronized internally. Once all
\textsf{Node::releaseFrame} have been called,
\textsf{Config::finishFrame} returns.

\begin{figure}[ht!]\center
  \includegraphics[width=\textwidth]{images/mainloop.pdf}
  {\caption{\label{fMainLoop}Synchronization of Frame Tasks}}
\end{figure}

\fig{fMainLoop} outlines the synchronization for the application, node
and pipe classes for an application node and one render client when
using the default \textsf{draw\_sync} thread model. Please note that
\textsf{Config::finishFrame} does block until the current frame has been
released locally and until the frame \textsf{current - latency} has been
released by all nodes. The window and channel synchronization are
similar and omitted for simplicity.

It is absolutely vital for the execution that \textsf{Node::start\-Fra\-me}
and \textsf{Node::release\-Fra\-me} are called, respectively. The default
implementations of the node task methods do take care of that.

\subsection{OpenGL Extension Handling}

Equalizer uses GLEW\footnote{\link{http://glew.sourceforge.net}} for
OpenGL extension handling, particularly the GLEW MX implementation
providing multi-context support.

Each \textsf{eq::Window} has a \textsf{GLEWContext}. This context can be
obtained by using \textsf{glewGetContext} on the window or channel. GLEW
MX uses this function to dispatch the functions to the correct
context. Equalizer (re-)initializes the GLEW context whenever a new
OpenGL context is set on the window.

Extended OpenGL functions called from a window or channel instance can
be called directly. GLEW will call the object's \textsf{glewGetContext}
to obtain the correct context:

{\footnotesize\begin{lstlisting}
void eqPly::Channel::_drawModel( const Model* model )
{
    ...
    glUseProgram( program );
    ...
}
\end{lstlisting}}

Functions called from another place need to define a macro or function
\textsf{glewGetContext} that returns the pointer to the GLEWContext of
the appropriate window:

{\footnotesize\begin{lstlisting}
// state has GLEWContext* from window
#define glewGetContext state.glewGetContext

/*  Set up rendering of the leaf nodes.  */
void VertexBufferLeaf::setupRendering( VertexBufferState& state,
                                       GLuint* data ) const
{
    ...
    glBindBuffer( GL_ARRAY_BUFFER, data[VERTEX_OBJECT] );
    glBufferData( GL_ARRAY_BUFFER, _vertexLength * sizeof( Normal ),
                    &_globalData.normals[_vertexStart], GL_STATIC_DRAW );
    ...
}
\end{lstlisting}}

\subsection{Advanced Window Initialization}

This section explains window initialization in detail. It discusses in
detail the handling of the different window systems. Window systems are
integrated through the \textsf{OSWindow} interface, which externalizes
the details of the windowing API from the \textsf{eq::Window}
implementation and facilitates the integration with new windowing API's
and custom applications.

Equalizer provides sample implementations for the supported window
systems. These sample implementations are intended to be sub-classed and
various steps in the window initialization can be overwritten and
customized.

An application typically choses to subclass the sample implementation if
only minor tweaks are needed for integration. For major changes or new
window systems, it is often easier to subclass directly from
\textsf{OSWindow} and implement the abstract methods of this interface
class.

The method \textsf{Window::configInitOSWindow} is used to instantiate
and initialize the \textsf{OSWindow} implementation during config
initialization. After a successful \textsf{OSWindow} initialization,
\textsf{Window::configInitGL} is called for the generic OpenGL state
setup.

Since window initialization is notoriously error-prone and hard to
debug, the sample implementation propagates the reason for errors
from the render clients back to the application. The \textsf{Pipe} and
\textsf{Window} classes have a \textsf{setErrorMessage} method, which is
used to set an error string. This string is passed to the
\textsf{Config} instance on the application node, where it can be
retrieved using \textsf{getErrorMessage}.

The sample implementations \textsf{AGLWindow}, \textsf{GLXWindow} and
\textsf{WGLWindow} all have similar, override-able methods for all
sub-tasks. This allows partial customization, without the need of
rewriting tedious window initialization code, e.g., the OpenGL pixel
format selection.

\subsubsection{\label{sDrawableConfig}Drawable Configuration}

OpenGL drawables have a multitude of buffer modes. A drawable might be
single-buffered, double-buffered or quad-buffered, have auxiliary image
planes such as stencil, accumulation and depth buffer or multisampling.

The OpenGL drawable is configured using window attributes. These
attributes are used by the method choosing the pixel format (or visual
in X11 speak) to select the correct drawable configuration.

Window attributes can either be configured through the configuration
file (see \aref{aFileFormat}), or programmatically. In the configuration
file, modes are selected which are not application-specific, for example
stereo formats for active stereo displays.

Applications which require certain drawable attributes can set the
corresponding window attribute hint during window initialization. The
Equalizer volume rendering example, \textsf{eVolve}, is such an
example. It does need alpha planes for rendering and compositing. The
window initialization of \textsf{eVolve} sets the attribute before
calling the default initialization method of Equalizer:

{\footnotesize\begin{lstlisting}
bool eVolve::Window::configInit( const uint32_t initID )
{
    // Enforce alpha channel, since we need one for rendering
    setIAttribute( IATTR_PLANES_ALPHA, 8 );

    return eq::Window::configInit( initID );
}
\end{lstlisting}}

\subsubsection{AGL Window Initialization}

AGL initialization happens in three steps: choosing a pixel format,
creating the context and creating a drawable.

Most AGL and Carbon calls are not thread-safe. The Equalizer methods
calling these functions use \textsf{Global::enterCarbon} and
\textsf{Global::leaveCarbon} to protect the API calls. Please refer to
\sref{sCarbonThread} for more details.

The pixel format is chosen based on the window's attributes. Some
attributes set to auto, e.g., stereo, cause the method first to request
the feature and then to back off and retry if it is not available. The
pixel format returned by \textsf{chooseAGLPixelFormat} has to be
destroyed using \textsf{destroyAGLPixelFormat}. When no matching pixel
format is found, \textsf{chooseAGLPixelFormat} returns \textsf{0} and
the AGL window initialization returns with a failure.

The context creation also uses the global Carbon lock. Furthermore, it
sets up the swap buffer synchronization with the vertical retrace, if
enabled by the corresponding window attribute hint. Again the window
initialization fails if the context could not be created.

The drawable creation method \textsf{configInitAGLDrawable} calls either
\textsf{configInitAGLFullscreen}, \textsf{configInitAGLWindow} or \textsf{configInitAGLPBuffer}

The top-level AGL window initialization code therefore looks as follows:

{\footnotesize\begin{lstlisting}
bool AGLWindow::configInit( )
{
    AGLPixelFormat pixelFormat = chooseAGLPixelFormat();
    if( !pixelFormat )
        return false;

    AGLContext context = createAGLContext( pixelFormat );
    destroyAGLPixelFormat ( pixelFormat );
    setAGLContext( context );

    if( !context )
        return false;

    return configInitAGLDrawable();
}
\end{lstlisting}}

\subsubsection{GLX Window Initialization}

GLX initialization is very similar to AGL initialization. Again the
steps are: choose visual (pixel format), create OpenGL context and then
create drawable. The only difference is that the data returned by
\textsf{chooseXVisualInfo} has to be freed using \textsf{XFree}:

{\footnotesize\begin{lstlisting}
bool GLXWindow::configInit( )
{
    XVisualInfo* visualInfo = chooseXVisualInfo();
    if( !visualInfo )
        return false;

    GLXContext context = createGLXContext( visualInfo );
    setGLXContext( context );

    if( !context )
        return false;

    const bool success = configInitGLXDrawable( visualInfo );
    XFree( visualInfo );

    if( success && !_xDrawable )
    {
        _window->setErrorMessage( 
            "configInitGLXDrawable did not set a X11 drawable");
        return false;
    }

    return success;    
}
\end{lstlisting}}

\subsubsection{WGL Window Initialization}

The WGL initialization requires another order of operations compared to
AGL or GLX. The following functions are used to initialize a WGL window:

\begin{enumerate}
\item\textsf{initWGLAffinityDC} is used to set up an affinity device
  context, which might be needed for window creation. The WGL window
  tracks potentially two device context handles, one for OpenGL context
  creation (the affinity DC), and one for \textsf{swapBuffers} (the
  window's DC).
\item\textsf{chooseWGLPixelFormat} chooses a pixel format based on the
  window attributes. If no device context is given, it uses the system
  device context. The chosen pixel format is set on the passed device
  context.
\item\textsf{configInitWGLDrawable} creates the drawable. The device
  context passed to \textsf{configInitWGLDrawable} is used to query the
  pixel format and is used as the device context for creating a
  PBuffer. If no device context is given, the display device context is
  used. On success, it sets the window handle. Setting a window handle
  also sets the window's device context.
\item\textsf{createWGLContext} creates an OpenGL rendering context using
  the given device context. If no device context is given, the window's
  device context is used. This function does not set the window's OpenGL
  context.
\end{enumerate}

The full \textsf{configInitWGL} task method, including error handling
and cleanup, looks as follows:

{\footnotesize\begin{lstlisting}
bool WGLWindow::configInit()
{
    if( !initWGLAffinityDC( ))
    {
        _window->setErrorMessage( "Can't create affinity dc" );
        return false;
    }

    int pixelFormat = chooseWGLPixelFormat();
    if( pixelFormat == 0 )
    {
        exitWGLAffinityDC();
        return false;
    }

    if( !configInitWGLDrawable( pixelFormat ))
    {
        exitWGLAffinityDC();
        return false;
    }

    if( !_wglDC )
    {
        exitWGLAffinityDC();
        setWGLDC( 0, WGL_DC_NONE );
        _window->setErrorMessage(
            "configInitWGLDrawable did not set a WGL drawable" );
        return false;
    }

    HGLRC context = createWGLContext();
    if( !context )
    {
        configExit();
        return false;
    }

    setWGLContext( context );
    makeCurrent();
    _initGlew();

    if( getIAttribute( Window::IATTR_HINT_SWAPSYNC ) != AUTO )
    {
        if( WGLEW_EXT_swap_control )
        {
            // set vsync on/off
            const GLint vsync = 
                ( getIAttribute( Window::IATTR_HINT_SWAPSYNC )==OFF ) ? 0 : 1;
            wglSwapIntervalEXT( vsync );
        }
        else
            EQWARN << "WGLEW_EXT_swap_control not supported, ignoring window "
                   << "swapsync hint" << std::endl;
    }

    if( !joinNVSwapBarrier( ))
    {
        _window->setErrorMessage( "Joining NV_swap_group failed" );
        return false;
    }

    if( getIAttribute( Window::IATTR_HINT_DRAWABLE ) == FBO )
        return configInitFBO();

    return true;
}
\end{lstlisting}}%>>

\subsection{\label{sTracking}Head Tracking}

The \textsf{eqPly} example contains rudimentary support for head
tracking to show how head tracking can be integrated with
Equalizer. Support for a wide range of tracking devices is not within
the scope of Equalizer. Other open source and commercial implementations
cover this functionality sufficiently and can easily be integrated with
Equalizer.

An Equalizer configuration has a vector of observers. Each observer has
its own head matrix and eye separation for tracking. In \textsf{eqPly},
all observers are updated with the same head matrix. A typical
application would assign different tracking data to different observers.

\begin{figure}[h!t]\center
  \subfigure[]{\label{fMono}
    \includegraphics[height=4cm]{images/mono}}\hfil
  \subfigure[]{\label{fStereo}
    \includegraphics[height=4cm]{images/stereo}}
  \subfigure[]{\label{fTracked}
    \includegraphics[height=4cm]{images/tracked}}\hfil
  \subfigure[]{\label{fHMD}
    \includegraphics[height=4cm]{images/hmd}}
  {\caption{\label{fImmersive}Monoscopic (a), Stereoscopic (b), Tracked (c) and
      HMD (d) Frusta}}
\end{figure}

\fig{fMono} illustrates a monoscopic view frustum. The viewer is
positioned at the origin of the global coordinate system, and the
frustum is completely symmetric. This is the typical view frustum for
non-stereoscopic applications.

In stereo rendering, the scene is rendered twice, with the two frusta
'moved' by the distance between the eyes, as shown in \fig{fStereo}.

In immersive visualization, the observer is tracked in and the view
frusta are adapted to the viewer's position and orientation, as shown
in \fig{fTracked}. The transformation $origin \rightarrow viewer$ is set by
the application using \textsf{Config::setHead\-Matrix}, which is used by
the server to compute the frusta. The resulting off-axis frusta are
positioned using the channel's head transformation, which can be
retrieved using \textsf{Channel::getHeadTransform}.

For head-mounted displays (HMD), the tracking information is used to
move the frusta with the observer, as shown in \fig{fHMD}. This results
in different projections compared to normal tracking with fixed
projection screens.

\subsection{\label{sLayoutAPI}Layout API}

The Layout API provides an abstraction for render surfaces (Canvas and
Segment) and the arrangement of rendering areas on them (Layout and
View). Its functionality has been described in \sref{sCanvas} and
\sref{sLayout}. This section focuses on how to use the Layout API
programmatically.

The application has read access to all canvases, segment, layouts and
views of the configuration. The render client has access to the current
view in the channel task methods. The layout entities can be sub-classed
using the \textsf{NodeFactory}. Currently the layout of a canvas and the
frustum of a view can be changed at runtime.

\subsubsection{Subclassing and Data Distribution}

Layout API entities (Canvas, Segment, Layout, View) are sub-classed like
all other Equalizer entities using the \textsf{NodeFactory}. Equalizer
registers the master instance of these entities on the application node,
so that the application can change mutable parameters and distribute its
own data to the render clients. \fig{fViewUML} shows the UML class
hierarchy for the \textsf{eqPly::View}.

The render clients can access a slave instance of the view using
\textsf{Channel::getView}. When called from one of the frame task
methods, this method will return the view of the current destination
channel for which the task method is executed. Otherwise it returns the
channel's native view, if it has one. Only destination channels of an
active canvas have a native view.

Equalizer commits dirty layout entities at the beginning of each
\textsf{Config::startFrame}, and synchronizes the slave instances on the
render clients correctly with the current frame.

The most common entity to subclass is the \textsf{View}, since the
application amends it by view-specific data. The view, like all other
layout entities, is indirectly sub-classed from \textsf{eq::Object}
through \textsf{eq::Frustum}. The design and usage of
\textsf{eq::Object} is explained in \sref{sEqObject}.

In \textsf{eqPly}, the application-specific data is the model
identifier. A new dirty bit for the model is defined, and the model
identifier is serialized when requested:

{\footnotesize\begin{lstlisting}
/** The changed parts of the view. */
enum DirtyBits
{
    DIRTY_MODEL       = eq::View::DIRTY_CUSTOM << 0,
};

void View::serialize( eq::net::DataOStream& os, const uint64_t dirtyBits )
{
    eq::View::serialize( os, dirtyBits );
    if( dirtyBits & DIRTY_MODEL )
        os << _modelID;
}

void View::deserialize( eq::net::DataIStream& is, const uint64_t dirtyBits )
{
    eq::View::deserialize( is, dirtyBits );
    if( dirtyBits & DIRTY_MODEL )
        is >> _modelID;
}

void View::setModelID( const uint32_t id )
{
    _modelID = id;
    setDirty( DIRTY_MODEL );
}
\end{lstlisting}}

\subsubsection{Dynamic Layout Switch}

The application can use a different layout on a canvas. This will cause
the running entities to be updated on the next frame. At a minimum, this
means the channels involved in the last layout on the canvas are
de-initialized, that is \textsf{configExit} and
\textsf{NodeFactory::releaseEntity} is called, and channels involved in
the new layout are initialized. If a layout does not cover fully a
canvas, the layout switch can also cause the (de-)initialization of
windows, pipes and nodes.

Due to the entity (de-)-initialization and the potential need to
initialize view-specific data, e.g., a model, a layout switch is
relatively expensive and will stall \textsf{eqPly} for about a second.

Initializing an entity can fail. If that happens, the server will exit
the whole configuration, and send an \textsf{EXIT} event to the
application node. The exit event will cause the application to exit,
since it resets the config's running state.

\subsubsection{View Frustum Update}

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/viewUML.pdf}
  {\caption{\label{fViewUML}UML Hierarchy of \textsf{eqPly::View}}}
\end{wrapfigure}
Changing view frustum parameters is typically done for non-fullscreen
application windows and multi-view layouts, where the rendering is not
meant to be viewed in real-world size in an immersive environment. A
typical use case is changing the field-of-view of the rendering.

A view is derived from \textsf{eq::Frustum} (\fig{fViewUML}), and the
application process can set the wall or projection parameters at
runtime. For a description of wall and projection parameters please refer
to \sref{sFrustum}. The new data will be effective for the next frame.

The frustum of a view overrides the underlying frustum of the
segments. The frustum of a view does not typically correspond to a
physical projection, and can therefore be changed at runtime. The
frustum of a segment does however correspond to a physical projection,
and is therefore immutable.

The default Equalizer event handling is using the view API to maintain
the aspect ratio of destination channels after a window resize. Without
updating the wall or projection description, the rendering would become
distorted.

\begin{wrapfigure}{l}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/viewUpdate.pdf}
  {\caption{\label{fViewUpdate}Event Flow during a View Update}}
\end{wrapfigure}
When a window is resized, a \textsf{CHANNEL\_RESIZE} event is eventually
generated. If the corresponding channel has a view,
\textsf{Channel::processEvent} sends a \textsf{VIEW\_RESIZE} event to
the application. This event contains the identifier of the view.

The config event is dispatched to \textsf{View::handleEvent}. Based on
the original size and wall or projection description of the view, a new
wall or projection is computed, keeping the aspect ratio and the height
of the frustum constant. This new frustum is automatically applied by
Equalizer at the next config frame.

\fig{fViewUpdate} shows a sequence diagram of such a view update.

\subsection{\label{sCompositing}Image Compositing for Scalable Rendering}

Two task methods are responsible for collecting and compositing the
result image during scalable rendering. Scalable rendering is a use case
of parallel rendering, when multiple channels are contributing to a single
view.

The source channels producing one or more \textsf{outputFrame}s use
\textsf{Channel::frame\-Read\-back} to read the pixel data from the frame
buffer. The channels receiving one or multiple \textsf{inputFrame}s use
\textsf{Channel::frameAssemb\-le} to assemble the pixel data into the
framebuffer. Equalizer takes care of the network transport of frame
buffer data between nodes.

Normally the programmer does not need to interfere with the image
compositing. Changes are sometimes required at a high level, for example
to order the input frames or to optimize the readback. The following
sections describe the image compositing API in Equalizer.

\subsubsection{Compression Plugins}

Compression plugins allow the creation of runtime-loadable modules for
image compression. Equalizer will search predefined directories during
\textsf{eq::init} for dynamic shared objects (DSO) containing compression
libraries (EqualizerCompressor*.dll on Windows, libeqCompressor*.dylib
on Mac OS X, libeqCompressor*.so on Linux).

The interface to a compression DSO is a C API, which allows to maintain
binary compatibility across Equalizer versions. Furthermore, the
definition of an interface facilitates the creation of new compression
codecs for developers.

Please refer to the Equalizer API documentation on the website for the
full specification for compression plugins. The Equalizer DSO doubles as
a compression plugin and implements a set of compression engines, which
can be used as a reference implementation.

Each compression DSO may contain multiple compression engines. The
number of compressors in the DSO is queried by Equalizer using
\textsf{EqCompressGetNumCompressors}.

For each compressor, \textsf{EqCompressorGetInfo} is called to retrieve
the information about the compressor. The information contains the API
version the DSO was written against, a unique name of the compressor,
the type of input data accepted as well as information about the
compressor's speed, quality and compression ratio.

Each image transported over the network allocates its own compressor or
decompressor instance. This allows compressor implementations to
maintain information in a thread-safe manner. The handle to a compressor
or decompressor instance is a void pointer, which typically hides a C++
object instantiated by the compression DSO.

A unit test is delivered with Equalizer which runs all compressors
against a set of images and provides performance information to
calculate the compressor characteristics.

\subsubsection{\label{sDirectSend}Parallel Direct Send Compositing}

To provide a motivation for the design of the image compositing
API, the direct send parallel compositing algorithm is introduced in this
section. Other parallel compositing algorithms, e.g. binary-swap, can
also be expressed through an Equalizer configuration file.

The main idea behind direct send is to parallelize the costly
recomposition for database (sort-last) decomposition. With each
additional source channel, the amount of pixel data to be composited
grows linearly. When using the simple approach of compositing all frames
on the destination channel, this channel quickly becomes the bottleneck
in the system. Direct send distributes this workload evenly across all
source channels, and thereby keeps the compositing work per channel
constant.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\label{fDirectSend}Direct Send Compositing}}
\end{wrapfigure}
In direct send compositing, each rendering channel is also responsible
for the sort-last composition of one screen-space tile. He receives
the framebuffer pixels for his tile from all the other channels. The
size of one tile decreases linearly with the number of source channels,
which keeps the total amount of pixel data per channel constant.

After performing the sort-last compositing, the color information is
transferred to the destination channel, similarly to a 2D (sort-first)
compound. The amount of pixel data for this part of the compositing
pipeline also approaches a constant value, i.e., the full frame buffer.

\fig{fDirectSend} illustrates this algorithm for three channels. The
Equalizer website contains a
presentation\footnote{\link{http://www.equalizergraphics.com/documents/EGPGV07.pdf}}
explaining and comparing this algorithm with the binary-swap algorithm.

The following operations have to be possible to perform this
algorithm:
\begin{itemize}
\item Selection of color and/or depth frame buffer attachments
\item Restricting the read-back area to a part of the rendered area
\item Positioning the pixel data correctly on the receiving channels
\end{itemize}

Furthermore it should be possible for the application to implement a
read back of only the relevant region of interest, that is, the 2D area
of the framebuffer actually updated during rendering. This optimization
will be fully supported by later versions of Equalizer.

\subsubsection{Frame, Frame Data and Images}

An \textsf{eq::Frame} references an \textsf{eq::Fra\-me\-Data}. The
frame data is the object connecting output with input frames. Output and
input frames with the same name within the same compound tree will
reference the same frame data.

The frame data is a holder for images and additional information, such
as output frame attributes and pixel data availability.

An \textsf{eq::Image} holds a
two-dimensional snapshot of the framebuffer and can contain color and/or
depth information.

The frame synchronization through the frame data allows the input frame
to wait for the pixel data to become ready, which is signaled by the
output frame after readback.

Furthermore, the frame data transports the inherited range of the output
frame's compound. The range can be used to compute the assembly order of
multiple input frames, e.g., for sorted-blend compositing in volume
rendering applications.

Readback and assemble operations on frames and images are designed to be
asynchronous. They have a start and finish method for both readback and
assemble to allow the initiation and synchronization of the operation.
Currently, only synchronous readback and assembly using
\textsf{glReadPixels} and \textsf{glDrawPixels} is implemented in the
respective start method of the image. Later versions of Equalizer will
implement asynchronous pixel transfers.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/assembly.pdf}
  {\caption{\label{fAssembly}Hierarchy of Assembly Classes}}
\end{wrapfigure}
The offset of input and output frames characterizes the position of the
frame data relative to the framebuffer, that is, the \textbf{window's}
lower-left corner. For output frames this is the position of the channel
relative to the window.

For output frames, the frame data's pixel viewport is the area of the
frame buffer to read back. It will transport the offset from the source
to the destination channel, that is, the frame data pixel viewport for
input frames position the pixel data on the destination. This has the
effect that a partial framebuffer readback will end up in the same place
in the destination channels.

The image pixel viewport signifies the region of interest that will be read
back. The default readback operation reads back one image using the full
pixel viewport of the frame data.

\fig{fAssembly} illustrates the relationship between frames, frame data
and images.

\subsubsection{\label{sCompositor}The Compositor}

The \textsf{Compositor} class gathers a set of static functions which
implement the various compositing algorithms and low-level
optimizations. \fig{fCompositor} provides a top-down functional overview
of the various compositor functions.

%\begin{wrapfigure}{r}{.618\textwidth}
\begin{figure}[h!t]\center
  \includegraphics[width=.618\textwidth]{images/compositor.pdf}
  {\caption{\label{fCompositor}Functional Diagram of the Compositor}}
\end{figure}
On a high level, the compositor combines multiple input frames using 2D
tiling, depth-compositing for polygonal data or sorted, alpha-blended
compositing for semitransparent volumetric data. These operations
composite either directly all images on the GPU, or use a CPU-based
compositor and then transfer the pre-integrated result to the GPU. The
high-level entry points automatically select the best algorithm. The
CPU-based compositor uses, based on availability, OpenMP and the
Paracomp compositing library to accelerate its operation.

On the next lower level, the compositor provides functionality to
composite a single frame, either using 2D tiling (possibly with blending
for alpha-blended compositing) or depth-based compositing.

The per-frame compositing in turn relies on the per-image compositing
functionality, which automatically decides on the algorithm to be used
(2D or depth-based). The concrete per-image assembly operation uses
OpenGL operations to composite the pixel data into the framebuffer,
potentially using GLSL for better performance.

\subsubsection{Custom Assembly in eVolve}

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\label{fSlices}Volume Rendering using 3D Textures}}\vspace{-2ex}
\end{wrapfigure}
The \textsf{eVolve} example is a scalable volume renderer. It uses 3D
texture-based volume rendering, where the 3D texture is intersected by
view-aligned slices. The slices are rendered back-to-front and blended
to produce the final image, as shown in \fig{fSlices}\footnote{Volume
  Data Set courtesy of: SFB-382 of the German Research Council (DFG)}.

When using 2D (sort-first) or stereo decompositions, no special
programming is needed to achieve good scalability, as \textsf{eVolve} is
mostly fill-limited and therefore scales nicely in these modes.

The full power of scalable volume rendering is however in DB (sort-last)
compounds, where the full volume is divided into separate bricks. Each
of the bricks is rendered like a separate volume. For recomposition, the
\textsf{RGBA} frame buffer data resulting from these render passes then
has to be assembled correctly.

Conceptually, the individual volume bricks of each of the source
channels produces pixel data which can be handled like one big 'slice'
through the full texture. Therefore they have to be blen\-ded
back-to-front in the same way as the slice planes are blended during
rendering.

DB compounds have the advantage of scaling any part of the volume
rendering pipeline: texture and main memory (smaller bricks for each
channel), fill rate (less samples per channel) and IO bandwidth for
time-dependent data (less data per time step and channel). Since the
amount of texture memory needed for each node decreases linearly, they
make it possible to render data sets which are not feasible to
visualize with any other approach.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/volResult.png}
  {\caption{\label{fVolResult}Result of \fig{fBlendPerspective}}}
\end{wrapfigure}
For recomposition, the 2D frame buffer contents are blended to form a
seamless picture. For correct blending, the frames are ordered in the
same back-to-front order as the slices used for rendering, and use the
same blending parameters. Simplified, the frame buffer images are
`thick' slices which are `rendered' by writing their content to the
destination frame buffer using the correct order.

For orthographic rendering, determining the compositing order of the
input frames is trivial. The screen-space orientation of the volume
bricks determines the order in which they have to be composited. The
bricks in \textsf{eVolve} are created by slicing the volume along one
dimension. Therefore the range of the resulting frame buffer images,
together with the sorting order, is used to arrange the frames during
compositing. \fig{fBlendOrtho} shows this composition for one view.

Finding the correct assembly order for perspective frusta is more
complex. The perspective distortion invalidates a simple orientation
criteria like the one used for orthographic frusta. For the view and
frustum setup shown in \fig{fBlendPerspective}\footnote{Volume Data Set
  courtesy of: AVS, USA} the correct compositing order is 4-3-1-2 or
1-4-3-2.

\begin{figure}[h!t]
  \subfigure[]{\label{fBlendOrtho}
    \includegraphics[width=.45\textwidth]{images/b2f_ortho.pdf}
  }\hfil
  \subfigure[]{\label{fBlendPerspective}
    \includegraphics[width=.55\textwidth]{images/b2f_perspective.pdf}
  }%
  {\caption{\label{fBlend}Back-to-Front Com\-po\-siting for
      Orthogonal and Perspective Frusta}}
\end{figure}

To compute the assembly order, \textsf{eVolve} uses the angle between
the $origin \rightarrow slice$ vector and the near plane, as shown in
\fig{fBlendPerspective}. When the angle becomes greater than
90\textdegree, the compositing order of the remaining frames has to be
changed. The result image of this composition naturally looks the same
as the volume rendering would when rendered on a single
channel. \fig{fVolResult} shows the result of the composition from
\fig{fBlendPerspective}.

The assembly algorithm described in this section also works with parallel
compositing algorithms such as direct-send.

\subsection{\label{sFSAA}Custom Subpixel Processing}

The eqPly example contains support for software anti-aliasing when being idle,
showing how to combine it with subpixel compounds.

\subsubsection{Supersampling Algorithm}

The supersampling algorith used by eqPly performs anti-aliasing, using a
technique which computes samples within a pixel by jittering the frustum by a
subpixel amount. The pixel is split into several, regular sub-pixels and a
sample is taken randomly within the sub-pixel.

The subpixel area is a function of the current jitter step and the channel's
\textsf{subpixel} description. To evenly distribute jitter values for all
channels over the total sub-pixel grid a precomputed prime number table is used
to pseudo-randomly select the subpixel:

{\footnotesize\begin{lstlisting}
eq::Vector2i Channel::_getJitterStep() const
{
    uint32_t channelID  = getSubPixel().index;
    uint32_t idx = ( channelID * _totalSteps );
    idx += ( _jitterStep * primeNumberTable[ channelID ] ) % _totalSteps;

    uint32_t sampleSize = _getSampleSize();
    const int dx = idx % sampleSize;
    const int dy = idx / sampleSize;

    return eq::Vector2i( dx, dy );
}
\end{lstlisting}}

Using this normalized subpixel center point and the size of a pixel on the near
plane, the jitter for the frustum is computed. A random position within the
sub-pixel is set as a sample position, which will be used to move the frustum:

{\footnotesize\begin{lstlisting}
eq::Vector2f Channel::_getJitter() const
{
    const eq::PixelViewport& pvp = getPixelViewport();
    float pvp_w = static_cast<float>( pvp.w );
    float pvp_h = static_cast<float>( pvp.h );
    float frustum_w = static_cast<float>(( getFrustum().get_width( )));
    float frustum_h = static_cast<float>(( getFrustum().get_height( )));

    float pixel_w = frustum_w / pvp_w;
    float pixel_h = frustum_h / pvp_h;

    uint32_t sampleSize = _getSampleSize();
    float subpixel_w = pixel_w / static_cast<float>( sampleSize );
    float subpixel_h = pixel_h / static_cast<float>( sampleSize );

    eq::Vector2i jitterStep = _getJitterStep();

    // Sample value randomly computed within the subpixel
    eq::base::RNG rng;
    float value_i = rng.get< float >() * subpixel_w
                    + static_cast<float>( jitterStep.x( )) * subpixel_w;

    float value_j = rng.get< float >() * subpixel_h
                    + static_cast<float>( jitterStep.y( )) * subpixel_h;

    return eq::Vector2f( value_i, value_j );
}
\end{lstlisting}}

\subsubsection{Idle Anti-Aliasing in eqPly}

The \textsf{FrameData} class holds the information whether or not the
application is in idle mode. The \textsf{Config} updates the idle mode
information depending on the application's state.  Each \textsf{Channel} will
perform anti-aliasing when no user event requires a redraw.

If the application is in idle mode, the \textsf{applyFrustum} operation will 
apply the computed jitter vector to the current frustum.

{\footnotesize\begin{lstlisting}
void Channel::applyFrustum() const
{
    const FrameData& frameData = _getFrameData();

    if( frameData.isIdle() && _jitterStep > 0 )
    {
        eq::Vector2f jitter = _getJitter();

        if( frameData.useOrtho( ))
        {
            eq::Frustumf ortho = getOrtho();
            ortho.apply_jitter( jitter );

            glOrtho( ortho.left(), ortho.right(),
                     ortho.bottom(), ortho.top(),
                     ortho.near_plane(), ortho.far_plane( ));
        }
        else
        {
            eq::Frustumf frustum = getFrustum();
            frustum.apply_jitter( jitter );

            glFrustum( frustum.left(), frustum.right(),
                       frustum.bottom(), frustum.top(),
                       frustum.near_plane(), frustum.far_plane( ));
        }
        return;
    }
    //else

    eq::Channel::applyFrustum();
}
\end{lstlisting}}

At the end of the frame, \textsf{frameViewFinish} accumulates the frame buffer
into an accumulation buffer. Since \textsf{frameViewFinish} is only called on
destination channels this operation is done only on the final rendering
result. After each frame, the result accumulated so far will be displayed. When
all the steps are done the config will stop rendering new frames.

If the current pixel viewport is different from the one saved in
\textsf{frameViewStart}, the accumulation buffer needs also to be resized and
the idle anti-aliasing is reset.

{\footnotesize\begin{lstlisting}
void Channel::frameViewFinish( const uint32_t frameID )
{
    const FrameData& frameData = _getFrameData();

    const eq::PixelViewport& pvp = getPixelViewport();
    bool isResized = _accum->resize( pvp.w, pvp.h );

    if( isResized )
    {
        _accum->clear();
        _jitterStep = _totalSteps;
        _subpixelStep = 0;
        return;
    }
    //else

    if( frameData.isIdle() && _needsTransfer )
    {
        setupAssemblyState();

        _accum->accum();
        _accum->display();

        resetAssemblyState();
    }

    _drawOverlay();
    _drawHelp();

    if( frameData.useStatistics())
        EQ_GL_CALL( drawStatistics( ));
}
\end{lstlisting}}

At the end of the frame for each channel, \textsf{frameFinish} will decrement
the jitter step and create a new event to inform the \textsf{Config} of how many
frames remain to draw before the idle anti-aliasing is complete.

\subsubsection{Transparent Software Anti-Aliasing}

Any Equalizer applications can benefit without modification from subpixel
compounds. Applications performing their own, configurable idle or non-idle
anti-aliasing can easily be integrated with subpixel compounds, as described in
the next section.

The default implementation of \textsf{applyFrustum} and \textsf{applyOrtho} in
\textsf{eq::Channel} apply jittering when a subpixel parameter is set.
Furthermore, the \textsf{Compositor} uses an \textsf{eq::Accum} buffer to
accumulate and display the results of a subpixel decomposition in \textsf{assembleFramesUnsorted} and \textsf{assembleFramesSorted}.

\subsubsection{Subpixel Compounds in eqPly}

Subpixel compounds accelerate the idle anti-aliasing computation in eqPly. The
accumulation buffer used in the \textsf{Compositor} in \textsf{frameAssemble} is
the same accumulation buffer used by eqPly in \textsf{frameViewFinish} for
collecting the idle anti-aliasing results.

When an application is passing an \textsf{eq::Accum} object to the compositor,
the compositor adds subpixel decomposition result into this accumulation
buffer. Otherwise it uses its own accumulation buffer, which it clears in the
beginning of the compositing process.

Since \textsf{eqPly} accumulates the results over multiple rendering frames,
each \textsf{Channel} manages its own accumulation buffer, which is passed to
the compositor in \textsf{frameAssemble} and used in \textsf{frameViewFinish} to
display the results accumulated so far.

The number of performed jitter steps is tracked in \textsf{frameDraw} and
\textsf{frameAssemble}, and is used in \textsf{frameFinish} to decrement the
outstanding anti-aliasing samples. Therefore, the idle anti-aliasing is finished
much faster if the current configuration uses a subpixel compound.

\subsection{\label{sStatistics}Statistics}

\subsubsection{Statistics Gathering}

Statistics are measured in milliseconds since the configuration was
initialized. On each node, the global configuration clock is synchronized
with the server clock. Each statistic event records the originator's
(channel, window, frame or config) unique identifier.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/statistics}
  {\caption{\label{fStatistics}Statistics for a two-Node 2D Compound}}
\end{wrapfigure}
Statistics are enabled per entity using an attribute hint. The hint
determines how precise the gathered statistics are. When set to
\textsf{fastest}, the per-frame clock is sampled directly when the event
occurs. When set to \textsf{nicest}, all OpenGL commands will be
finished before sampling the event. This incurs a performance penalty,
but gives more correct results. The default setting is fastest in
release builds, and nicest in debug builds.

The events are processed by the channel's and window's
\textsf{processEvent} method. The default implementation sends these
events to the config using \textsf{Config::sendEvent}, as explained in
\sref{sEventHandling}. When the default implementation of
\textsf{Config::handleEvent} receives the statistics event, it sorts the
event per frame and per originator. When a frame has been finished, the
events are pushed to the local (app-)node for visualization.

\fig{fStatistics} shows the visualization of statistics events in an
overlay\footnote{3D model courtesy Cyberware, Inc.}.

\subsubsection{Statistics Overlay}

The Equalizer examples render a statistics overlay using the gathered
statistics events directly before
\textsf{Window::swapBuffers}. Statistics are toggled on and off using
the 's' key. The overlay rendering method,
\textsf{Channel::drawStatistics}, can be called directly by any
application to render the overlay. It can also be overwritten by a
custom implementation.

\fig{fStatisticsDetail} shows a detail view of \fig{fStatistics}. The
statistics shown are for a two-node 2D compound. The destination channel
is on the \textsf{appNode} and contributes to the rendering.

\begin{figure}[h!t]
  \includegraphics[width=\textwidth]{images/statisticsDetail}
  {\caption{\label{fStatisticsDetail}Detail of the Statistics from
      \fig{fStatistics}.}}
\end{figure}

This configuration executes two \textsf{Channel::frameDraw} tasks, one
\textsf{Channel::frame\-Read\-back} task on the remote node, one
\textsf{Channel::frameAssemble} task on the local node, as well as frame
transmission and compression.

The X axis is the time. One pixel on the screen corresponds to a given
time unit, here one millisecond per pixel. The scale is zoomed dynamically
to powers-of-ten milliseconds to fit the statistics into the available
viewport. This allows easy and accurate evaluations of bottlenecks or
misconfigurations in the rendering pipeline. The scale of the statistics
is printed.

The rightmost pixel is the most current time. On the Y axis are the
entities: channels, windows, frames and the config. The upper channel is
the local channel since it executes \textsf{frameAssemble}, and the
lower channel is the remote channel, executing \textsf{frameReadback}.

The draw time has been artificially inflated by using immediate mode
rendering to provide more expressive statistics.

To facilitate the understanding, older frames are gradually grayed
out. The rightmost, current frame is brighter than the
frame-before-last.

The configuration used has a latency of one frame. Consequently, the
execution of two frames overlaps. This can be observed in the early
execution of the remote channel's \textsf{frameDraw}, which starts while
the local channel is still assembling the previous frame.

The beginning of a frame is marked by a vertical green line, and the end
of a frame by a vertical gray line. These lines are also attenuated. The
brightness and color matches the event for \textsf{Config::startFrame}
and \textsf{Config::finishFrame}, respectively. The event for
\textsf{startFrame} is typically not visible, since it takes less than
one millisecond to execute. If no idle processing is done by the
application, the event for \textsf{finishFrame} occupies a full frame,
since the config is blocked here waiting for the frame to complete.

In the above example, the local channel finishes drawing the frame
early, and therefore spends a considerable amount of time waiting for
the input frame from the remote channel. These wait events, rendered
red, are a sub-event of the yellow \textsf{frameAssemble} task.

The frame statistics provide information on the image transmission for
scalable rendering. They are grouped per frame, not per node, that is,
the transmit and compress part are executed by one node and the receive
by another, even though the statistics are drawn in the same line. The
decompression time is part of the receive statistic. The total
compression ratio is printed with each compression statistic, in this
case the image has been compressed to 11\% of its original size.

The white \textsf{Window::swapBuffers} task might also take also a long
time, since the execution of the swap buffer is locked to the vertical
retrace of the display, and enforced by \textsf{glFinish} before
sampling the end time of the event. In release builds, this does not
happen and all OpenGL commands are pipelined with the execution on the
CPU.

Note that the remote source window does not execute the swapBuffers in
this configuration, since it is a single-buffered PBuffer.

\subsection{\label{sCUDA}GPU Computing with CUDA}

The Equalizer parallel rendering framework can be used to scale GPU 
computing applications (also known as GPGPU) based on CUDA. The necessary
steps to write a distributed GPGPU application using Equalizer are described
in the following.

Please note that the support for CUDA is only enabled if the CUDA libraries are found
in their default locations. Otherwise please adapt the makefiles accordingly.

\subsubsection{CUDA Initialization}

Equalizer automatically initializes a CUDA context on each pipe (using the specified
device number) whenever \textsf{hint\_cuda\_GL\_interop} is set. The initialization is 
currently done using the CUDA runtime API in order to support device emulation. 
On the other side, if this hint is not set the OpenGL interop feature is not enabled, and 
furthermore, the GPU which executes CUDA code defaults to device 0.

Once CUDA is initialized correctly, C for CUDA can be used as any other GPU 
programming interface.

\subsubsection{CUDA Memory Management}

Equalizer does not provide any facility to perform memory transfer from and to the CUDA
devices. This is entirely left to the programmer.

\section{\label{sNetwork}The Network Layer}

The Equalizer networking layer provides a peer-to-peer communication
infrastructure. It is used by Equalizer to communicate between the
application node, the server and the render clients. It can also be
used by applications to implement distributed processing independently
or complementary to the core Equalizer functionality.

The network layer is implemented in the \textsf{eq::net} namespace. It
provides layered functionality, which provide higher level functionality
for the programmer. The main primitives in the network layer are:

\begin{description}
\item[Connection] A stream-oriented point-to-point communication
  line. Different implementations of a connection exist.
\item[Node] The abstraction of a process in the cluster. Nodes
  communicate with each other using connections.
\item[Session] Provides unique identifiers and object mapping to a set
  of nodes.
\item[Object] Provides object-oriented data distribution of C++ objects
  between nodes within a session.
\end{description}

\subsection{Connections}

The \textsf{eq::net::Connection} is the basic primitive used for
communication between processes in Equalizer. It provides a
stream-oriented communication between two endpoints.

A connection is either closed, connected or listening. A closed
connection cannot be used for communications. A connected connection
can be used to read or write data to the communication peer. A listening
connection can accept connection requests.

An \textsf{eq::net::ConnectionSet} is used to manage multiple
connections. The typical use case is to have one or more listening
connections for the local process, and a number of connected connections
for communicating with other processes.

The connection set is used to select one connection which requires some
action. This can be a connection request on a listening connection,
pending data on a connected connection or the notification of a
disconnect.

The connection and connection set can be used by applications to
implement other network-related functionality, e.g., to communicate
with a sound server on a different machine.

\subsection{Packet Handling}

Nodes, sessions and objects communicate using packets derived from
\textsf{eq::net::Packet}. The basic packet dispatch and handling is
implemented in the \textsf{eq::net::Base} class, from which
\textsf{eq::net::Node}, \textsf{eq::net::Session} and
\textsf{eq::net::Object} is sub-classed.

The network base class allows the registration of a packets with a
dispatch queue and an invocation method. Each packet has a command
identifier, which is used to identify the registered queue and
method. The method \textsf{dispatchCommand} pushes the packet to the
registered queue, whereas the method \textsf{invokeCommand} calls the
registered command function.

A command function groups the method and \textsf{this} pointer, allowing
to call a C++ method on a concrete instance. If no queue is registered
for a certain command, \textsf{dispatchCommand} directly calls the
registered command function.

This dispatch and invocation functionality is used within Equalizer to
dispatch commands from the receiver thread to the appropriate node or
pipe thread, and then to invoke the command when it is processed by
these threads.

\subsection{Nodes}

The \textsf{eq::net::Node} is the abstraction of one process in the
cluster. Each node has a unique identifier, based on the UUID
standard\footnote{\link{http://en.wikipedia.org/wiki/Universally\_Unique\_Identifier}}. This
identifier is used to address nodes, e.g., to query connection
information to connect to the node. Nodes use connections to communicate
with each other by sending packets sub-classed from
\textsf{eq::net::NodePacket}.

Nodes, like connections, are either in the closed, listening or
connected state. A listening node uses one or more listening connections
in conjunction with a connection set to wait for data or connection
requests. For each connection request, a connection and node are created
to represent this node on the local machine. This node is in the
connected state, and its connection is added to the connection set of
the listening node.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/netNode.pdf}
  {\caption{\label{fNetNode}Communication between two Nodes}}
\end{wrapfigure}
\fig{fNetNode} shows the communication between two nodes. The nodes in
the listening state also have a listening connection, which is not shown
in the Figure for complexity reasons.

When the remote node sends a packet, the listening node receives the
packet and dispatches it from the receiver thread using the method
\textsf{dispatchCommand}. The default implementation knows how to
dispatch packets of type node, session or object. Applications can
define custom data types for packets, and then have to extend
\textsf{dispatchCommand} to handle these custom data types.

Node packets are directly dispatched using
\textsf{Base::dispatchCommand}. For session and object packets, the
appropriate session is found and \textsf{Session::dispatchCommand} will
dispatch the packet.

If \textsf{dispatchCommand} returns false, the packet will be
re-dispatched later. This is used if an object has not been mapped
locally, and therefore the packet could not be dispatched.

If an application wants to extend communication on the node level, it
can either define its own datatype for packets, or to define custom node
packets. Packets with a custom datatype are handled by overriding
\textsf{Node::dispatchCommand}. Custom node packets can use any command
greater than \textsf{CMD\_NODE\_CUSTOM} as the packet command. By
registering a \textsf{eq::net::CommandFunc} for these commands, the
\textsf{eq::net::Base} dispatch and invoke mechanism can be used
automatically.

\subsection{Sessions}

The \textsf{eq::net::Session} gathers provides higher-level
functionality to a set of nodes. The master instance of a session is
registered to a node, which makes this node the session server and
assigns a node-unique identifier to the session. All other nodes map
their instance of the session to their local listening node using the
session's identifier. One given node can participate in any number of
sessions, both as master and slave. This allows the flexible usage of
sessions to group different aspects of an application.

A session provides the facilitate to allocate session-unique
identifiers. These identifiers can be used by the application to
uniquely address data within the cluster. They are also used by the
session to address distributed objects, which are described in the
following section.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/netSession.pdf}
  {\caption{\label{fNetSession}Communication between two Sessions}}
\end{wrapfigure}
Sessions instances can communicate to each other directly using an
\textsf{eq::net::Ses\-sion\-Packet}, which is derived from the
\textsf{NodePacket} and extends it by the session identifier. The
session ID is used by the node to dispatch the packet to the session.

The session, similar to the node, uses the methods
\textsf{dispatchCommand} and \textsf{invokeCommand} to dispatch and
execute packets using either the \textsf{eq::net::Base} class for
session packets or to dispatch and invoke object packets on the
appropriate \textsf{eq::net::Object}. Application developers can use
this mechanism by creating custom session packets with a command greater
than \textsf{CMD\_SESSION\_CUSTOM} and registering a command function
for these commands.

\subsection{\label{sNetObject2}Objects}

Distributed objects provide powerful, object-oriented data distribution for C++
objects. They facilitate the implementation of data distribution in a cluster
environment. Their functionality and an example use case for parallel rendering
has been described in \sref{sNetObject}.

Distributed objects are created by subclassing from \textsf{eq::Object} or
\textsf{eq::net::Object}. The application programmer implements serialization
and deserialization of the distributed data. Objects are dynamically assigned to
a session, which assigns session-unique identifiers to objects and manages the
network communication between different instances of the same distributed
object.

The master version of a distributed object is registered with a
\textsf{eq::net::Session}, which assigns a session-unique identifier to
the object. This identifier can be used by other nodes to map their
instance of the object, thus synchronizing the object's data with the
remotely registered master version.

Distributed objects can be static (immutable) or dynamic. Dynamic objects are
versioned. New versions are commited from the master instance, which sends
the delta between the previous and current version to all mapped slave
objects. The slave objects sync the queued deltas when they need a version.

\subsubsection{Common Usage for Parallel Rendering}

Distributed objects are addressed using session-unique identifiers,
because pointers to other objects cannot be distributed directly; they
have no meaning on remote nodes. The session used to register
distributed objects is normally the \textsf{eq::Config}, which is
sub-classed from \textsf{eq::net::Session}.

The entry point for shared data on a render client is the identifier
passed by the application to \textsf{Config::init}. This identifier
typically contains the identifier of a static distributed object, and is
passed by Equalizer to all \textsf{configInit} task methods. Normally
this initial object is mapped by the render clients in
\textsf{Node::configInit}. It normally contains identifiers of other
shared data objects.

The distributed data objects referenced by the initial data object are
often versioned objects, to keep them in sync with the rendered
frames. Similar to the initial identifier passed to
\textsf{Config::init}, an object identifier or object version can be
passed to \textsf{Config::startFrame}. Equalizer will pass this
identifier to all \textsf{frameStart} task methods. In \textsf{eqPly},
the frame-specific data, e.g., the global camera data, is versioned. The
frame data identifier is passed in the initial data, and the frame data
version is passed with each new frame request.

\begin{wrapfigure}{r}{.382\textwidth}
%\begin{figure}[h!t]
  \subfigure[]{\label{fObjectDist1}
    \includegraphics[width=.382\textwidth]{images/umlObjectSubclass}}\hfil
  \subfigure[]{\label{fObjectDist2}
    \includegraphics[width=.382\textwidth]{images/umlObjectProxy}}\hfil
  \subfigure[]{\label{fObjectDist3}
    \includegraphics[width=.382\textwidth]{images/umlObjectMI}}%
  {\caption{\label{fObjectDist}Object Distribution using Subclassing,
      Proxies or Multiple Inheritance}}
\end{wrapfigure}
There are multiple ways of implementing data distribution for an
existing C++ class hierarchy, such as a scene graph:
\begin{description}
\item[Subclassing] The classes to be distributed inherit from
  \textsf{eq::net::Object} and implement the serialization methods. This
  approach is recommended if the source code of existing classes can be
  modified. It is used for \textsf{eqPly::InitData} and
  \textsf{eqPly::FrameData}. (\fig{fObjectDist1})
\item[Proxies] For each object to be distributed, a proxy object is
  created which manages data distribution for its associated
  object. This requires the application to track changes on the object
  separately from the object itself. The model data distribution of
  eqPly is using this pattern. (\fig{fObjectDist2})
\item[Multiple Inheritance] A new class inheriting from the class to be
  distributed and from \textsf{eq::net::Object} implements the data
  distribution. This requires the application to instantiate a different
  type of object instead of the existing object, and to create wrapper
  methods in the superclass calling the original method and setting the
  appropriate dirty flags. This pattern is not used in
  eqPly. (\fig{fObjectDist3})
\end{description}

\subsubsection{Change Handling and Serialization}

Equalizer determines the way changes are to be handled by calling
\textsf{Object::get\-Change\-Type} during the registration of the master version
of a distributed object. The change type determines the memory footprint and the
contract for the serialization methods. The deserialization always happens in
the application thread causing the deserialization. The following change types
are possible:

\begin{description}
  \item[STATIC] The object is not versioned. The instance data is serialized
    whenever a new slave instance is mapped. The serialization happens in the
    command thread on the master node, i.e., in parallel to the application
    threads. Since the object's distributed data is supposed to be static, this
    should not create any race conditions. No additional data is stored.
  \item[INSTANCE] The object is versioned, and the instance and delta data is
    identical, that is, only instance data is serialized. The serialization
    always happens before \textsf{Session::registerObject} or
    \textsf{Object::commit} returns. Previous instance data is saved to be able
    to map old versions.
  \item[DELTA] The object is versioned, and the delta data is typically smaller
    than the instance data. Both the delta and instance data are serialized
    before \textsf{Object::commit} returns. Previous instance data is saved to
    be able to map old versions.
  \item[UNBUFFERED] The object is versioned. No data is stored, and no previous
    versions can be mapped. The instance data is serialized from the command
    thread whenever a new slave instance is mapped, i.e., in parallel to
    application threads. The application has to ensure that this does not create
    any thread conflicts. The delta data is serialized before
    \textsf{Object::commit} returns. The application may choose to use a
    different, more optimal implementation to pack deltas by using a different
    implementation for \textsf{getInstanceData} and \textsf{pack}.
\end{description}

\subsubsection{\label{sSlaveCommit}Slave Object Commit}

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-2ex}\includegraphics[width=.618\textwidth]{images/slaveWrites.pdf}
  {\caption{\label{fSlaveCommit}Slave Commit Communication Sequence}}
\end{wrapfigure}
Versioned slave objects may also commit data to the master instance. However,
this requires the implementation to take special care to resolve conflicts
during deserialization. The algorithm and constraints for slave object commits
are explained in this section.

When using \textsf{commit} on slave object instances, multiple modifications on
the same database may happen simultaneously. In order to avoid excessive
synchronization in the network layer, Equalizer a light-weight approach which
requires the application to avoid or handle conflicting changes on the same
object.

When a slave object instance commits data, \textsf{pack} is called on the
object. The data serialized by \textsf{pack} is send to the master instance of
the object. On the master instance, \textsf{notifyNewVersion} is called from the
command thread to signal the data availability. No new version is generated on
the slave instance.

The master instance can \textsf{sync} delta data received from slaves. The data
is synchronized in the order it has been received from the slaves. If a specific
order is needed, it is the application's responsibity to serialize the slave
commits accordingly. When using \textsf{VERSION\_NEXT} as the version parameter
to \textsf{sync}, one slave commit is unpacked. This call may block if no data
is pending. When using \textsf{VERSION\_HEAD} all pending versions are
unpacked. This call will never block.

Synchronizing slave data on a master instance does not create a new version. The
application has to explicitly call \textsf{commit} to pack the changed data,
which will create a new version send to all slaves.

This approach requires the application to handle or avoid, if possible, the
following situations:
\begin{itemize}
\item A slave instance will always receive its own changes when the master
  commits the slave change.
\item A slave instance commits data which does not correspond to the last
  version of the master instance.
\item Two different slave instances commit changes simultaneously which may
  result in conflicts.
\item A master instance applies data to an already changed database.
\item The master instance commits a new version while a slave instance is
  modifying its instance. 
\end{itemize}

\fig{fSlaveCommit} depicts one possible communication sequence. In general, it
is best to adhere to the following rules:
\begin{itemize}
\item Handle the case that a slave instance will receive its own changes again.
\item Write your application code so that only one object instance is modified
  at any given time. 
\item When multiple slave instances have to modify the same object at the same
  time, try to modify and serialize disjunct data of the object.
\item Before modifying a slave instance, \textsf{sync} it to the head version.
\item Write serialization code defensively, e.g.:
  \begin{itemize}
  \item When adding a new child to an object, check that it has not been added
    already.
    \item When removing a child, transmit the child's identifier instead of its
      index for removal.
    \item Modify a variable by serializing its value (instead of an operation on
      it) so that the last change will win.
  \end{itemize}
\end{itemize}

\subsubsection{Communication between Object Instances}

Objects can send command packets to each other by subclassing
\textsf{eq::net::Ob\-ject\-Packet}, which is derived from
\textsf{SessionPacket}. The object packet extends the session packet by
the object identifier and an object instance identifier.

Multiple instances of the same object (identifier) can be mapped on the
same node and session simultaneously. Each object instance has a
node-unique instance identifier. When the instance ID is set to
\textsf{EQ\_ID\_ANY} (the default value), a command is dispatch to all
object instances registered to the session with the correct object
ID. Multiple object instance per node are used when different threads on
the node need to process a different version of the same object.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/netObject.pdf}
  {\caption{\label{fNetObject}Communication between two Objects}}
\end{wrapfigure}
Distributed objects provide data synchronization functionality, as described
above. Additionally, applications can send command packets by creating custom
object packets with a command greater than \textsf{CMD\_OBJECT\_CUSTOM}.

Distributed object communication relies on the lower node and session
communication layers, as shown in \fig{fNetObject}.

\subsection{Barrier}

The \textsf{eq::net::Barrier} provides a networked barrier primitive. It
is an \textsf{eq::net::Object} used by Equalizer for software swap
barrier support, but it can be used as a generic barrier in application
code.

The barrier uses both the data distribution for synchronizing its data,
as well as custom command packets to implement the barrier logics. A
barrier is a versioned object. Each version can have a different height,
and enter request are automatically grouped together by version.

\subsection{\label{sNetUsage}Usage in Equalizer}

The Equalizer client library and server are built on top of the network
layer. They influence the design of it, and serve as a sample
implementation on how to use classes in the \textsf{eq::net} namespace.

\subsubsection{\label{sEqObject}eq::Object}

Equalizer uses the \textsf{eq::net::Object} data distribution mechanism
for implementing data distribution between the server, application and
render clients. To do this, \textsf{eq::Object} implements one usage
pattern for the \textsf{eq::net::Object}, allowing inheritance with data
distribution. It is easier to use, but imposes one typical way to implement data
distribution.

The \textsf{eq::Object} data distribution is based on the concept of
dirty bits. Dirty bits form a 64-bit mask which marks which parts of the
object have to be distributed during the next commit.

For serialization, the default \textsf{eq::net::Object} serialization
functions are implemented by \textsf{eq::Object}, which (de-)serializes
and resets the dirty bits, and calls \textsf{serialize} or
\textsf{deserialize} with the bit mask specifying which data has to be
transmitted or received. During a commit or sync, the current dirty bits
are given, whereas during object mapping all dirty bits are passed to
the serialization methods.

To use \textsf{eq::Object}, the following steps have to be taken:

\begin{description}
\item[Inherit from \textsf{eq::Object}:] The base class will provide the
  dirty bit management and call serialize and deserialize
  appropriately. By overriding \textsf{getChangeType}, the default
  versioning strategy might be changed.
\item[Define new dirty bits:] Define dirty bits for each data item by
  starting at \textsf{Object::DIRTY\_CUSTOM}, shifting this value
  consecutively by each new dirty bit.
\item[Implement serialize and deserialize:] For each object-specific
  dirty bit which is set, stream the corresponding data item to or from
  the provided stream. Call the parent method first in both
  functions. For application-specific objects, write a (de-)serialization
  function.
\item[Mark dirty data:] In each 'setter' method, call \textsf{setDirty}
  with the corresponding dirty bit.
\end{description}

The inheritance chain \textsf{eq::View} $\rightarrow$
\textsf{eq::Frustum} $\rightarrow$ \textsf{eq::Object} $\rightarrow$
\textsf{eq::net::Object} serves as sample implementation for this
process. The \textsf{eq::Object} base class provides the facility to set
a name on the object, which is automatically distributed.

The registration and mapping of \textsf{eq::Object}s is done in the same
way as for \textsf{eq::net::Object}s, which has been described in
\sref{sNetObject2}. Equalizer objects derived from \textsf{eq::Object},
e.g., \textsf{eq::View}, are typically registered and mapped by
Equalizer.

\newpage
\appendix
\section{\label{aFileFormat}File Format}

The current Equalizer file format is a one-to-one representation of the
server's internal data structures. Its purpose is intermediate, that is,
it will gradually be replaced by automatic resource detection and
automatic configuration. However the core scalability engine will always
use a similar structure as currently exposed by the file format.

The file format represents an ASCII deserialization of the
server. Streaming an \textsf{eq::server::Server} to an \textsf{eq::base::Log}
ostream produces a valid configuration file. Likewise, loading a
configuration file produces an \textsf{eq::server::Server}.

The file format uses the same syntactical structure as VRML. If your
editor supports syntax highlighting and formatting for VRML, you can use
this mode for editing \textsf{.eqc} files.

The configuration file consist of an optional global section and a
server configuration. The global section defines default values for
various attributes. The server section represents an
\textsf{eq::server::Server}.

\subsection{File Format Version}

\subsection{\label{sGlobal}Global Section}

The global section defines default values for attributes used by the
individual entities in the server section. The naming convention for
attributes is:

{\footnotesize\begin{lstlisting}
EQ_<ENTITY>_<DATATYPE>ATTR_<ATTR_NAME>
\end{lstlisting}}

The entity is the capitalized name of the corresponding section later in
the configuration file: connection, config, pipe, window, channel or
compound. The connection is used by the server and nodes.

The datatype is one capital letter for the type of the attribute's
value: \textsf{S} for strings, \textsf{C} for a character, \textsf{I}
for an integer and \textsf{F} for floating-point values. Enumeration
values are handled as integers. Strings have always to be surrounded by
double quotes '\textsf{''}'. A character has to be surrounded by single
quotes '\textsf{'}'.

The attribute name is the capitalized name of the entities attribute, as
discussed in the following sections.

Global attribute values have useful default parameters, which can be
overridden with an environment variable of the same name. For enumeration
values the corresponding integer value has to be used. The global values
in the config file override environment variables, and are in turn
overridden by the corresponding attributes sections of the specific
entities.

The globals section starts with the token \textsf{global} and an open
curly brace '\textsf{\{}', and is terminated with a closing curly brace
'\textsf{\}}'. Within the braces, globals are set using the attribute's
name followed by its value. The following attributes are available:

\begin{center}
\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_SATTR\_HOSTNAME\\
  \textbf{Value} & string\\
  \textbf{Default} & ''localhost''\\
  \textbf{Details} & The hostname or IP address used to connect the
  server or node. When used for the server, the listening port of the
  server is bound to this address. When used for a node, the server
  first tries to connect to the render client node using this hostname,
  and then tries to launch the render client executable on this host.\\
  \textbf{See also} & EQ\_NODE\_SATTR\_LAUNCH\_COMMAND
  EQ\_CONNECTION\_IATTR\_PORT
  EQ\_CONNECTION\_IATTR\_TYPE\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_IATTR\_TYPE\\
  \textbf{Value} & TCPIP \textbar \ SDP \textbar \ RSP \textbar \ 
  PIPE [Win32 only]\\
  \textbf{Default} & TCPIP\\
  \textbf{Details} & The protocol for connections. SDP programmatically
  selects the socket direct protocol (AF\_INET\_SDP) provided by most
  InfiniBand protocol stacks, TCPIP uses normal TCP
  sockets (AF\_INET). RSP provides reliable UDP multicast. PIPE uses a named
  pipe to communicate between two processes on the same machine.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_IATTR\_PORT\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & 0\\
  \textbf{Details} & The listening port used by the server or
  node. For nodes, the port can be used to contact pre-started, resident
  render client nodes or to use a specific port for the node. If 0 is
  specified, a random port is chosen. Note that a server with no
  connections automatically creates a default connection using the
  server's default port.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONNECTION\_IATTR\_FILENAME\\
  \textbf{Value} & string\\
  \textbf{Default} & none\\
  \textbf{Details} & The filename of the named pipe used by the server or
  node. The filename has to be unique on the local host.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_NODE\_SATTR\_LAUNCH\_COMMAND\\
  \textbf{Value} & string\\
  \textbf{Default} & ssh -n \%h \%c \textgreater\& \%h.\%n.log [POSIX]\\
                   & ssh -n \%h \%c [WIN32]\\
  \textbf{Details} & The command used by the server to auto-launch nodes
  which could not be connected. The launch command is executed from a
  process forked from the server process. The \% tokens are replaced by
  the server at runtime with concrete data: \%h is replaced by the
  hostname, \%c by the render client command to launch, including command line
  arguments and \%n by a node-unique identifier. Each command line
  argument is surrounded by launch command quotes.\\
  \textbf{See also} & EQ\_NODE\_SATTR\_LAUNCH\_COMMAND\_QUOTE 
  EQ\_NODE\_IATTR\_LAUNCH\_TIMEOUT\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_NODE\_CATTR\_LAUNCH\_COMMAND\_QUOTE\\
  \textbf{Value} & character\\
  \textbf{Default} & ' [POSIX]\\
                   & '' [WIN32]\\
  \textbf{Details} & The server uses command line arguments to launch
  render client nodes correctly. Certain launch commands or shells use
  different conventions to separate command line arguments. These
  arguments might contain white spaces, and therefore have to be
  surrounded by quotes to identify their borders. This option is mostly
  used on Windows.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_NODE\_IATTR\_LAUNCH\_TIMEOUT\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & 60'000 (1 minute)\\
  \textbf{Details} & Defines the timeout in milliseconds to wait for
  an auto-launched node. If the render client process did not contact
  the server within that time, the node is considered to be unreachable
  and the initialization of the configuration fails.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CONFIG\_FATTR\_EYE\_BASE\\
  \textbf{Value} & float\\
  \textbf{Default} & 0.05\\
  \textbf{Details} & The default distance in meters between the left and
  the right eye, i.e., the eye separation. The eye base influences the
  frustum during stereo rendering. See \sref{sTracking} for details.\\
  \textbf{See also} & EQ\_WINDOW\_IATTR\_HINT\_STEREO
  EQ\_COMPOUND\_IATTR\_STEREO\_MODE\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_NODE\_IATTR\_THREAD\_MODEL\\
  \textbf{Value} & ASYNC \textbar \ DRAW\_SYNC \textbar \ LOCAL\_SYNC\\
  \textbf{Default} & DRAW\_SYNC\\
  \textbf{Details} & The threading model for node synchronization. See 
  \sref{sThreadModel} for details.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_PIPE\_IATTR\_HINT\_THREAD\\
  \textbf{Value} & OFF \textbar \ ON\\
  \textbf{Default} & ON\\
  \textbf{Details} & Determines if all task methods for a pipe and its
  children are executed from a separate operating system thread
  (default) or from the node main thread. Non-threaded pipes have
  certain performance limitations and should only be used where necessary.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_PIPE\_IATTR\_HINT\_CUDA\_GL\_INTEROP\\
  \textbf{Value} & OFF \textbar \ ON\\
  \textbf{Default} & OFF\\
  \textbf{Details} & Determines if the pipe thread enables interoperability between CUDA
  and OpenGL or not. When set to ON, the CUDA device is configured on which the 
  thread executes the CUDA device code. If no graphics adapter is specified in the config 
  file, the one providing the highest compute performances on the particular node is chosen. 
  When set to OFF, no CUDA device is configured at all.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_STEREO\\
  \textbf{Value} & OFF \textbar \ ON \textbar \ AUTO\\
  \textbf{Default} & AUTO\\
  \textbf{Details} & Determines if the window selects a quad-buffered
  stereo visual. When set to AUTO, the default window initialization methods try
  to allocate a stereo visual for windows, but fall back to a mono
  visual if allocation fails. For PBuffers, AUTO selects a mono visual.\\
  \textbf{See also} & EQ\_COMPOUND\_IATTR\_STEREO\_MODE\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_DOUBLEBUFFER\\
  \textbf{Value} & OFF \textbar \ ON \textbar \ AUTO\\
  \textbf{Default} & AUTO\\
  \textbf{Details} & Determines if the window selects a double-buffered
  stereo visual. When set to AUTO, the default window initialization methods try
  to allocate a double-buffered visual for windows, but fall back to a
  single-buffered visual if allocation fails. For PBuffers, AUTO selects
  a single-buffered visual.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_DECORATION\\
  \textbf{Value} & OFF \textbar \ ON\\
  \textbf{Default} & ON\\
  \textbf{Details} & When set to OFF, window borders and other
  decorations are disabled, and typically the window cannot be moved or
  resized. This option is useful for source windows during
  decomposition. The implementation is window-system specific.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_FULLSCREEN\\
  \textbf{Value} & OFF \textbar \ ON\\
  \textbf{Default} & OFF\\
  \textbf{Details} & When set to ON, the window displays in
  fullscreen. This option forces window decorations to be OFF. The
  implementation is window-system specific.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_SWAPSYNC\\
  \textbf{Value} & OFF \textbar \ ON\\
  \textbf{Default} & ON\\
  \textbf{Details} & Determines if the buffer swap is synchronized with
  the vertical retrace of the display. This option is currently not
  implemented for GLX. For WGL, the WGL\_EXT\_swap\_control extension is
  required. For optimal performance, set swap synchronization to OFF for
  source-only windows. This option has no effect on single-buffered windows.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_DRAWABLE\\
  \textbf{Value} & window \textbar \ pbuffer \textbar \ FBO\\
  \textbf{Default} & window\\
  \textbf{Details} & Selects the window's drawable type. A window is an
  on-screen, window system-dependent window with a full-window OpenGL
  drawable. PBuffers are off-screen drawables created using window
  system-dependent PBuffer API's. FBO are off-screen frame buffer
  objects. To calculate the PBuffer or FBO size on unconnected devices,
  a pipe viewport size of 4096x4096 is assumed,
  unless specified otherwise using the pipe's viewport parameter.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_STATISTICS\\
  \textbf{Value} & OFF \textbar \ FASTEST [ON] \textbar \ NICEST\\
  \textbf{Default} & FASTEST [Release Build]\\
                   & NICEST [Debug Build]\\
  \textbf{Details} & Determines how statistics are gathered. OpenGL
  buffers commands, which causes the rendering to be executed at an
  arbitrary point in time. Nicest statistics gathering executes a
  \textsf{Window::finish}, which calls by default \textsf{glFinish}, in
  order to accurately account the rendering operations to the sampled
  task method. However, calling \textsf{glFinish} has a performance
  impact. Therefore, the fastest statistics gathering samples the task
  statistics directly, without finishing the OpenGL commands first. Some
  operations, e.g., frame buffer readback, inherently finish all
  previous OpenGL commands.\\
  \textbf{See also} & EQ\_NODE\_IATTR\_HINT\_STATISTICS 
  EQ\_CHANNEL\_IATTR\_HINT\_STATISTICS\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_COLOR\\
  \textbf{Value} & unsigned \textbar \ RGBA16F \textbar \ RGBA32F\\
  \textbf{Default} & AUTO\\
  \textbf{Details} & Determines the number of color planes for the
  window. The interpretation of this value is window system-specific, as
  some window systems select a visual with the closest match to this
  value, and some select a visual with at least the number of color
  planes specified. RGBA16F and RGBA32F select floating point
  framebuffers with 16 or 32 bit precision per component,
  respectively. AUTO selects a visual with a reasonable quality,
  typically eight bits per color.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_ALPHA\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & UNDEFINED\\
  \textbf{Details} & Determines the number of alpha planes for the
  window. The interpretation of this value is window system-specific, as
  some window systems select a visual with the closest match to this
  value, and some select a visual with at least the number of alpha
  planes specified. By default no alpha planes are requested.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_DEPTH\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & AUTO\\
  \textbf{Details} & Determines the precision of the depth buffer. The
  interpretation of this value is window system-specific, as some window
  systems select a visual with the closest match to this value, and some
  select a visual with at least the number of depth bits specified. AUTO
  select a visual with a reasonable depth precision, typically 24 bits.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_STENCIL\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & AUTO\\
  \textbf{Details} & Determines the number of stencil planes for the
  window. The interpretation of this value is window system-specific, as
  some window systems select a visual with the closest match to this
  value, and some select a visual with at least the number of stencil
  planes specified. AUTO tries to select a visual with at least one
  stencil plane, but falls back to no stencil planes if allocation
  fails. Note that for depth-compositing and pixel-compositing at least
  one stencil plane is needed.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_ACCUM\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & UNDEFINED\\
  \textbf{Details} & Determines the number of color accumulation buffer
  planes for the window. The interpretation of this value is window
  system-specific, as some window systems select a visual with the
  closest match to this value, and some select a visual with at least
  the number of accumulation buffer planes specified.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_ACCUM\_ALPHA\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & UNDEFINED\\
  \textbf{Details} & Determines the number of alpha accumulation buffer
  planes for the window. The interpretation of this value is window
  system-specific, as some window systems select a visual with the
  closest match to this value, and some select a visual with at least
  the number of accumulation buffer planes specified. If this attribute
  is undefined, the value of EQ\_WINDOW\_IATTR\_PLANES\_ACCUM is used to
  determine the number of alpha accumulation buffer planes.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_SAMPLES\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & UNDEFINED\\
  \textbf{Details} & Determines the number of samples used for multisampling.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CHANNEL\_IATTR\_HINT\_STATISTICS\\
  \textbf{Value} & OFF \textbar \ FASTEST [ ON ] \textbar \ NICEST\\
  \textbf{Default} & FASTEST [Release Build]\\
                   & NICEST [Debug Build]\\
  \textbf{Details} & See EQ\_WINDOW\_IATTR\_HINT\_STATISTICS.\\
  \textbf{See also} & EQ\_NODE\_IATTR\_HINT\_STATISTICS 
  EQ\_WINDOW\_IATTR\_HINT\_STATISTICS\\
  \hline
\end{tabularx}\\%\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_COMPOUND\_IATTR\_STEREO\_MODE\\
  \textbf{Value} & QUAD \textbar \ ANAGLYPH\\
  \textbf{Default} & QUAD\\
  \textbf{Details} & Selects the algorithm used for stereo
  rendering. QUAD-buffered stereo uses the left and right buffers of a
  stereo window (active stereo). Anaglyphic stereo uses
  \textsf{glColorMask} to mask colors for individual eye passes, used in
  conjunction with colored glasses.\\
  \hline
\end{tabularx}\\%\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_COMPOUND\_IATTR\_STEREO\_ANAGLYPH\_LEFT\_MASK\\
  \textbf{Value} & [ RED GREEN BLUE ]\\
  \textbf{Default} & [ RED ] \\
  \textbf{Details} & Select the color mask for the left eye pass during
  anaglyphic stereo rendering.\\
  \hline
\end{tabularx}\\%\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_COMPOUND\_IATTR\_STEREO\_ANAGLYPH\_RIGHT\_MASK\\
  \textbf{Value} & [ RED GREEN BLUE ]\\
  \textbf{Default} & [ GREEN BLUE ]\\
  \textbf{Details} & Select the color mask for the right eye pass during
  anaglyphic stereo rendering.\\
  \hline
\end{tabularx}\\%\vfill
\end{center}

\subsection{Server Section}

The server section consists of connection description parameters for the
server listening sockets, and a number of configurations for this
server. Currently only the first configuration is used.

\subsubsection{\label{sConnectionDescription}Connection Description}

A connection description defines the network parameters of an Equalizer
process. Currently TCP/IP, SDP, RSP and PIPE connection types are
supported. TCP/IP creates a TCP socket. SDP is very similar, except that the
address family \textsf{AF\_INET\_SDP} instead of \textsf{AF\_INET} is used to
enforce a SDP connection. RSP provides reliable UDP multicast. PIPE uses a named
pipe for fast interprocess communication on Windows.

Note that you can also use the transparent mode provided by most
InfiniBand implementations to use SDP with TCP connections.

Furthermore, a port for the socket can be specified. When no port is
specified for the server, the default port 4242 (+UID on Posix systems)
is used. When no port is specified for a node, a random port will be
chosen by the operating system. For pre-launched render clients, a port
has to be specified for the server to find the client node.

The hostname is the IP address or resolvable host name. A server or
node may have multiple connection descriptions, for example to use a
named pipe for local communications and TCP/IP for remote nodes.

The interface is the IP address or resolvable host name of the adapter to which
multicast traffic is send.

A server listens on all provided connection descriptions. If no hostname
is specified for a server connection description, it listens to
INADDR\_ANY, and is therefore reachable on all network interfaces. If
the server's hostname is specified, the listening socket is bound only
to this address. If any of the given hostnames is not resolvable, or any
port cannot be used, server initialization will fail.

For a node, all connection descriptions are used while trying to
establish a connection to the node. When auto-launched by the server,
all connection descriptions of the node are passed to the launched node
process, which will cause it to bind to all provided descriptions.

{\footnotesize\begin{lstlisting}
server
{
    connection # 0-n times, listening connections of the server
    {
        type       TCPIP | SDP | RSP | PIPE [Win32 only]
        port       unsigned
        filename   string
        hostname   string
        interface  string
    }

    config {...} # 1-n times, currently only the first one is used by the server
}
\end{lstlisting}}

\subsubsection{Config Section}

A configuration has a number of parameters, nodes, observers, layouts,
canvases and compounds.

The nodes and their children describe the rendering resources in a
natural, hierarchical way. Observers, layouts and canvases describe the
properties of the physical projection system. Compounds use rendering
resources (channels) to execute rendering tasks.

For an introduction to writing configurations and the concepts of the
configuration entities please refer to \sref{sConfig}.

The latency of a config defines the maximum number of frames the
slowest operation may fall behind the application thread. A latency of 0
synchronizes all rendering tasks started by \textsf{Config::startFrame}
in \textsf{Config::finishFrame}. A latency of one synchronizes all
rendering tasks started one frame ago in \textsf{finishFrame}.

For a description of config attributes please refer to \sref{sGlobal}.

{\footnotesize\begin{lstlisting}
config
{
    latency int    # Nr. of frames nodes may fall behind application, default 1
    attributes
    {
         eye_base float # distance between left and right eye
    }

    appNode {...}  # 0|1 times, launches render threads within app process
    node {...}     # 1-n times, a system in the cluster
    observer {...} # 0-n times, one tracked entity
    layout {...}   # 0-n times, logical views on canvases
    canvas {...}   # 0-n times, physical segment layout of a projection surface
    compound {...} # 0-n times, rendering description
}
\end{lstlisting}}

\subsubsection{Node Section}

A node represents a machine in the cluster, and is one process. It has a
name, a number of connection descriptions and at least one pipe. The
name of the node can be used for debugging, it has no influence on the
execution of Equalizer. For a description of node and connection
attributes please refer to \sref{sGlobal}.

{\footnotesize\begin{lstlisting}
node
{
    name     string
    connection # 0-n times, possible connections to this node
    {
        type          TCPIP | SDP | PIPE
        TCPIP_port    unsigned
        PIPE_filename string
        hostname      string
        command       string      # render client launch command
        command_quote 'character' # launch command argument quote char
        timeout       unsigned    # timeout in milliseconds for launch
    }
    attributes
    {
        thread_model    ASYNC | DRAW_SYNC | LOCAL_SYNC
    }
    pipe {...} # 1-n times
}
\end{lstlisting}}

\subsubsection{Pipe Section}

A pipe represents a graphics card (GPU), and is one execution thread. It
has a name, GPU settings and attributes. The name of a pipe can be used
for debugging, it has no influence on the execution of Equalizer.

The GPU is identified by two parameters, a port and a device. The port
is only used for the GLX window system, and identifies the port number
of the X server, i.e., the number after the colon in the DISPLAY
description (':\textbf{0}.1').

The device identifies the graphics adapter. For the GLX window system
this is the screen number, i.e., the number after the dot in the DISPLAY
description (:0.\textbf{1}). The OpenGL output is always restricted by
glX to the GPU attached to selected screen.

For the AGL window system, the device selects the \textit{n}th display
in the list of online displays. The OpenGL output is optimized for the
selected display, but not restricted to the attached GPU.

For the WGL window system, the device selects the \textit{n}th GPU in
the system. The GPU can be offline, in this case only PBuffer windows
can be used. To restrict the OpenGL output to the GPU, the
\textsf{WGL\_NV\_gpu\_affinity} extension is used. If the extension is
not present, the window is opened on the \textit{n}th monitor, but
OpenGL commands are sent to all GPU's.

The viewport of the pipe can be used to override the pipe
resolution. The viewport is defined in pixels. The x and y parameter of
the viewport are currently ignored. The default viewport is
automatically detected. For offline GPU's, a default of 4096x4096 is
used.

For a description of pipe attributes please refer to \sref{sGlobal}.

{\footnotesize\begin{lstlisting}
pipe
{
    name     string
    port     unsigned     # X server number or ignored
    device   unsigned     # graphics adapter number
    viewport [ viewport ] # default: autodetect
    attributes
    {
        hint_thread            OFF | ON   # default ON
        hint_cuda_GL_interop   OFF | ON   # default OFF
    }

    window {...} # 1-n times
}
\end{lstlisting}}

\subsubsection{Window Section}

A window represents an OpenGL drawable and holds an OpenGL context. It
has a name, a viewport and attributes. The name of a window can be used
for debugging, it has no influence on the execution of Equalizer, other
then it being used as the window title by the default window creation
methods.

The viewport of the window is relative to the pipe. It can be specified
in relative or absolute coordinates. Relative coordinates are normalized
coordinates with respect to the pipe, e.g., a viewport of \textsf{[ 0.25
  0.25 0.5 0.5 ]} creates a window in the middle of the screen, using
50\% of the pipe's size. Absolute coordinates are integer pixel values,
e.g., a viewport of \textsf{[ 50 50 800 600 ]} creates a window 50 pixels
from the upper-left corner, sized 800x600 pixels, regardless of the
pipe's size. The default viewport is \textsf{[ 0 0 1 1 ]}, i.e., a
full-screen window.

For a description of window attributes please refer to \sref{sGlobal}.

{\footnotesize\begin{lstlisting}
window
{
    name     string
    viewport [ viewport ] # wrt pipe, default full screen

    attributes
    {
        hint_stereo         OFF | ON | AUTO
        hint_doublebuffer   OFF | ON | AUTO
        hint_decoration     OFF | ON
        hint_fullscreen     OFF | ON
        hint_swapsync       OFF | ON         # AGL, WGL only
        hint_drawable       window | pbuffer | FBO
        hint_statistics     off | fastest [on] | nicest
        planes_color        unsigned | RGBA16F | RGBA32F
        planes_alpha        unsigned
        planes_depth        unsigned
        planes_stencil      unsigned
        planes_accum        unsigned
        planes_accum_alpha  unsigned
        planes_samples      unsigned
    }

    channel {...} # 1-n times
}
\end{lstlisting}}

\subsubsection{Channel Section}

A channel is a two-dimensional area within a window. It has a name,
viewport and attributes. The name of the channel is used to identify the
channel in the respective segments or compounds. It should be unique
within the config.

The viewport of the channel is relative to the window. As for windows,
it can be specified in relative or absolute coordinates. The default
viewport is \textsf{[ 0 0 1 1 ]}, i.e., fully covering its window.

The channel can have an alternate drawable description. Currently, the
window's framebuffer can be replaced by framebuffer objects bound to the
window's OpenGL context. The window's default framebuffer can be
partially overwritten with framebuffer objects.

For a description of channel attributes please refer to \sref{sGlobal}.

{\footnotesize\begin{lstlisting}
channel
{
    name     string
    viewport [ viewport ] #wrt window, default full window
    drawable [ FBO_COLOR FBO_DEPTH FBO_STENCIL ]

    attributes
    {
        hint_statistics     OFF | FASTEST [ON] | NICEST
    }
}
\end{lstlisting}}

\subsubsection{Observer Section}

An observer represents a tracked entity, i.e, one user. It has a name
and an eye separation. The name of an observer can be used for
debugging, it has no influence on the execution of Equalizer. It can be
used to reference the observer in views, in which case the name should
be unique. Not all views have to be tracked by an observer.

{\footnotesize\begin{lstlisting}
observer # 0...n times
{
    name    string
    eyeBase float
}
\end{lstlisting}}

\subsubsection{Layout Section}

A layout represent a set of logical views on one or more canvases. It
has a name and child views. The name of a layout can be used for
debugging, it has no influence on the execution of Equalizer. It can be
used to reference the layout, in which case the name should be unique.

A layout is applied to a canvas. If no layout is applied to a canvas,
nothing is rendered on this canvas, i.e, the canvas is inactive.

The layout assignment can be changed at runtime by the application. The
intersection between views and segments defines which output
(sub-)channels are available. These output channels are typically used
as destination channels in a compound. They are automatically created
during configuration loading or creation.

{\footnotesize\begin{lstlisting}
layout  # 0...n times
{
    name string
    view {...}  # 1...n times
}
\end{lstlisting}}

\subsubsection{View Section}

A view represents a 2D area on a canvas. It has a name, viewport,
observer and frustum. The name of a view can be used for debugging, it
has no influence on the execution of Equalizer. It can be used to
reference the view, in which case the name should be unique.

A view can have a frustum description. The view's frustum overrides
frusta specified at the canvas or segment level. This is typically used
for nonphysically correct rendering, e.g., to compare two models
side-by-side. If the view does not specify a frustum, the corresponding
destination channels will use the sub-frustum resulting from the
view/segment intersection.

A view is a view on the application's model, in the sense used by the
Model-View-Controller pattern. It can be a scene, viewing mode, viewing
position, or any other representation of the application's data.

{\footnotesize\begin{lstlisting}
view  # 1...n times
{
    name     string
    observer observer-ref
    viewport [ viewport ]

    wall       // frustum description
    {
        bottom_left  [ float float float ]
        bottom_right [ float float float ]
        top_left     [ float float float ]
        type         fixed | HMD
    }
    projection // alternate frustum description, last one wins
    {
        origin       [ float float float ]
        distance     float
        fov          [ float float ]
        hpr          [ float float float ]
    }
}
\end{lstlisting}}

\subsubsection{Canvas Section}

A canvas represents a logical projection surface of multiple
segments. It has a name, frustum, layouts, and segments. The name of a
canvas can be used for debugging, it has no influence on the execution
of Equalizer. It can be used to reference the canvas, in which case the
name should be unique.

Each canvas is made of one or more segments. Segments can be planar or
non-planar to each other, and can overlap or have gaps between each
other. A canvas can define a frustum, which will create default, planar
sub-frusta for its segments.

The layouts referenced by the canvas can be applied to it by the
application at runtime. One layout can be referenced by multiple
canvases. The first layout is the layout active by default.

{\footnotesize\begin{lstlisting}
canvas  # 0...n times
{
    name    string
    layout  layout-ref   # 1...n times

    wall       
    {
            bottom_left  [ float float float ]
            bottom_right [ float float float ]
            top_left     [ float float float ]
            type         fixed | HMD
    }
    projection
    {
            origin       [ float float float ]
            distance     float
            fov          [ float float ]
            hpr          [ float float float ]
    }

    segment {...}   # 1...n times
}
\end{lstlisting}}

\subsubsection{Segment Section}

A segment represents a single display, i.e., a projector or monitor. It
references a channel, has a name, viewport and frustum. The name of a
segment can be used for debugging, it has no influence on the execution
of Equalizer. It can be used to reference the segment, in which case the
name should be unique.

The channel referenced by the segment defines the output channel. The
viewport of the section defines the 2D area covered by the channel on
the canvas. Segments can overlap each other, e.g., when edge-blended
projectors or passive stereo is used.

A segment can define a frustum, in which case it overrides the default
frustum calculated from the canvas frustum and segment viewport.

{\footnotesize\begin{lstlisting}
segment # 1...n times
{
    channel  string
    name     string
    viewport [ viewport ]

    wall       // frustum description
    {
        bottom_left  [ float float float ]
        bottom_right [ float float float ]
        top_left     [ float float float ]
        type         fixed | HMD
    }
    projection // alternate frustum description, last one wins
    {
        origin       [ float float float ]
        distance     float
        fov          [ float float ]
        hpr          [ float float float ]
    }
}
\end{lstlisting}}

\subsubsection{Compound Section}

Compounds are the basic data structure describing the rendering
setup. They use channels for rendering. Please refer to
\sref{sCompounds} for a description of compound operation logics.

The name of the compound is used for the default names of swap barriers
and output frames.

A channel reference is either the name of the channel in the resource
section if no canvases are used, or the destination channel reference of
a view/segment intersection. Channel segment references are delimited by
braces, in which the canvas, segment, layout and view describing the
channel are named, i.e, 'channel ( canvas "PowerWall" segment 0 layout
"Simple" view 0 )'.

Compound tasks describe the operations the compound executes. The
default is all tasks for compounds with no children (leaf compounds) and
\textsf{READBACK ASSEMBLE} for all others. The readback and assemble
tasks are only executed if the compound has output frames or input
frames, respectively. Tasks are not inherited by the children of a
compound.

The buffer defines the default frame buffer attachments read back by
output frames. Output frames may change the buffer attachments used.

The viewport restricts the rendering to the area relative to the parent
compound. The range restricts the database range, relative to the
parent. The pixel setting selects the pixel decomposition kernel,
relative to the parent. The subpixel defines the jittering applied to
the frustum, relative to the parent. The eye pass select the mono, left
or right eye passes rendered by the compound. The zoom scales the parent
pixel viewport resolution. The DPlex period defines that $\frac{1}{period}$
frames are rendered, and the phase defines when in the period the
rendering starts. All these attributes are inherited by the children of
a compound. Viewport, range, pixel and period parameters are cumulative.

Equalizers are used to automatically optimize the decomposition. A 2D,
horizontal or vertical load equalizer adjusts the viewport of all direct
children of the compound each frame. A DB load equalizer adjusts the
range of all direct children. A dynamic framerate (DFR) equalizer adjusts
the zoom for a constant framerate. A framerate equalizer
smoothens the framerate of the compound's window to produce a steady
output framerate, typically for DPlex compounds. A monitor equalizer
adjusts the output image zoom to monitor another canvas.

For a description of compound attributes please refer to \sref{sGlobal}.

A wall or projection description is used to define the view frustum of
the compound. The frustum is inherited and typically only defined on the
topmost compound. The last specified frustum description is used. Sizes
are specified in meters. \fig{fFrusta} illustrates the frustum
parameters.

A swap barrier is used to synchronize the output of multiple
compounds. All swap barriers of the same name are synchronized The
default name is \textsf{barrier[.root\-Com\-pound\-Name]}. All windows
using the same \textsf{NV\_group} on a single system are synchronized
with each other using hardware synchronization. All groups using the
same \textsf{NV\_barrier} across systems are synchronized with each
other using hardware synchronization. When using hardware
synchronization, the barrier name is ignored.

Output frames transport frame buffer contents to input frames of the
same name. If the compound has a name, the default frame name is
\textsf{frame.compoundName}, otherwise the default name is
\textsf{frame.channelName}. The frame buffer attachments to read back
are inherited from the compound, but can be overridden by output
frames. Frames of type \textsf{texture} copy the framebuffer contents to
a texture, and can only be used to composite frames between windows of
the same pipe.

{\footnotesize\begin{lstlisting}
compound
{
    name     string
    channel  channel-ref  # where the compound tasks are executed

    task     [ CLEAR DRAW READBACK ASSEMBLE ]
    buffer   [ COLOR DEPTH ]         # default COLOR

    viewport [ viewport ]            # wrt parent compound, sort-first
    range    [ float float ]         # DB-range for sort-last
    pixel    [ int int int int ]     # pixel decomposition (x y w h)
    subpixel [ int int ]	     # subpixel decomposition (index size)
    eye      [ CYCLOP LEFT RIGHT ]   # monoscopic or stereo view
    zoom     [ float float ]         # up/downscale of parent pvp
    period   int                     # DPlex period, every nth frame rendered
    phase    int                     # DPlex phase, when to start rendering

    load_equalizer # adjust 2D tiling or DB range of children
    {
        mode 2D | DB | VERTICAL | HORIZONTAL
        damping float # 0: no damping, 1: no changes
        boundary [ x y ] | float  # x,y: tile boundary for 2D, float: range
    }
    DFR_equalizer # adjust ZOOM to achieve constant framerate
    {
        framerate float # target framerate
        damping float   # 0: no damping, 1: no changes
    }
    framerate_equalizer {} # smoothen window swapbuffer rate (DPlex)
    monitor_equalizer {}   # set frame zoom when monitoring other views

    attributes
    {
        stereo_mode                  QUAD | ANAGLYPH    # default QUAD
        stereo_anaglyph_left_mask    [ RED GREEN BLUE ] # default red
        stereo_anaglyph_right_mask   [ RED GREEN BLUE ] # df green blue
    }

    wall       # frustum description
    {
        bottom_left  [ float float float ]
        bottom_right [ float float float ]
        top_left     [ float float float ]
        type         fixed | HMD
    }
    projection # alternate frustum description, last one wins
    {
        origin       [ float float float ]
        distance     float
        fov          [ float float ]
        hpr          [ float float float ]
    }

    child-compounds

    swapbarrier  # compounds with the same barriername sync swap
    {
        name string
        NV_group OFF | ON | unsigned
        NV_barrier OFF | ON | unsigned
    }
    outputframe
    {
        name   string
        buffer [ COLOR DEPTH ]
        type texture | memory
    }
    inputframe
    {
        name string # corresponding output frame
    }
}

channel-ref:  'string' | '(' channel-segment-ref ')'
channel-segment-ref: ( canvas-ref ) segment-ref ( layout-ref ) view-ref
canvas-ref:   'string' | 'index'
segment-ref:  'string' | 'index'
layout-ref:   'string' | 'index'
view-ref:     'string' | 'index'
observer-ref: 'string' | 'index'
\end{lstlisting}}

\if 0
{\footnotesize\begin{lstlisting}
\end{lstlisting}}
\fi

\cleardoublepage
\pagestyle{empty}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaserBack.pdf}
\end{textblock}

\end{document}